### 2018-03-21 2018-tsd-brno/webapp.tex
init

\documentclass{llncs}
%\usepackage[cp1250]{inputenc}
%\usepackage[utf8]{inputenc}
%\usepackage[czech]{babel}
\usepackage{graphicx}
%\usepackage{tipa}

\bibliographystyle{splncs}

\begin{filecontents}{citace.bib}
@inproceedings{kruuza2012making,
  title={Making Community and ASR Join Forces in Web Environment},
  author={Kr{\uu}za, Old{\v{r}}ich and Peterek, Nino},
  booktitle={International Conference on Text, Speech and Dialogue},
  pages={415--421},
  year={2012},
  organization={Springer}
}
\end{filecontents}

\begin{document}
\newtheorem{Definition}{Definition}
\title{ASR-Output-Correcting Webapp After 6 Years}

\author{Oldřich Krůza and Vladislav Kuboň}
\institute{Charles University in Prague\\
           Faculty of Mathematics and Physics\\
           Institute of Formal and Applied Linguistics\\
           Malostranské nám. 25, Prague, Czech Republic\\
           \{kruza,vk\}@ufal.mff.cuni.cz}

\maketitle

\begin{abstract}

This paper presents a next-generation web application that enables users to
contribute corrections to automatically acquired transcription of long speech
recordings. We describe differences from similar settings, compare our solution
with others and reflect on the development from the first, 6 years old version
in the light of the work done, lessons learned and the new technologies
available in the browser.

\end{abstract}

\section{Introduction}

In 2012\cite{kruuza2012making}, we presented a setting where a community of
users contributed corrections to automatically transcribed talks of a single
speaker. Now that the browser technologies evolved drastically and we could
observe the usage patterns and discover shortcomings of the solution at hand, we
have created a next generation of the programme. We shall describe the steps
taken and discuss their motivation and impact.

\section{}

\section{Conclusion}

\section*{Acknowledgments}

The research was supported by SVV project number 265 314.\\
\\
This work has been using language resources stored
by the LINDAT-Clarin project of the Ministry of
Education of the Czech Republic (project LM2010013).
%\bibliography{paper}

\bibliography{citace}

\end{document}

### 2018-03-22 2018-tsd-brno/webapp.tex
tsd2018: writing

@@
 
-\section{}
+\section{Differences to other settings}
+
+In our setting, we have a large spoken corpus (about 1000 hours) of a single
+speaker. Our aim is to have a transcription as good as possible for the purpose
+of searching and further, higher-level processing of the data. There is a pool
+of people interested in the talks, who on one hand are the force we can try to
+employ and on the other hand are the consumers of our effort, our target group
+so to speak.
+
+The web application should therefore combine the two purposes: 1. serve its user
+with making the content available in a manner as good as possible and 2. animate
+the user to give as much and as high-quality contribution as possible.
+
+To our best knowledge, there is no other project with a comparable setting.
+However, we can compare single aspects found in other applications.
+
+\subsection{Transcribing apps}
+
+The main differences to common transcribing software are that
+\begin{enumerate}
+\item{they are optimized for the case where there is no transcription available
+and it must be acquired from scratch;}
+\item{no quality control: the user is free to enter whatever transcription she
+pleases and the ultimate measure is her content; in contrast, we need the
+transcription to be accurate because it is used as training data for the
+acoustic model;}
+\item{the alignment is on the level of phrases, if any; we use alignment on the
+level of phones;}
+\item{standalone transcribing applications are user-centric: the user
+transcribes whatever acoustic data they choose; our system is data-centric: the
+whole application with all its tools and persons revolves around the data set;}
+\item{common transcription software assumes the user wants to transcribe; we
+assume the user want to listen and possibly read along and we want to animate
+her to submit transcriptions.}
+\item{user collision}
+\end{enumerate}
+
+Our first reference is
+Transcriber\footnote{sourceforge.net/projects/trans}. Transcriber is an
+open-source program written in TCL.
 

### 2018-03-22 2018-tsd-brno/webapp.tex
tsd2018: more text

--- 2018-03-21 2018-tsd-brno/webapp.tex
@@
 \usepackage{graphicx}
+\usepackage{tabularx}
 %\usepackage{tipa}
@@
 
-\section{}
+\section{Differences to other settings}
+
+In our setting, we have a large spoken corpus (about 1000 hours) of a single
+speaker. Our aim is to have a transcription as good as possible for the purpose
+of searching and further, higher-level processing of the data. There is a pool
+of people interested in the talks, who on one hand are the force we can try to
+employ and on the other hand are the consumers of our effort, our target group
+so to speak.
+
+The web application should therefore combine the two purposes: 1. serve its user
+with making the content available in a manner as good as possible and 2. animate
+the user to give as much and as high-quality contribution as possible.
+
+To our best knowledge, there is no other project with a comparable setting.
+However, we can compare single aspects found in other applications.
+
+\subsection{Transcribing apps}
+
+The main differences to common transcribing software are that
+
+\noindent
+\begin{tabularx}{\textwidth}{
+    @{\hspace{1.5em}}% Space for left bullet
+    >{\leavevmode\llap{\textbullet~}\raggedright}% Left bullet + formatting of column
+    X% Left column specification
+    @{\quad\hspace{1.5em}}% Space between columns + right bullet space
+    >{\leavevmode\llap{\textbullet~}\raggedright\arraybackslash}% Right bullet + formatting of column
+    X% Right column specification
+    @{}% No column space on right
+  }
+  \em{common transcription applications} & \em{our application} \\
+  are optimized for the case where there is no transcription available and it
+  must be acquired from scratch &
+    always assumes a transcription is available \\
+  need no quality control: the user is free to enter whatever transcription she
+  pleases and the ultimate measure is her satisfaction &
+    needs the transcription to be accurate because it is used as training data
+    for the acoustic model \\
+  use alignment on the level of phrases, if any &
+    uses alignment on the level of words \\
+  are user-centric: the user transcribes whatever acoustic data they choose &
+    is data-centric: the whole application with all its tools and persons
+    revolves around the data set \\
+  assumes the user wants to transcribe &
+    assume the user want to listen and possibly read along and we want to
+    animate her to submit transcriptions \\
+  has no shared data between users &
+    must count with collisions
+\end{tabularx}
+
+We can still learn a lot from transcribing software. The ease of performing
+common tasks, like pausing, resuming and rewinding is crucial for the user
+experience and in effect for the amount of submissions that we receive. Also,
+the way the text is displayed synchronously to the audio played has a big impact
+and the approaches have a lot of space for variation.
+
+\subsection{Wiki}
+
+Where our application diverts from transciption software, it mostly resembles a
+wiki: a community platform that serves its users including the contributors but
+where the quality of the contributions is essential, while the contributor's
+satisfaction alone is of little importance.
+
+%The need for quality control has wide implications. We need to perform forced
+%alignment every time a segment is transcribed. Hence, it must be clear what part
+%of the audio \em{exactly} the segment corresponds to. Since each word knows its
+%exact timestamp, and we know which words from the old transcription are to be
+%replaced, we can perform the forced alignment very precisely. This is also one
+%of the reasons why our application always assumes an existing transcription.
+
+%What concerns animating the user to provide a transcription, there must be more
+%research done. The basic idea is though that when the user sees the
+%transcription synchronously and it contains errors, they get the urge to fix it.
+
+%Our first reference is
+%Transcriber\footnote{sourceforge.net/projects/trans}. Transcriber is an
+%open-source program written in TCL.
 

### 2018-03-23 2018-tsd-brno/webapp.tex
motivation

@@
 taken and discuss their motivation and impact.
+
+\subsection{Motivation}
+
+Our project focuses on the collection of recordings of Karel Mako\v{n} *1912
+\textdagger1993, the author of numerous books, translations and comments to
+works of spiritual and religious nature, who was influenced by trances during
+recurring surgery without anesthesy in the age of 6, ecstasies in the youth and
+finally facing and surviving certain death in a Nazi concentration camp, after
+which he experienced enlightenment. He gave talks in a narrow circle of friends
+and the recordings in our care have been taken between early 70's and 1991,
+spanning about 1000 hours in total.
+
+All of his work deals more or less directly with a single topic: entering the
+eternal life before the physical death. He draws mainly from the Christian
+symbolism, builds up on Christian mysticism and ancient tradition of India.
+
+Basically the meat of his teachings is systematically encompassed in his written
+works, whereas the talks contain the sauce: talks tailored to the audience,
+answering questions, personal experiences, behind-the-scenes to the books etc.
 
@@
 
+One major difference to a wiki is that wiki is creative, whereas our task is
+mechanical. The user has basically no room for their own invention: providing a
+different than correct transcription is seen as an error.
+
+Major wikis have good measures for edit conflicts, which is where we could learn
+some lessons. However, so far there was no need to do that because
+\begin{enumerate}
+\item{when we always simply take the most recent version
+of a segment, the result stays consistent even if a piece from user A comes into
+a larger transcription of user B;}
+\item{our user base is so far limited to a small community who have no problem
+coordinating with each other. We plan to expand to broader public soon though.}
+\end{enumerate}
+
+% alert when correcting humanic
+
 %The need for quality control has wide implications. We need to perform forced

### 2018-03-23 2018-tsd-brno/webapp.tex
interface desc

@@
     always assumes a transcription is available \\
+  allow annotation of speakers &
+    assumes all utterances come from the same speaker \\
   need no quality control: the user is free to enter whatever transcription she
@@
 
+\section{Description of the web application}
+
+The application consists of several views:
+\begin{enumerate}
+\item{the start page where all recordings are listed and each points to a detail
+view,}
+\item{the detail view, where a recording can be played back, its transcription
+is displayed and can be corrected by the user,}
+\item{search page, where hits to a search query are listed and point to
+corresponding positions in the recordings,}
+\item{static pages with general information, contact etc.}
+\end{enumerate}
+
+We shall only discuss the detail view as the others are not relevant to this
+article. Figure~\ref{fig:scn1lab} shows the interface during playback.
+Figure~\ref{fig:scn2lab} shows the interface while a segment is being edited.
+The interface in the figures is conveniently shown in English, although in
+reality it is in Czech.
+
+\begin{figure}[htpb]
+\includegraphics[scale=0.6]{rc/radio-makon-en-1-lab.png}
+\caption{Web interface during playback}
+\label{fig:scn1lab}
+\end{figure}
+
+Legend to Figure~\ref{fig:scn1lab}:
+\begin{enumerate}
+\item{
+    Header with
+    \begin{itemize}
+    \item{app name linking to start page,}
+    \item{about link,}
+    \item{search field and}
+    \item{username input field;}
+    \end{itemize}
+}
+\item{Identifier of the recording;}
+\item{Automatically transcribed segments in gray;}
+\item{Manually transcribed segments in black;}
+\item{Currently played-back word highlighted by yellow background;}
+\item{Marked word highlighted in regent st. blue;}
+\item{
+    Marked word info:
+    \begin{itemize}
+    \item{
+        occurrence: the word with contextual capitalization and
+        punctuation as it appeard in the text, currently being edited as the
+        selected initial letter reveals,
+    }
+    \item{form: normalized word form as it appears in the word list,}
+    \item{pronunciation: Czech phonetic transcription of the word,}
+    \item{
+        position: time of the beginning of the word in seconds from the
+        start of the recording;
+    }
+    \end{itemize}
+}
+\item{
+    Tools for storing:
+    \begin{itemize}
+    \item{direct links to the audio files,}
+    \item{selecting the whole transcription for easy pasting,}
+    \item{storing the decoded recording in the browser's IndexedDB;}
+    \end{itemize}
+}
+\item{Graphical equalizer for compensating narrow-band noise;}
+\item{
+    Audio playback controls:
+    \begin{itemize}
+    \item{play/pause button,}
+    \item{current playback position,}
+    \item{playback scrollbar,}
+    \item{total recording length;}
+    \end{itemize}
+}
+\item{Current position reflected in URL fragment.}
+\end{enumerate}
+
+\begin{figure}[htpb]
+\includegraphics[scale=0.6]{rc/radio-makon-en-2-lab.png}
+\caption{Interface in the state of editing a segment, with labels}
+\label{fig:scn2lab}
+\end{figure}
+
+Legend to Figure~\ref{fig:scn2lab}:
+\begin{enumerate}
+\item{
+    Selecting a text range with the mouse defines the segment the user is about
+    to transcribe;
+}
+\item{
+    The edit tool with
+    \begin{itemize}
+    \item{textarea prefilled with the current transcription,}
+    \item{playback button that plays the corresponding segment,}
+    \item{save button and}
+    \item{download-segment button, which initiates a file-save action for the
+    corresponding segment.}
+    \end{itemize}
+}
+\end{enumerate}
+
+\subsection{Displaying the transcription}
+
+Most transcribing programs show the transcription as a vertical list of
+utterances, see Figure~\ref{fig:transcriber1} for an example of
+{\em{Transcriber}}\footnote{trans.sourceforge.net}, a veteran open-source
+transcription tool. We attribute this to the fact that the atomic elements of
+the transcription are the user-entered utterances and their boundaries are
+reliable. In our case, the atomic elements are words. There are sentences, sure,
+but the segmentation to sentences by the ASR is very unreliable, so we want it
+to be natural to transcribe a segment overlapping sentence boundaries.
+
+This is one of the reasons why we display the transcription basically as a
+single wrapped line.
+
+The transcription display was designed to have these features:
+\begin{itemize}
+\item{Currently played-back word should be highlighted;}
+\item{
+    Manually transcribed segments should be clearly distinct from automatically
+    transcribed ones;
+    \label{feats:item:manualdistinct}
+}
+\item{
+    Selecting one or more words with the mouse should trigger transcription mode
+    for the selected text;
+    upon a successful save, this should be merged into the display;
+}
+\item{
+    Clicking a word should bring up its context info (we call this the
+    \em{marked word} as the term \em{selected word} is already taken);
+}
+\item{
+    The whole transcription should be shown at once for easy searching;
+    \label{feats:item:showall}
+}
+\item{
+    The page should be responsive;\label{feats:item:speed}
+}
+\end{itemize}
+
+These requirements are harder to combine than it may seem. Notably
+responsiveness is hard to combine with all of the other ones. Why is that so?
+
+\begin{figure}[htpb]
+\includegraphics[scale=0.4]{rc/transcriber1.png}
+\caption{A screenshot of Transcriber}
+\label{fig:transcriber1}
+\end{figure}
+
 % alert when correcting humanic

### 2018-03-24 2018-tsd-brno/webapp.tex
s/transcribing software/transcription software/

@@
 
-\subsection{Transcribing apps}
+\subsection{Transcription apps}
 
-The main differences to common transcribing software are that
+The main differences to common transcription software are that
 
@@
 
-We can still learn a lot from transcribing software. The ease of performing
+We can still learn a lot from transcription software. The ease of performing
 common tasks, like pausing, resuming and rewinding is crucial for the user
@@
 
-Most transcribing programs show the transcription as a vertical list of
+Most transcription programs show the transcription as a vertical list of
 utterances, see Figure~\ref{fig:transcriber1} for an example of

### 2018-03-24 2018-tsd-brno/webapp.tex
more text

@@
 %\usepackage{tipa}
+\usepackage{enumitem}
 
@@
 
-\author{Oldřich Krůza and Vladislav Kuboň}
+\author{Old\v{r}ich Kr\r{u}za and Vladislav Kubo\v{n}}
 \institute{Charles University in Prague\\
@@
 
+\subsubsection{Performance challenge}
+
 The transcription display was designed to have these features:
-\begin{itemize}
-\item{Currently played-back word should be highlighted;}
+\begin{enumerate}
+\item{
+    Currently played-back word should be highlighted;
+    \label{feats:item:curword}
+}
 \item{
@@
     upon a successful save, this should be merged into the display;
+    \label{feats:item:selectable}
 }
@@
     Clicking a word should bring up its context info (we call this the
-    \em{marked word} as the term \em{selected word} is already taken);
+    {\em marked word} as the term {\em selected word} is already taken);
+    \label{feats:item:clickable}
 }
@@
 }
-\end{itemize}
+\end{enumerate}
 
@@
 responsiveness is hard to combine with all of the other ones. Why is that so?
+
+Points~\ref{feats:item:curword} through \ref{feats:item:clickable}
+call for every word to be wrapped in its own element.
+Point~\ref{feats:item:showall} and the median count of words in a transcript of
+about 6000 yield 6000 \texttt{<span>} elements just to show the text.
+
+Although this may not seem like a big deal, it does make affect the
+responsiveness and memory footprint of the page.
+
+Kr\r{u}za 2012~\cite{kruuza2012making} solve this by sacrificing point~5:
+% TODO obrázek MakonFM
+only 3
+lines of text are shown with the current word kept on the middle line. Thanks to
+the development in the web standards and their support from popular browsers, a
+solution is possible.
+
+\subsubsection{Solution}
+
+We can use the fortunate fact that manually transcribed words and automatically
+transcribed ones tend to form larger chunks. The average number of words per
+submitted segment is 7.9. Furthermore, the absolute majority of such segments
+are adjacent to other manually transcribed chunks. 
+% TODO max fragmentation
+Hence, wrapping each chunk of consecutive manually or automatically transcribed
+words in an HTML element is no problem, which solves
+point~2.
+
+Point~\ref{feats:item:selectable} can be implemented using
+\texttt{document.selection} and the \texttt{Range} objects, which let us find
+out the innermost HTML element and text offset of the start and end of the
+textual selection. Since we know the length of the words, this allows us to map
+the selection to the corresponding words in the transcription.
+
+Points~\ref{feats:item:curword} and~\ref{feats:item:clickable} can be
+implemented in two ways: We could either wrap the current and marked word in a
+dedicated element or we could draw a highlighting rectangle beneath the word.
+
+Wrapping the word would definetely be more robust and less error-prone but the
+constant changes in the DOM during playback with possible frequent reflows speak
+against it. Finding the exact position of each word and drawing a rectangle
+precisely beneath it (beneath on the z-axis; over it in the x-y sense), avoiding
+positioning issues and keeping the rectangle position synced even after scrolling
+/ window resizing is definitely a challenge but we chose this way nonetheless.
+The performance gain for the majority of the usage time outweighs the possible
+errors in the corner cases,
 

### 2018-03-24 2018-tsd-brno/webapp.tex
a bit of bullshit

@@
 
+\begin{figure}[htpb]
+\includegraphics[scale=0.4]{rc/transcriber1.png}
+\caption{A screenshot of Transcriber}
+\label{fig:transcriber1}
+\end{figure}
+
 This is one of the reasons why we display the transcription basically as a
@@
 The performance gain for the majority of the usage time outweighs the possible
-errors in the corner cases,
+errors in the corner cases, moreso since the eventual errors are not critical
+and mostly remedied by further playback.
 
-\begin{figure}[htpb]
-\includegraphics[scale=0.4]{rc/transcriber1.png}
-\caption{A screenshot of Transcriber}
-\label{fig:transcriber1}
-\end{figure}
+The efficiency of repositioning a rectangle is supported by the fact that we can
+calculate the coordinates of all rendered words once and only recalculate them
+in two cases: 1) In the rare event of screen resize and 2) when a corrected
+segment is merged into the transcription, in which case we only need to
+recalculate for the words further in the document.
+
+\subsection{Implementation Details}
+
+\subsubsection{Audio Engine}
+
+The adoption of Web Audio API allowed for big improvements in comparison with
+the original implementation described by Kruza 2012~\cite{kruuza2012making}.
+There are four major differences between using the HTML \texttt{<audio>} tag
+and the Web Audio API.
+
+\begin{itemize}
+\item{It is now possible to precisely replay the selected audio span.}
+\item{
+    We could implement a graphical equalizer. Some recordings suffer from loud
+    noise in the low frequency spectrum. A systematic approach to acoustic
+    normalising of the material is a point of future work. Until that, the
+    equalizer is a huge relief for the users.
+}
+\item{
+    Thanks to the \texttt{OfflineAudioContext}, it is possible to store the
+    recording in the browser's storage and avoid downloading or decoding it
+    again after reload.
+}
+\item{
+    We have also implemented saving the audio corresponding to the selected text
+    segment as a sound file.
+}
+\end{itemize}
+
+\subsubsection{App State Management}
+
+We use React as the view library and Redux for state management. The good thing
+about Redux is that it makes it easy to keep minimal state as the single source
+of truth and everything that can be computed is computed, while avoiding
+needless calculations. This is of course nothing new -- basically it is what we
+know from database design as the normal representation. Also in our case, this
+approach makes the program more predictable, less error-prone and, as the modern
+programming jargon lovingly expresses, {\em easier to reason about}.
+
+We are squeezing as much juice from these fruits as we can.
+
+Among the states the app can enter is simple playback, inspecting a word,
+transcribing a segment and replaying the selected audio span. Now, the only
+thing we actually store is the sub
 

### 2018-03-25 2018-tsd-brno/webapp.tex
eh nix
bla bla redux

@@
 needless calculations. This is of course nothing new -- basically it is what we
-know from database design as the normal representation. Also in our case, this
-approach makes the program more predictable, less error-prone and, as the modern
-programming jargon lovingly expresses, {\em easier to reason about}.
+know from database design as the normal representation. It is the first time
+this approach reached the web front-end in such degree of popularity though.
 
-We are squeezing as much juice from these fruits as we can.
+Also in our case, this approach makes the program more predictable, less
+error-prone and, as the modern programming jargon lovingly expresses, {\em
+easier to reason about}. But some of our features make this a bit complicated.
 
@@
 transcribing a segment and replaying the selected audio span. Now, the only
-thing we actually store is the sub
+relevant things we actually store is the redux state are:
+\begin{enumerate}
+\item{
+    The array of transcription words, each of which bears the flag whether it is
+    automatically or manually transcribed. This defines the manual - automatic
+    chunks of words that in turn define the HTML elements wrapping them.
+}
+\item{
+    The beginning and end of the selection in terms of chunk number and
+    character offset in the chunk, which is basically what we get from the DOM
+    upon a \texttt{mouseup} event.
+}
+\item{Whether playback is running and the current position.
+\end{enumerate}
+
+Whether a word is marked or a segment is being edited is determined solely by
+the boundaries of the selected words. If there is a selection and the beginning
+and end are identical, it means a word was simply clicked and its detail is
+shown (it is {\em marked}). If the boundaries span at least one character, then
+all words that intersect this span are {\em selected} for correction.
+
+Simple as it sounds, a slight problem arises when a correction is accepted and
+the corrected subtitles are merged into the view. In the time after the
+correction is accepted and reflected in the redux state, but before the new
+chunks are rendered in the document, selection changes cannot be reliably mapped
+to logical chunks. Simple null defaults solve this problem.
+
+\section{Future Work}
+
+Next up in our to-do list is releasing the application to wider public and
+researching the
 

### 2018-03-25 2018-ftc-vancouver/webapp.tex
added FTC-vancouver paper

--- 2018-03-24 2018-tsd-brno/webapp.tex
@@
-\documentclass{llncs}
-%\usepackage[cp1250]{inputenc}
-%\usepackage[utf8]{inputenc}
-%\usepackage[czech]{babel}
+%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% sample root file for your "contribution" to a proceedings volume
+%
+% Use this file as a template for your own input.
+%
+%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass{svproc}
+%
+% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+
+% to typeset URLs, URIs, and DOIs
+\usepackage{url}
 \usepackage{graphicx}
 \usepackage{tabularx}
-%\usepackage{tipa}
 \usepackage{enumitem}
-
-\bibliographystyle{splncs}
-
-\begin{filecontents}{citace.bib}
-@inproceedings{kruuza2012making,
-  title={Making Community and ASR Join Forces in Web Environment},
-  author={Kr{\uu}za, Old{\v{r}}ich and Peterek, Nino},
-  booktitle={International Conference on Text, Speech and Dialogue},
-  pages={415--421},
-  year={2012},
-  organization={Springer}
-}
-\end{filecontents}
+\def\UrlFont{\rmfamily}
 
 \begin{document}
-\newtheorem{Definition}{Definition}
-\title{ASR-Output-Correcting Webapp After 6 Years}
-
+\mainmatter              % start of a contribution
+%
+\title{Second-Generation Web Interface to Correcting ASR output}
+%
+\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
+%                                     also used for the TOC unless
+%                                     \toctitle is used
+%
 \author{Old\v{r}ich Kr\r{u}za and Vladislav Kubo\v{n}}
+%
+\authorrunning{Kr\r{u}za, Kubo\v{n}} % abbreviated author list (for running head)
+%
+%%%% list of authors for the TOC (use if author list has to be modified)
+%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
+%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
+%
 \institute{Charles University in Prague\\
@@
 
-\maketitle
+\maketitle              % typeset the title of the contribution
 
 \begin{abstract}
-
 This paper presents a next-generation web application that enables users to
@@
 available in the browser.
-
+\keywords{Speech recognition, community-driven, web standards}
 \end{abstract}
@@
 
-In 2012\cite{kruuza2012making}, we presented a setting where a community of
+Kr\r{u}za 2012\cite{kruuza2012making} presented a setting where a community of
 users contributed corrections to automatically transcribed talks of a single
@@
 needless calculations. This is of course nothing new -- basically it is what we
-know from database design as the normal representation. Also in our case, this
-approach makes the program more predictable, less error-prone and, as the modern
-programming jargon lovingly expresses, {\em easier to reason about}.
+know from database design as the normal representation.
 
-We are squeezing as much juice from these fruits as we can.
+Also in our case, this approach makes the program more predictable, less
+error-prone and, as the modern programming jargon lovingly expresses, {\em
+easier to reason about}.
 
@@
 thing we actually store is the sub
+
+\section{Future Work}
+
+Next up in our to-do list is releasing the application to wider public and
+researching the
 
@@
 Education of the Czech Republic (project LM2010013).
-%\bibliography{paper}
-
-\bibliography{citace}
 

### 2018-03-25 2018-ftc-vancouver/webapp.tex
ftc2018 roughly done

@@
 %
-\institute{Charles University in Prague\\
+\institute{Charles University\\
            Faculty of Mathematics and Physics\\
            Institute of Formal and Applied Linguistics\\
-           Malostranské nám. 25, Prague, Czech Republic\\
+           Malostransk\'{e} n\'{a}m. 25, Prague, Czech Republic\\
            \{kruza,vk\}@ufal.mff.cuni.cz}
@@
   }
-  \em{common transcription applications} & \em{our application} \\
+  \em{common transcription applications}: & \em{our application}: \\
   are optimized for the case where there is no transcription available and it
-  must be acquired from scratch &
-    always assumes a transcription is available \\
-  allow annotation of speakers &
-    assumes all utterances come from the same speaker \\
+  must be acquired from scratch; &
+    always assumes a transcription is available; \\
+  allow annotation of speakers; &
+    assumes all utterances come from the same speaker; \\
   need no quality control: the user is free to enter whatever transcription she
-  pleases and the ultimate measure is her satisfaction &
+  pleases and the ultimate measure is her satisfaction; &
     needs the transcription to be accurate because it is used as training data
-    for the acoustic model \\
-  use alignment on the level of phrases, if any &
-    uses alignment on the level of words \\
-  are user-centric: the user transcribes whatever acoustic data they choose &
+    for the acoustic model; \\
+  use alignment on the level of phrases, if any; &
+    uses alignment on the level of words; \\
+  are user-centric: the user transcribes whatever acoustic data they choose; &
     is data-centric: the whole application with all its tools and persons
-    revolves around the data set \\
-  assumes the user wants to transcribe &
-    assume the user want to listen and possibly read along and we want to
-    animate her to submit transcriptions \\
-  has no shared data between users &
-    must count with collisions
+    revolves around the data set; \\
+  assumes the user wants to transcribe; &
+    assume the user wants to listen and possibly read along and we want to
+    animate her to submit transcriptions; \\
+  has no shared data between users; &
+    must count with collisions.
 \end{tabularx}
@@
 
-Major wikis have good measures for edit conflicts, which is where we could learn
-some lessons. However, so far there was no need to do that because
+Popular wikis have good measures for edit conflicts, which is where we could
+learn some lessons. However, so far there was no need to do that because
 \begin{enumerate}
@@
     \item{download-segment button, which initiates a file-save action for the
-    corresponding segment.}
+    audio segment corresponding the the selected text.}
     \end{itemize}
@@
 \item{
-    The page should be responsive;\label{feats:item:speed}
+    The page should be responsive.\label{feats:item:speed}
 }
@@
 The adoption of Web Audio API allowed for big improvements in comparison with
-the original implementation described by Kruza 2012~\cite{kruuza2012making}.
+the original implementation described by Kr\r{u}za 2012~\cite{kruuza2012making}.
 There are four major differences between using the HTML \texttt{<audio>} tag
@@
 needless calculations. This is of course nothing new -- basically it is what we
-know from database design as the normal representation.
+know from database design as the normal representation. It is the first time
+this approach reached the web front-end in such degree of popularity though.
+% TODO: cite normal representation
 
@@
 error-prone and, as the modern programming jargon lovingly expresses, {\em
-easier to reason about}.
+easier to reason about}. But some of our features make this a bit complicated.
 
@@
 transcribing a segment and replaying the selected audio span. Now, the only
-thing we actually store is the sub
+relevant things we actually store is the redux state are:
+\begin{enumerate}
+\item{
+    The array of transcription words, each of which bears the flag whether it is
+    automatically or manually transcribed. This defines the manual - automatic
+    chunks of words that in turn define the HTML elements wrapping them.
+}
+\item{
+    The beginning and end of the selection in terms of chunk number and
+    character offset in the chunk, which is basically what we get from the DOM
+    upon a \texttt{mouseup} event.
+}
+\item{Whether playback is running and the current position.}
+\end{enumerate}
+
+Whether a word is marked or a segment is being edited is determined solely by
+the boundaries of the selected words. If there is a selection and the beginning
+and end are identical, it means a word was simply clicked and its detail is
+shown (it is {\em marked}). If the boundaries span at least one character, then
+all words that intersect this span are {\em selected} for correction.
+
+Simple as it sounds, a slight problem arises when a correction is accepted and
+the corrected subtitles are merged into the view. In the time after the
+correction is accepted and reflected in the redux state, but before the new
+chunks are rendered in the document, selection changes cannot be reliably mapped
+to logical chunks. Simple null defaults solve this problem.
 
@@
 
-Next up in our to-do list is releasing the application to wider public and
-researching the
+We plan to focus on optimizing the app for wider audience. Experience confirms
+that the Mako\v{n}'s talks are of interest to some people, and our aim is to
+remove as many obstacles as possible to potentially interested people reaching
+the material. The benefit from technical point of view would be clear: A web app
+for listening to recordings and correcting their transcription is nice but one
+that's really easy to use and inviting to people to submit corrections is nicer.
 
-% alert when correcting humanic
+One of the aspects we want to explore is enabling people to naturally share
+catching segments of talks on social networks.
 
-%The need for quality control has wide implications. We need to perform forced
-%alignment every time a segment is transcribed. Hence, it must be clear what part
-%of the audio \em{exactly} the segment corresponds to. Since each word knows its
-%exact timestamp, and we know which words from the old transcription are to be
-%replaced, we can perform the forced alignment very precisely. This is also one
-%of the reasons why our application always assumes an existing transcription.
-
-%What concerns animating the user to provide a transcription, there must be more
-%research done. The basic idea is though that when the user sees the
-%transcription synchronously and it contains errors, they get the urge to fix it.
-
-%Our first reference is
-%Transcriber\footnote{sourceforge.net/projects/trans}. Transcriber is an
-%open-source program written in TCL.
+Another point of near future endeavor is higher-level work with the contents. By
+this we mean that we'd like to use both automatic processing methods and the
+users to do semantic analysis of the talks: What topic is covered where? What
+topics are covered at all? Which talks relate to which written works?, and
+similar questions.
 
@@
 
+Our work is motivated by the will to make the most out of a specific set of
+talks of spiritual nature by a single speaker. The setting and path taken are
+not common and we hope to make solution that can be used by others or to inspire
+others at solving their own tasks, however distantly related.
+
+The web interface to consuming and editing audio transcription we have created
+is first of its kind that we know of and we can but admire the recent
+developments in the web standards, their adoption by major browsers and the
+open-source JavaScript ecosystem that make this possible.
+
 \section*{Acknowledgments}
 
-The research was supported by SVV project number 265 314.\\
+The research was supported by SVV project number 260 453.\\
 \\
-This work has been using language resources stored
-by the LINDAT-Clarin project of the Ministry of
-Education of the Czech Republic (project LM2010013).
+use the prescribed wording: This work has been using language resources stored
+and distributed  by the LINDAT/CLARIN project of the Ministry of Education,
+Youth and Sports of the Czech Republic (project LM2015071).
 

### 2018-03-26 2018-ftc-vancouver/webapp.tex
cite codd normal form

@@
 \usepackage{enumitem}
+
+\begin{filecontents}{citace.bib}
+@inproceedings{kruuza2012making,
+  title={Making Community and ASR Join Forces in Web Environment},
+  author={Kr{\uu}za, Old{\v{r}}ich and Peterek, Nino},
+  booktitle={International Conference on Text, Speech and Dialogue},
+  pages={415--421},
+  year={2012},
+  organization={Springer}
+}
+@article{codd1970relational,
+  title={A relational model of data for large shared data banks},
+  author={Codd, Edgar F},
+  journal={Communications of the ACM},
+  volume={13},
+  number={6},
+  pages={377--387},
+  year={1970},
+  publisher={ACM}
+}
+\end{filecontents}
+
 \def\UrlFont{\rmfamily}
@@
 needless calculations. This is of course nothing new -- basically it is what we
-know from database design as the normal representation. It is the first time
-this approach reached the web front-end in such degree of popularity though.
-% TODO: cite normal representation
+know from database design as the normal representation.~\cite{codd1970relational}
+It is the first time this approach reached the web front-end in such degree of
+popularity though.
 

### 2018-03-26 2018-ftc-vancouver/webapp.tex
o systému nejen webappce

@@
 
-In our setting, we have a large spoken corpus (about 1000 hours) of a single
+The application we describe is a part of a larger system that deals with
+Mako\v{n}'s recordings. It consists roughly of (1) the corpus itself, (2) an ASR
+system trained specially for it and (3) a web interface for the users. These
+three parts form a whole where the ASR gives a baseline transcription, the users
+correct it and the corrections are fed as further training data for the acoustic
+and language models. In this paper, we focus on the web interface.
+
+The spoken corpus is about 1000 hours of a single
 speaker. Our aim is to have a transcription as good as possible for the purpose

### 2018-03-26 2018-ftc-vancouver/webapp.tex
citations

@@
 }
+@article{adenot2013web,
+  title={Web audio api},
+  author={Adenot, Paul and Wilson, Chris and Rogers, Chris},
+  journal={W3C, October},
+  volume={10},
+  year={2013}
+}
+@article{abramov2015redux,
+  title={Redux},
+  author={Abramov, D},
+  journal={React Community, c},
+  year={2015}
+}
+
 \end{filecontents}
@@
 
-The adoption of Web Audio API allowed for big improvements in comparison with
-the original implementation described by Kr\r{u}za 2012~\cite{kruuza2012making}.
-There are four major differences between using the HTML \texttt{<audio>} tag
-and the Web Audio API.
+The adoption of Web Audio API~\cite{adenot2013web} allowed for big improvements
+in comparison with the original implementation described by Kr\r{u}za
+2012~\cite{kruuza2012making}. There are four major differences between using the
+HTML \texttt{<audio>} tag and the Web Audio API.
 
@@
 
-We use React as the view library and Redux for state management. The good thing
-about Redux is that it makes it easy to keep minimal state as the single source
-of truth and everything that can be computed is computed, while avoiding
-needless calculations. This is of course nothing new -- basically it is what we
-know from database design as the normal representation.~\cite{codd1970relational}
-It is the first time this approach reached the web front-end in such degree of
-popularity though.
+We use React as the view library and Redux\cite{abramov2015redux} for state
+management. The good thing about Redux is that it makes it easy to keep minimal
+state as the single source of truth and everything that can be computed is
+computed, while avoiding needless calculations. This is of course nothing new --
+basically it is what we know from database design as the normal
+representation.~\cite{codd1970relational} It is the first time this approach
+reached the web front-end in such degree of popularity though.
 
@@
 
+\bibliography{citace}
+\bibliographystyle{splncs03}
+
 \end{document}

### 2018-03-28 2018-ftc-vancouver/webapp.tex
dingir citation

@@
   title={Making Community and ASR Join Forces in Web Environment},
-  author={Kr{\uu}za, Old{\v{r}}ich and Peterek, Nino},
+  author={Kr\r{u}za, Old{\v{r}}ich and Peterek, Nino},
   booktitle={International Conference on Text, Speech and Dialogue},
@@
 }
-
+@article{hajek2007cesky,
+  title={\v{C}esk\'{y} mystik Karel Mako\v{n}},
+  author={Jurik H\'{a}jek},
+  journal={Dingir},
+  year={2007},
+  volume={2007/4},
+  pages={142--143},
+  publisher={Dingir, s.r.o.},
+  ISSN={1212-1371},
+  editor={Zden\v{e}k Vojt\'{i}\v{s}ek}
+}
 \end{filecontents}
@@
 recordings. We describe differences from similar settings, compare our solution
-with others and reflect on the development from the first, 6 years old version
-in the light of the work done, lessons learned and the new technologies
-available in the browser.
+with others and reflect on the development from the now 6 years old work we
+build upon in the light of the progress made, lessons learned and the new
+technologies available in the browser.
 \keywords{Speech recognition, community-driven, web standards}
@@
 
-Our project focuses on the collection of recordings of Karel Mako\v{n} *1912
-\textdagger1993, the author of numerous books, translations and comments to
-works of spiritual and religious nature, who was influenced by trances during
-recurring surgery without anesthesy in the age of 6, ecstasies in the youth and
-finally facing and surviving certain death in a Nazi concentration camp, after
-which he experienced enlightenment. He gave talks in a narrow circle of friends
-and the recordings in our care have been taken between early 70's and 1991,
-spanning about 1000 hours in total.
+Our project focuses on the collection of recordings of Karel
+Mako\v{n}~\cite{hajek2007cesky} *1912 \textdagger1993, the author of numerous
+books, translations and comments to works of spiritual and religious nature, who
+was influenced by trances during recurring surgery without anesthesy in the age
+of 6, ecstasies in the youth and finally facing and surviving certain death in a
+Nazi concentration camp, after which he experienced enlightenment. He gave talks
+in a narrow circle of friends and the recordings in our care have been taken
+between early 70's and 1991, spanning about 1000 hours in total.
 
@@
 
+\bibliographystyle{splncs03}
 \bibliography{citace}
-\bibliographystyle{splncs03}
 

### 2018-03-31 2018-ftc-vancouver/webapp.tex
several enhancements, notably
- describe correction submitting
- more factual conclusion
- Title Case

@@
 %
-\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
+\titlerunning{Second-Generation Web Interface to Correcting ASR output}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 technologies available in the browser.
-\keywords{Speech recognition, community-driven, web standards}
+\keywords{Speech recognition, transcription, community-driven, web standards}
 \end{abstract}
@@
 Our project focuses on the collection of recordings of Karel
-Mako\v{n}~\cite{hajek2007cesky} *1912 \textdagger1993, the author of numerous
+Mako\v{n}\cite{hajek2007cesky} *1912 \textdagger1993, the author of numerous
 books, translations and comments to works of spiritual and religious nature, who
@@
 
-All of his work deals more or less directly with a single topic: entering the
-eternal life before the physical death. He draws mainly from the Christian
-symbolism, builds up on Christian mysticism and ancient tradition of India.
+All of Mako\v{n}'s work deals more or less directly with a single topic:
+entering the eternal life before the physical death. He draws mainly from the
+Christian symbolism, builds up on Christian mysticism and ancient tradition of
+India and China.
 
@@
 works, whereas the talks contain the sauce: talks tailored to the audience,
-answering questions, personal experiences, behind-the-scenes to the books etc.
+answers to questions, personal experiences, behind-the-scenes to the books etc.
 
-\section{Differences to other settings}
+\section{Differences to Other Settings}
 
@@
 three parts form a whole where the ASR gives a baseline transcription, the users
-correct it and the corrections are fed as further training data for the acoustic
+correct it and the corrections are fed as further training data to the acoustic
 and language models. In this paper, we focus on the web interface.
@@
 
-\subsection{Transcription apps}
+\subsection{Transcription Apps}
 
@@
   must be acquired from scratch; &
-    always assumes a transcription is available; \\
+    always assumes a prior transcription is available; \\
   allow annotation of speakers; &
@@
   has no shared data between users; &
-    must count with collisions.
+    must count with collisions. \\
 \end{tabularx}
 
-We can still learn a lot from transcription software. The ease of performing
-common tasks, like pausing, resuming and rewinding is crucial for the user
-experience and in effect for the amount of submissions that we receive. Also,
-the way the text is displayed synchronously to the audio played has a big impact
-and the approaches have a lot of space for variation.
+\vspace{5mm}
+
+Despite of these differences, we can still learn a lot from transcription
+software. The ease of performing common tasks, like pausing, resuming and
+rewinding is crucial for the user experience and in effect for the amount of
+submissions that we receive. Also, the way the text is displayed synchronously
+to the audio played has a big impact and the approaches have a lot of space for
+variation.
 
@@
 where the quality of the contributions is essential, while the contributor's
-satisfaction alone is of little importance.
+satisfaction alone is of less importance.
 
@@
 
-\section{Description of the web application}
+\section{Description of the Web Application}
+
+We have pointed out the differences in our setting to others. Now we shall
+describe how we implemented the program.
 
@@
     \item{download-segment button, which initiates a file-save action for the
-    audio segment corresponding the the selected text.}
+    audio segment corresponding the the selected text. The synthesis of the
+    downloaded file takes place in the browser.}
     \end{itemize}
@@
 
-\subsection{Displaying the transcription}
+Most common tasks have keyboard shortcuts: Ctrl-Space for play/pause and
+Ctrl-Enter for submitting a correction.
+
+\subsection{Displaying the Transcription}
 
@@
 
-\subsubsection{Performance challenge}
+\subsubsection{Performance Challenge}
 
@@
 Kr\r{u}za 2012~\cite{kruuza2012making} solve this by sacrificing point~5:
-% TODO obrázek MakonFM
-only 3
-lines of text are shown with the current word kept on the middle line. Thanks to
+only 3 lines of text are shown with the current word kept on the middle line as
+shown on Figure~\ref{fig:makonfm}.\footnote{The current word is on the top line
+on the screenshot because it is at the beginning of the recording.} Thanks to
 the development in the web standards and their support from popular browsers, a
 solution is possible.
+
+\begin{figure}[htpb]
+\includegraphics[scale=0.7]{rc/makonfm-en-1.png}
+\caption{Original web interface from 2012}
+\label{fig:makonfm}
+\end{figure}
 
@@
 submitted segment is 7.9. Furthermore, the absolute majority of such segments
-are adjacent to other manually transcribed chunks. 
-% TODO max fragmentation
+are adjacent to other manually transcribed chunks.\footnote{
+The median number of chunks is 1 (most recordings have no manually corrected
+segments), maximum is 1109. Median only counting touched recordings is 8.}
 Hence, wrapping each chunk of consecutive manually or automatically transcribed
@@
 out the innermost HTML element and text offset of the start and end of the
-textual selection. Since we know the length of the words, this allows us to map
+textual selection. Since we know the length of each word, this allows us to map
 the selection to the corresponding words in the transcription.
@@
 
+\subsection{Mechanics of Submitting a Corrected Segment}
+
+As stated above, when the user selects at least one character with the mouse,
+the application enters the state of correcting the selected transcription padded
+to whole words. In this mode, the transcription to correct is shown in a
+textarea and the global playback controls are replaced by those that only allow
+playback of audio corresponding to the selected transcription.
+
+Once the user believes that the content of the textarea corresponds precisely
+to the words uttered, she hits the {\em save} button or the Ctrl-Enter keyboard
+shortcut. This starts an asynchronous HTTP request to the back-end, where
+parametrized (MFCC) versions of the recordings are stored, along with the new
+transcription and the time positions of the beginning and end of the segment.
+
+The server then cuts off the corresponding segment from the parametrized
+recording, runs forced alignment on it with the provided transcription with a
+threshold to reject bad matches. If the forced alignment fails, an error
+response is sent back and the transcription is not merged into the original. In
+the case of a success, the correction is merged on one hand on the server side
+and pushed to a CDN, on the other hand it is merged into the transcription word
+array in the JavaScript application. This redundancy warrants that we don't have
+to reload the whole transcription every time a segment is corrected.
+
+React ensures the updating of the chunks, and the coordinates of the words
+further in the document are recalculated for word-highlighting purposes.
+
+Apart from this, the version of the transcription to the recording is updated.
+This is because the transcriptions have a long cache time because normally, they
+don't change at all. At the page load, the versions of all transcriptions are
+loaded and used as cache busters. This enables us to use an external CDN and
+cache effectively.
+
 \subsection{Implementation Details}
@@
 
-The adoption of Web Audio API~\cite{adenot2013web} allowed for big improvements
-in comparison with the original implementation described by Kr\r{u}za
-2012~\cite{kruuza2012making}. There are four major differences between using the
-HTML \texttt{<audio>} tag and the Web Audio API.
+The adoption of Web Audio API\cite{adenot2013web} allowed for big improvements
+in comparison with the original implementation. There are four major differences
+between using the HTML \texttt{<audio>} tag and the Web Audio API.
 
@@
     recording in the browser's storage and avoid downloading or decoding it
-    again after reload.
+    again after reload. We use \texttt{IndexedDB} as the storage method because
+    \texttt{localStorage} has too low quota of about 10MB and the
+    \texttt{FileSystem API} is not yet widely enough supported.
 }
@@
 basically it is what we know from database design as the normal
-representation.~\cite{codd1970relational} It is the first time this approach
+representation.\cite{codd1970relational} It is the first time this approach
 reached the web front-end in such degree of popularity though.
@@
 We plan to focus on optimizing the app for wider audience. Experience confirms
-that the Mako\v{n}'s talks are of interest to some people, and our aim is to
-remove as many obstacles as possible to potentially interested people reaching
+that Mako\v{n}'s talks are of interest to some people, and our aim is to
+remove as many obstacles as possible from potentially interested people reaching
 the material. The benefit from technical point of view would be clear: A web app
@@
 
-Our work is motivated by the will to make the most out of a specific set of
-talks of spiritual nature by a single speaker. The setting and path taken are
-not common and we hope to make solution that can be used by others or to inspire
-others at solving their own tasks, however distantly related.
-
-The web interface to consuming and editing audio transcription we have created
-is first of its kind that we know of and we can but admire the recent
-developments in the web standards, their adoption by major browsers and the
-open-source JavaScript ecosystem that make this possible.
+With our web application, a user can listen to recorded speech, see its
+transcription with the currently played word highlighted, commit corrections to
+the transcription, and inspect a word as well as a span of text and audio. The
+corrections are checked by a forced-alignment mechanism on the server side. Our
+solution overcomes performance challenges and is a serious improvement from the
+original version. More widespread usage as well as employing it on another data
+set is to be done.
 
@@
 \\
-use the prescribed wording: This work has been using language resources stored
+This work has been using language resources stored
 and distributed  by the LINDAT/CLARIN project of the Ministry of Education,

### 2018-03-31 2018-ftc-vancouver/webapp.tex
stop pretending we aren't we

@@
 
-Kr\r{u}za 2012\cite{kruuza2012making} presented a setting where a community of
+In 2012\cite{kruuza2012making}, we have presented a setting where a community of
 users contributed corrections to automatically transcribed talks of a single
@@
 
-Kr\r{u}za 2012~\cite{kruuza2012making} solve this by sacrificing point~5:
+In the original version, we solved this by sacrificing point~5:
 only 3 lines of text are shown with the current word kept on the middle line as

### 2018-03-31 2018-ftc-vancouver/webapp.tex
compare specific transcription apps

@@
 Basically the meat of his teachings is systematically encompassed in his written
-works, whereas the talks contain the sauce: talks tailored to the audience,
+works, whilst the talks contain the sauce: talks tailored to the audience,
 answers to questions, personal experiences, behind-the-scenes to the books etc.
@@
 \section{Differences to Other Settings}
+
+% wreally: pause interval
+%          auto loop
+%          templates
+%          pedál
+%          Fka
+%          bez alignmentu
 
@@
 
-The main differences to common transcription software are that
+The best widespread match to our task is that of creating an application for
+transcribing speech recordings. Let's compare the two tasks, pointing out the
+main points of difference. For reference, we take (1)
+Transcriber\footnote{trans.sourceforge.net}, a classical open-source program
+written in TCL, (2) oTranscribe\footnote{otranscribe.com}, a free modern
+web-based transcription tool and (3) Transcribe\footnote{transcribe.wreally.com}
+a commercial web-based transcription tool.
+
+The numbers in the bullet list denote the programs our statement applies to. For
+example, of the three only Transcriber allows speaker annotation, hence there is
+only the number (1) standing at the second list item.
 
@@
   }
-  \em{common transcription applications}: & \em{our application}: \\
-  are optimized for the case where there is no transcription available and it
+  \em{transcription applications}: & \em{our application}: \\
+  (1,2,3) are optimized for the case where there is no transcription available and it
   must be acquired from scratch; &
     always assumes a prior transcription is available; \\
-  allow annotation of speakers; &
+  (1) allow annotation of speakers; &
     assumes all utterances come from the same speaker; \\
-  need no quality control: the user is free to enter whatever transcription she
-  pleases and the ultimate measure is her satisfaction; &
+  (1,2,3) need no quality control: the user is free to enter whatever
+  transcription she pleases and the ultimate measure is her satisfaction; &
     needs the transcription to be accurate because it is used as training data
     for the acoustic model; \\
-  use alignment on the level of phrases, if any; &
+  (1)\footnote{Transcriber explicitly aligns the text with speech, while the
+  other two merely automate addition of timestamps into the transcription.} use
+  alignment on the level of phrases, if any; &
     uses alignment on the level of words; \\
-  are user-centric: the user transcribes whatever acoustic data they choose; &
+  (1,2,3) are user-centric: the user transcribes whatever acoustic data they
+  choose; &
     is data-centric: the whole application with all its tools and persons
     revolves around the data set; \\
-  assumes the user wants to transcribe; &
+  (1,2,3) assumes the user wants to transcribe; &
     assume the user wants to listen and possibly read along and we want to
     animate her to submit transcriptions; \\
-  has no shared data between users; &
+  (1,2)\footnote{Transcribe supports team co-operation.} has no shared data
+  between users; &
     must count with collisions. \\
@@
 utterances, see Figure~\ref{fig:transcriber1} for an example of
-{\em{Transcriber}}\footnote{trans.sourceforge.net}, a veteran open-source
-transcription tool. We attribute this to the fact that the atomic elements of
+Transcriber. We attribute this to the fact that the atomic elements of
 the transcription are the user-entered utterances and their boundaries are

### 2018-03-31 2018-ftc-vancouver/webapp.tex
tidy the tabularx source code

@@
     @{\hspace{1.5em}}% Space for left bullet
-    >{\leavevmode\llap{\textbullet~}\raggedright}% Left bullet + formatting of column
+    >{{\hsize=0.9\hsize}\leavevmode\llap{\textbullet~}\raggedright}% Left bullet + formatting of column
     X% Left column specification
+    @{\hspace{0.2em}}
+    >{\hsize=0.2\hsize}
+    X
     @{\quad\hspace{1.5em}}% Space between columns + right bullet space
@@
   }
-  \em{transcription applications}: & \em{our application}: \\
-  (1,2,3) are optimized for the case where there is no transcription available and it
+  \em{transcription applications}: & & \em{our application}: \\
+  are optimized for the case where there is no transcription available and it
   must be acquired from scratch; &
+    (1,2,3) &
       always assumes a prior transcription is available; \\
-  (1) allow annotation of speakers; &
+
+  allow annotation of speakers; &
+    (1) &
       assumes all utterances come from the same speaker; \\
-  (1,2,3) need no quality control: the user is free to enter whatever
-  transcription she pleases and the ultimate measure is her satisfaction; &
+
+  need no quality control: the user is free to enter whatever transcription she
+  pleases and the ultimate measure is her satisfaction; &
+    (1,2,3) &
       needs the transcription to be accurate because it is used as training data
       for the acoustic model; \\
-  (1)\footnote{Transcriber explicitly aligns the text with speech, while the
-  other two merely automate addition of timestamps into the transcription.} use
-  alignment on the level of phrases, if any; &
+
+  use alignment on the level of phrases, if any; &
+    (1)\footnote{
+        Transcriber explicitly aligns the text with speech, while the
+        other two merely automate addition of timestamps into the transcription.
+    } &
       uses alignment on the level of words; \\
-  (1,2,3) are user-centric: the user transcribes whatever acoustic data they
-  choose; &
+
+  are user-centric: the user transcribes whatever acoustic data they choose; &
+    (1,2,3) &
       is data-centric: the whole application with all its tools and persons
       revolves around the data set; \\
-  (1,2,3) assumes the user wants to transcribe; &
+
+  assumes the user wants to transcribe; &
+    (1,2,3) &
       assume the user wants to listen and possibly read along and we want to
       animate her to submit transcriptions; \\
-  (1,2)\footnote{Transcribe supports team co-operation.} has no shared data
-  between users; &
+
+  has no shared data between users; &
+    (1,2)\footnote{Transcribe supports team co-operation.} &
       must count with collisions. \\

### 2018-03-31 2018-ftc-vancouver/webapp.tex
row margin for tabularx
rant about ergonomy

@@
 \usepackage{tabularx}
+\setlength{\extrarowheight}{6pt}
 \usepackage{enumitem}
@@
 \subsection{Transcription Apps}
+\label{ssec:diff:trans}
 
@@
 
+\subsection{Ergonomy}
+
+It is clear that the ease of use is crucial in our case where the user is
+supposed to perform a requiring, tedious task with repeated steps, especially
+since it is our interest more than hers that she performs them. We compared our
+setting with that of transcription apps in section~\ref{ssec:diff:trans},
+pointing out lessons to learn. Let's now look at some specific points and their
+actual (lack of) implementation.
+
+\subsubsection{Keyboard Shortcuts}
+
+One of the most profound measures in ergonomy are keyboard shortcuts. The most
+common task is pausing and resuming playback. Both oTranscribe and Transcribe
+use the \texttt{esc} key for that, and Transcriber uses the \texttt{tab} key.
+We chose \texttt{ctrl+space} combination. We argue that \texttt{esc} is not the
+best of options because the distance the fingers have to travel from the
+alphanumeric keys causes a noticeable delay. This can lead to missing a pause
+between words. The \texttt{tab} key as chosen by Transcriber is a stellar choice
+from the ergonomy point of view and there is no reason not to use it in a
+dedicated user interface. However, in the browser, where the \texttt{tab} key
+has as native use, re-binding it could lead to confusion and irritation. The
+space bar is probably the easiest-to-find key in all situations and dedicating
+\texttt{ctrl} to all application-specific commands as opposed to single keys
+lends a sense of consistency, we believe.
+
+This is speculation though, as we had no resources so far to perform serious
+research to support these statements.
+
+The only other keyboard shortcut we support is \texttt{ctrl+enter} for
+submitting the correction. We chose this to stay consistent using the
+\texttt{ctrl} key and because this shortcut is familiar to users of many instant
+messengers, like the Facebook chat or the once popular official ICQ client.
+Also, requiring a key combination prevents accidental submission, which is
+desirable as we only want double-checked, guaranteed-precise ones. In
+comparison, Transcriber uses the base \texttt{enter} key to separate utterances.
+oTranscribe and Transcribe allow free formatting with no explicit alignment, so
+using the enter key to split utterances by lines is the user's choice.
+
+\subsubsection{Missing Features}
+
+One of the features that Transcribe, the only commercial tool in our reference
+list, offers is setting up keyboard shortcuts for common words. We haven't
+implemented this because ideally, common words should be covered by the speech
+recognition. However, it could be sensible to implement it anyway. The reason is
+that a word can be very rare globally and thus poorly recognized by ASR but very
+common in a specific passage. This particularly regards named entities.
+
+Another point in our ergonomy to-do list is lifting the need to select a segment
+prior to correcting it. If the transcription was simply editable, it could
+increase the ease of use rapidly. We would have to automate the selection of
+segment to send for forced alignment but we could probably do a better job than
+the user in the end.
+
 \subsection{Mechanics of Submitting a Corrected Segment}

### 2018-03-31 2018-ftc-vancouver/webapp.tex
minor edits plus grey/black font choice

@@
 
-The numbers in the bullet list denote the programs our statement applies to. For
-example, of the three only Transcriber allows speaker annotation, hence there is
-only the number (1) standing at the second list item.
+The numbers in the bullet list below denote the programs our statement applies
+to. For example, of the three only Transcriber allows speaker annotation, hence
+there is only the number (1) standing at the second list item.
 
@@
         Transcriber explicitly aligns the text with speech, while the
-        other two merely automate addition of timestamps into the transcription.
+        other two merely support addition of timestamps into the transcription.
     } &
@@
 
-Where our application diverts from transciption software, it mostly resembles a
+Where our application diverts from transcription software, it mostly resembles a
 wiki: a community platform that serves its users including the contributors but
@@
 
-One major difference to a wiki is that wiki is creative, whereas our task is
-mechanical. The user has basically no room for their own invention: providing a
-different than correct transcription is seen as an error.
+One major point of difference to a wiki is that wiki is creative, whereas our
+task is mechanical. The user has basically no room for their own invention:
+providing a different than correct transcription is seen as an error.
 
@@
 \begin{enumerate}
-\item{when we always simply take the most recent version
+\item{if we always simply take the most recent version
 of a segment, the result stays consistent even if a piece from user A comes into
@@
 is displayed and can be corrected by the user,}
-\item{search page, where hits to a search query are listed and point to
+\item{the search page, where hits to a search query are listed and point to
 corresponding positions in the recordings,}
@@
 \item{Identifier of the recording;}
-\item{Automatically transcribed segments in gray;}
+\item{Automatically transcribed segments in grey;}
 \item{Manually transcribed segments in black;}
@@
         occurrence: the word with contextual capitalization and
-        punctuation as it appeard in the text, currently being edited as the
-        selected initial letter reveals,
+        punctuation as it appeard in the text (currently being edited as the
+        selected initial letter reveals),
     }
@@
 
-Most common tasks have keyboard shortcuts: Ctrl-Space for play/pause and
-Ctrl-Enter for submitting a correction.
+Most common tasks have keyboard shortcuts: \texttt{ctrl+space} for play/pause
+and \texttt{ctrl+enter} for submitting a correction.
 
@@
 
-Most transcription programs show the transcription as a vertical list of
+Many transcription programs show the transcription as a vertical list of
 utterances, see Figure~\ref{fig:transcriber1} for an example of
@@
 
+\subsubsection{Manual / Automatic Distinction}
+
+As shown on Figure~\ref{fig:scn1lab}, we draw automatic transcription in grey
+and manual one in black. Why did we choose this instead of normal / boldface?
+Firstly, the normal font is optimal for reading. Boldface is meant to highlight
+spots in text. It becomes bulky when applied on long continuous passages. The
+automatic transcription contains many errors, so there is no sense in optimizing
+it for best reading experience.
+
+There is also another practical reason. When the two font variants only differ
+in color, and a segment of automatic transcription is left intact and submitted
+as correct transcription, its merge-down into the displayed text causes no
+reflow, which saves us computations and raises responsiveness. It may seem like
+a rare use case but we believe that identifying correctly recognized words is a
+legitimate way of contribution, so why not optimize for it?
+
+Still, the underlying HTML tags are \texttt{<span>} and \texttt{<b>} because
+that way the distinction persists when copy-pasting the text from the web page
+to a rich text editor.
+
 \subsection{Ergonomy}
@@
 alphanumeric keys causes a noticeable delay. This can lead to missing a pause
-between words. The \texttt{tab} key as chosen by Transcriber is a stellar choice
+between words. The \texttt{tab} key as chosen by Transcriber is a splendid choice
 from the ergonomy point of view and there is no reason not to use it in a
@@
 
-This is speculation though, as we had no resources so far to perform serious
-research to support these statements.
+This is mere personal experience though, as we had no resources so far to
+perform serious research to support these statements.
 
@@
 desirable as we only want double-checked, guaranteed-precise ones. In
-comparison, Transcriber uses the base \texttt{enter} key to separate utterances.
+comparison, Transcriber uses the bare \texttt{enter} key to separate utterances.
 oTranscribe and Transcribe allow free formatting with no explicit alignment, so
@@
 Once the user believes that the content of the textarea corresponds precisely
-to the words uttered, she hits the {\em save} button or the Ctrl-Enter keyboard
-shortcut. This starts an asynchronous HTTP request to the back-end, where
-parametrized (MFCC) versions of the recordings are stored, along with the new
-transcription and the time positions of the beginning and end of the segment.
+to the words uttered, she hits the {\em save} button or the \texttt{ctrl+enter}
+keyboard shortcut. This starts an asynchronous HTTP request to the back-end,
+where parametrized (MFCC) versions of the recordings are stored, along with the
+new transcription and the time positions of the beginning and end of the
+segment.
 
@@
 
-Among the states the app can enter is simple playback, inspecting a word,
-transcribing a segment and replaying the selected audio span. Now, the only
-relevant things we actually store is the redux state are:
+Among the states the app can enter is simple playback, inspecting a word and
+transcribing a segment. Now, the only relevant things we actually keep in the
+Redux store are:
+
 \begin{enumerate}
@@
 }
-\item{Whether playback is running and the current position.}
 \end{enumerate}
@@
 
-Another point of near future endeavor is higher-level work with the contents. By
+Another point of near-future endeavor is higher-level work with the contents. By
 this we mean that we'd like to use both automatic processing methods and the

### 2018-03-31 2018-ftc-vancouver/webapp-anon.tex
anonymized version

--- 2018-03-31 2018-ftc-vancouver/webapp.tex
@@
 
-\begin{filecontents}{citace.bib}
+\begin{filecontents}{citace-anon.bib}
 @inproceedings{kruuza2012making,
-  title={Making Community and ASR Join Forces in Web Environment},
-  author={Kr\r{u}za, Old{\v{r}}ich and Peterek, Nino},
-  booktitle={International Conference on Text, Speech and Dialogue},
-  pages={415--421},
+  title={The unnamed previous version},
+  author={John Doe and Joan Doe},
+  booktitle={The Example Conference},
+  pages={42--4242},
   year={2012},
-  organization={Springer}
+  organization={Example Publisher}
 }
@@
 %
-\author{Old\v{r}ich Kr\r{u}za and Vladislav Kubo\v{n}}
+\author{John Doe and Joan Doe}
 %
-\authorrunning{Kr\r{u}za, Kubo\v{n}} % abbreviated author list (for running head)
+\authorrunning{John Doe and Joan Doe} % abbreviated author list (for running head)
 %
@@
 %
-\institute{Charles University\\
-           Faculty of Mathematics and Physics\\
-           Institute of Formal and Applied Linguistics\\
-           Malostransk\'{e} n\'{a}m. 25, Prague, Czech Republic\\
-           \{kruza,vk\}@ufal.mff.cuni.cz}
+\institute{Unnamed University\\
+           Faculty of Placeholder Arts\\
+           Example Institute\\
+           Street 42, City, Country\\
+           \{john.doe,joan.doe\}@example.com}
 
@@
 
-The research was supported by SVV project number 260 453.\\
+The research was supported by Example project number 123456.\\
 \\
-This work has been using language resources stored
-and distributed  by the LINDAT/CLARIN project of the Ministry of Education,
-Youth and Sports of the Czech Republic (project LM2015071).
+The research was supported by Example project number 123456.\\
+\\
+The research was supported by Example project number 123456.
 
 \bibliographystyle{splncs03}
-\bibliography{citace}
+\bibliography{citace-anon}
 

### 2018-04-05 2018-ftc-vancouver/webapp-anon.tex
correct case for title

@@
 %
-\title{Second-Generation Web Interface to Correcting ASR output}
+\title{Second-Generation Web Interface to Correcting ASR Output}
 %
-\titlerunning{Second-Generation Web Interface to Correcting ASR output}  % abbreviated title (for running head)
+\titlerunning{Second-Generation Web Interface to Correcting ASR Output}  % abbreviated title (for running head)
 %                                     also used for the TOC unless

### 2018-04-08 2018-clarin-pisa/makon-data.tex
start writing the 2018 clarin extended abstract

%!TEX encoding = UTF-8 Unicode
%!TEX TS-program = xelatex
%
% File CLARIN2018.tex
%
% Adapted for modern XeLaTeX and Unicode by desmedt@uib.no
% Otherwise contact: jwagner@computing.dcu.ie
%% 
%% Based on the style files for CLARIN2015, which were, in turn,
%% Based on the style files for CAC2014, which were, in turn,
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[a4paper,11pt]{article}
\usepackage{CLARIN2018}
% - - - - - - - IMPORTANT - - - - - - - -
% The next three lines allow XeLaTeX, graphics import and hyperlinks, set font and language
\usepackage{xltxtra,polyglossia,graphicx,hyperref}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setdefaultlanguage{english}
% If for some reason the above three lines are not compatible with your LaTeX installation,
% comment out the above three instructions and uncomment the following four instead:
%\usepackage{times}
%\usepackage{url}
%\usepackage{latexsym}
%\usepackage[english]{babel}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\usepackage{covington} % if needed, for linguistic examples

\title{Spoken Corpus of Karel Mako\v{n}}

% - - - - - - - IMPORTANT - - - - - - -
% Leave the author information empty until your paper has been accepted

% Uncomment the following line ONLY if you need two author rows
%\setlength\titlebox{80mm}

\author{First Author \\
  Department (optional)\\
  University of City, Country \\
  {\tt email@domain} \\\And % if needed: this makes a second column
  Second Author \\
  Department (optional)\\
  University Name without city \\
  City, Country \\
 {\tt email@domain} \\
% \AND % if needed: this makes a second row
%  Third Author \\
%  Department (optional)\\
%  University of City, Country \\
%  {\tt email@domain} \\\And
%  Fourth Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
%  {\tt email@domain} \\
}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  We present a detailed view on the spoken corpus of Karel Mako\v{n}. This paper
  briefly outlines the system for processing the corpus, which has been
  published in a previous work, and elaborates on the corpus, the process of its
  acquisition and its annotation.
\end{abstract}

\section{Introduction} \label{intro}

% - - - - - - - IMPORTANT - - - - - - -
% The following footnote without marker is needed for the camera-ready
% version of the paper.
% Comment out the instructions (first text) and uncomment the 8 lines
% under "final paper" for your variant of English.
%
\blfootnote{
    %
    % for review submission
    %
    %\hspace{-0.65cm}  % space normally used by the marker
This work is licenced under a Creative Commons Attribution 4.0 International Licence.
Licence details:
\url{http://creativecommons.org/licenses/by/4.0/}
    % % final paper: en-uk version (to license, a licence)
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licensed under a Creative Commons
    % Attribution 4.0 International Licence.
    % Licence details:
    % \url{http://creativecommons.org/licenses/by/4.0/}
    %
    % % final paper: en-us version (to licence, a license)
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licenced under a Creative Commons
    % Attribution 4.0 International License.
    % License details:
    % \url{http://creativecommons.org/licenses/by/4.0/}
}

The spoken corpus of Karel Mako\v{n} is a collection of talks given in a circle
of friends in the course of late 60's or early 70's till 1991. The recordings
have been kept on magnetophone tapes until their digitization that took place
between 2010 and 2012. A complete transcription was obtained using a dedicated
ASR system and the work of the community around K.M's legacy. The corpus is
about 1000 hours in total length, of which about 66 have been transcribed
manually.

\section{Content of the Corpus}

\subsection{Karel Mako\v{n}}

The author of the talks, Mr. Karel Mako\v{n} *1912 \textdagger{}1993, was giving
talks in a strife to share his awareness of the proverbial meaning of life and
to give a manual to eternal life after his release from a concentration camp
during the WWII.

His steep spiritual path began in early childhood when he was
subject to repeated arm surgery without anesthesy. The pain the child suffered
made him escape mentally from inside his body, which led him to what he calls
``the yellow light''. As a result, he felt obliged always to {\em do the right
thing}.

This culminated in the age of 17, in a perceived life crissis, when he read in a
book that ``this life is the bridge to the eternal life''. The thought
captivated him and gave him a new life direction, which also led to several
ecstasies.

The next culmination happened in 1939, after nine years of constant prayer for a
stronger love to God, when he was doported to the nazi concentration camp in
Sachsenhausen, as a university student. Here, he was promised certain death by a
Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
life completely to God and the attacker turned around in awe and fled.

It was in that moment that Mako\v{n} obtained enlightenment, accompanied with
profound understanding of Christian symbolism. He was experiencing absolute
freedom and abundance ever since, even in the concentration camp. He also
understood the general spiritual mechanisms that conditioned his deliverance and
devoted much of his subsequent life to enabling others to learn from his
experience.

\subsection{Giving and Recording the Talks}

The mystical, religious and spiritual nature of Mako\v{n}'s talks put K.M. in
direct antagonism with the communist regime. The opposition was strictly
unidirectional though, as Mako\v{n} also understood the purpose of the regime
and was far from wasting his energy on a fight against it. His way of dealing
with the conflict was by staying out of wide popularity. Hence, most of the
talks have been given during regular meetings in various places of
Czechoslovakia and later the Czech Republic, in a close circle of friends.

One of the fortunate consequences of the seclusion of the talks is that there is
very little background noise. Also, most of the recordings
% TODO: kolik je od Elgra?
have been taken by a single person who used the available state-of-the-art
amateur recording technology and methods. He was also very systematic in
labeling and archiving the recordings. As a result, the majority of the base for
digitization was a set of reel-to-reel tapes labeled with ordinal indexes and a
set of cassettes labeled by year and ordinal index. The reel-to-reel tapes have
been acquired in a generous speed of 9cm/second, which granted them a very good
shield against the inevitable aging.

\subsection{Topics}

In general, the whole corpus deals with a single topic that can be summarized as
a howto for entering the eternal life before the physical death. On a
finer-grained look, we can identify recurring sub-topics, like

\begin{itemize}
\item{interpretation of some passages in the New Testament, notably the parable
of the prodigal son, the parable of the talents, and the Our Father prayer,}
\item{milestone personal experiences, like the one in the concentration camp,}
\item{explaining symbolism, the apostles as symbols for human abilities,}
\item{references to specific people, like St. Teresa of \'{A}vila or Padre Pio.}
\end{itemize}

A list of recurring topics and their identification in the sound files is a
point of future work.

\subsection{Acoustic Quality}

Despite of the relatively good recording technique and mostly quiet environment,
the resulting acoustic quality varies wildly. Many recordings are clearly
intelligible, while some suffer from various defects like
\begin{itemize}
\item{background noise in varying intensity and bandwidth,}
\item{altered speed,}
\item{clipping or faitness,}
\item{echo,}
\item{overlap of two recordings.}
\end{itemize}

Sometimes the quality changes abruptly inside a file, when recording was stopped
and resumed with different conditions.

Expectably, the age of a recording correlates with its quality but other factors
play a significant role also. Apart of obvious ones like the relative position
of the speaker to the microphone, the acoustics of the room and quality of the
recording device, also the tape brand seems to be quite important.

\section{Processing}

I have focused on two major aspects of processing the material: 1. Making it
accessible to the public and 2. acquiring a complete transcription. The
% akustická úprava, indexy, vyhledávání

%\section*{Acknowledgements}
%
%This document has been adapted from the instructions for the 
%CAC2014 proceedings, which are, in turn based on those for the
%COLING 2014 proceedings, which are, in turn, based on the ACL-2014
%proceedings compiled by Alexander Koller and Yusuke Miyao,
%which are, in turn, based on the instructions for earlier ACL proceedings,
%including those for ACL-2012 by Maggie Li and Michael
%White, those from ACL-2010 by Jing-Shing Chang and Philipp Koehn,
%those for ACL-2008 by Johanna D. Moore, Simone Teufel, James Allan,
%and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal
%Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and
%earlier ACL and EACL formats. Those versions were written by several
%people, including John Chen, Henry S. Thompson and Donald
%Walker. Additional elements were taken from the formatting
%instructions of the {\em International Joint Conference on Artificial
%  Intelligence}.

% The bibliography section below is for illustration only! Include your own bib file like this:
%\bibliography{yourbibfile}

\begin{thebibliography}{}

\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
Alfred~V. Aho and Jeffrey~D. Ullman.
\newblock 1972.
\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
\newblock Prentice-{Hall}, Englewood Cliffs, NJ.

\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
{American Psychological Association}.
\newblock 1983.
\newblock {\em Publications Manual}.
\newblock American Psychological Association, Washington, DC.

\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
{Association for Computing Machinery}.
\newblock 1983.
\newblock {\em Computing Reviews}, 24(11):503--512.

\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
\newblock 1981.
\newblock Alternation.
\newblock {\em Journal of the Association for Computing Machinery},
  28(1):114--133.

\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
Dan Gusfield.
\newblock 1997.
\newblock {\em Algorithms on Strings, Trees and Sequences}.
\newblock Cambridge University Press, Cambridge, UK.

\end{thebibliography}

\end{document}

### 2018-04-09 2018-clarin-pisa/analysis/bin/add-ratios-to-wl.pl
script for data analysis for 2018 clarin

@@ -1,0 +1,20 @@
+#!/usr/bin/perl
+
+# input line format: wordform humanic_count automatic_count
+# with AUTT and HUMT constants for total number of automatic and humanic words,
+# write (humanic_count / HUMT) / (automatic_count / AUTT)
+
+use 5.010;
+use strict;
+use warnings;
+use utf8;
+
+my $HUMT = 620615;
+my $AUTT = 7574994;
+my $w = $AUTT / $HUMT;
+
+while (<>) {
+    chomp;
+    my ($wf, $hum, $aut) = split / /;
+    say join ' ', $wf, $hum, $aut, $w * $hum / $aut;
+}

### 2018-04-09 2018-clarin-pisa/analysis/bin/find-occurrence.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+ag -i '\{[^{]*\boccurrence\W+'"$1"'[^}]*\}' /home/sixtease/skola/phd/webapp/MakonFM/root/static/subs/*.sub.js /home/sixtease/skola/phd/webapp/MakonFM/root/static/subs/prevzate/*.sub.js

### 2018-04-09 2018-clarin-pisa/analysis/bin/getwl.pl

@@ -1,0 +1,17 @@
+#!/usr/bin/perl
+
+use 5.010;
+use strict;
+use warnings;
+use utf8;
+
+my %dict;
+
+while (<>) {
+    /wordform"?\s*:\s*"([^"]*)"/ and $dict{$1}++;
+}
+
+my $n = 0;
+my @sorted = sort keys %dict;
+
+print "$_ $dict{$_}\n" for @sorted;

### 2018-04-09 2018-clarin-pisa/analysis/bin/sort-joined-wl.pl

@@ -1,0 +1,15 @@
+#!/usr/bin/perl
+
+use 5.010;
+use strict;
+use warnings;
+use utf8;
+
+my @lines;
+
+while (<>) {
+    (undef, undef, undef, my $rat) = split / /;
+    push @lines, [+$rat, $_];
+}
+
+print for map $_->[1], sort {$b->[0] <=> $a->[0]} @lines;

### 2018-04-09 2018-clarin-pisa/makon-data.tex
crap to delete for clarin 2018

@@
 I have focused on two major aspects of processing the material: 1. Making it
-accessible to the public and 2. acquiring a complete transcription. The
-% akustická úprava, indexy, vyhledávání
+accessible to the public and 2. acquiring a complete transcription. To a lesser
+extent, I have also worked towards 1. overcoming the acoustic difficulties for
+listeners, 2. topically indexing some recordings and 3. enablind a full-text
+search over the whole corpus.
+
+\subsection{Access to the Corpus}
+
+The speech recordings are available through two sources. First and foremost,
+it is in the Clarin repository, which guarantees it will be easily available no
+matter what fate this project will take.
+
+Secondly, I have created a web application for listening to the recordings
+directly in the browser. The web application is also capable of displaying the
+transcription synchronously. When there is an error in the transcription (which
+is frequently so in case of automatically acquired ones), the user can submit a
+correction.
+
+\subsection{Complete Transcription}
+
+Transcription of the whole corpus has been the main focus for the majority of
+the time spent on the project. The concept was to transcribe a few minutes from
+scratch, train a HMM-based ASR model on that, evaluate the bulk of the corpus
+and feed that to the web application, as a base for the users to correct.
+
+We only have a single speaker, a restricted domain, relatively consistent
+conditions, an extensive collection of the author's original books,
+translations, comments and letters to build a well-fitting language model. So
+even very little training data should give satisfactory results and the error
+rate should drop rapidly as the training data would increase.
+
+Well, the error rate did drop as the training data increased but the accuracy of
+the speech recognition was more of a challenge than I had thought. Fot the most
+part, the users were actually using the web interface to provide a new
+transcription from scratch, ignoring the way too erroneous output of the ASR,
+instead of merely correcting single errors, as was meant.
+
+Still, the community managed to provide over 66 hours of high-quality manual
+transcriptions, which have been used gradually as training data. Now, the
+acoustically intact recordings are automatically transcribed well enough to
+allow reading along and to do full-text search.
+
+The word error rate on our test set is currently 42\% but it is probably lower
+on clean recordings and also most mistakes are in word endings, so the meaning
+is comprehensible. I am currently training a DNN acoustic model, so chances are
+the ASR score will soon rise noticeably.
+
+\subsection{Secondary Processing Attempts}
+
+\subsubsection{Mitigating Acoustic Defects}
+
+In a new version of the web interface (paper pending), I have implemented a
+user-controllable parametric graphical equalizer. This can help significantly in
+the case of narrow-band noise, which occurs on some recordings, predominantly in
+the bass range.
+
+This was a cheap solution for a low-hanging fruit but a more holistic approach
+is
 

### 2018-04-11 2018-clarin-pisa/makon-data.tex
bla
table for clarin 2018

@@
 
+We can take the current transcription %and the few notes some listeners made
+%about topics and their alignment with the audio
+as a source for a baseline.
+Topic is inherently a vague concept, which I shall address when I focus on it
+but for the baseline evaluation, we can take a look at some easily defined cases
+like named entities. The ease about them is that we can look for their specific
+word form and if it is not found, then it means the transcription is not talking
+about it.
+
+Of course, there are exceptions, like saying ``the capital of Japan'' instead of
+``Tokyo'' but firstly, I chose terms where this is not likely to happen and
+secondly, this can only lower the recall of our baseline score, not the
+precision, and we have no way of telling the precision given no annotated data
+for this task.\footnote{The 50-minute test set is way too small for this kind of
+task.}
+
+I have chosen the following terms:
+
+\begin{enumerate}
+\item{Lazarus,}
+\item{Mithra, Mithraism,}
+\item{Satan,}
+\item{St. Teresa,}
+\item{fairy-tale.}
+\end{enumerate}
+
+I chose fairy-tale (Czech: \emph{poh\'{a}dka}) as a recurring topic with a
+unifying word, for an example of a non-named-entity.
+
+The search for stems of these words in the automatic part of the transcription
+yielded results summed up in Table~\ref{tab:topicsearch}. For terms where there
+were more than 20 hits, I have checked the first 20 but each from a different
+file.
+
+\begin{table}[htpb]
+\begin{center}
+\begin{tabular}{|l|l|r|r|r|}
+\hline
+Term & Query & hits in automatic transcription & hits in manual transcription & Precision \\
+\hline
+Lazarus & \texttt{lazar.*} & 16 & 14 & 8/10   \\
+Mithra & \texttt{mithra.*|mitra.*} & 0 & 0 & n/a   \\
+Satan & \texttt{satan.*} & 308 & 129 & 16/20   \\
+St. Teresa & \texttt{terez.*} & 882 & 91 & 15/20   \\
+fairy-tale & \texttt{pohádk.*} & 216 & 55 & 18/20   \\
+\hline
+\multicolumn{4}{|l|}{average precision} & 81.25\%\\
+\hline
+\end{tabular}
+\caption{Text search hits in the automatically-acquired part of the
+transcription}\label{tab:topicsearch}
+\end{center}
+\end{table}
+
 \subsection{Acoustic Quality}
@@
 This was a cheap solution for a low-hanging fruit but a more holistic approach
-is
+is very desirable.
+
+% 7.6\% of the corpus is annotated with 620615 : 7574994 hum : auto
 

### 2018-04-11 2018-clarin-pisa/analysis/bin/find-wordform.sh
some data anal. scripts for clarin2018

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+ag '\{[^{]*\bwordform\W+'"$1"'[^}]*\}' /home/sixtease/skola/phd/webapp/MakonFM/root/static/subs/*.sub.js /home/sixtease/skola/phd/webapp/MakonFM/root/static/subs/prevzate/*.sub.js

### 2018-04-11 2018-clarin-pisa/analysis/bin/play-ackd.pl

@@ -1,0 +1,15 @@
+#!/usr/bin/perl
+
+use 5.010;
+use strict;
+use warnings;
+use utf8;
+
+while (<>) {
+    next if not /timestamp\D+(\d+)/;
+    my $ts = $1;
+    m{([^/]+)\.sub\.js};
+    my $stem = $1;
+    $ts--;
+    system "mplayer -ss $ts /home/sixtease/kroupy/Music/Makon/all/mp3/$stem.mp3";
+}

### 2018-04-11 2018-clarin-pisa/analysis/bin/split-hum-aut.pl

@@ -1,0 +1,23 @@
+#!/usr/bin/perl
+
+use 5.010;
+use strict;
+use warnings;
+use utf8;
+
+my $hum_fn = shift @ARGV;
+my $aut_fn = shift @ARGV;
+open my $hum_fh, '>', $hum_fn or die "Cannot open hum file $hum_fn: $!";
+open my $aut_fh, '>', $aut_fn or die "Cannot open aut file $aut_fn: $!";
+
+local $/ = '}';
+while (<>) {
+    if (/occurrence/) {
+        if (/humanic/) {
+            print {$hum_fh} $_;
+        }
+        else {
+            print {$aut_fh} $_;
+        }
+    }
+}

### 2018-04-12 2018-clarin-pisa/makon-data.tex
some blurb and data for clarin 2018

@@
 
+Most of the previous work was focused on acquiring the automatic transcription
+and development of a web interface to allow users to access the talks and
+provide corrections to the existing transcription, in order to have high-quality
+transcription as well as to gather more training data.
+
+The actual content of the corpus with respect to topics, references and
+statements, is on one hand quite well known firstly because there are many books
+by the author with descriptive indexes like any other book, which largely
+correspond to what is said in the talks, secondly because the the domain stays
+more or less consistent across the whole set. On the other hand, it is only
+known vaguely and a systematic effort to analyze it is yet to be carried out.
+
 \section{Content of the Corpus}
@@
 \item{milestone personal experiences, like the one in the concentration camp,}
-\item{explaining symbolism, the apostles as symbols for human abilities,}
+\item{explaining symbolism, like the apostles as symbols for human abilities,}
 \item{references to specific people, like St. Teresa of \'{A}vila or Padre Pio.}
@@
 as a source for a baseline.
-Topic is inherently a vague concept, which I shall address when I focus on it
+Topic is inherently a vague concept %, which I shall address when I focus on it
 but for the baseline evaluation, we can take a look at some easily defined cases
@@
 secondly, this can only lower the recall of our baseline score, not the
-precision, and we have no way of telling the precision given no annotated data
+precision, and we have no way of telling the recall given no annotated data
 for this task.\footnote{The 50-minute test set is way too small for this kind of
-task.}
+task. Another source could be manual notes available to some of the recordings
+but preparing them for this purpose requires some more effort.}
 
-I have chosen the following terms:
+I have chosen the following example topics:
 
@@
 
+The precision of over 80\% seems quite promising and hints that the current
+quality of transcription allows for reasonable searching through the corpus.
+However, the hits were all clearly audible. I met none in an acoustically defect
+(see the next subsection) file, which leads to the hypothesis that those are not
+yet searchable to a distantly satisfying level.
+
+\subsubsection{Correspondence between Talks and Books}
+
+The proclamation at the end of the introduction~\ref{intro}, that the
+topics of the talks correspond largely to that of the concurrently written
+books, deserves a kind of support. Mako\v{n} translated the book \emph{El
+Castillo Interior}, English \emph{The Interior Castle} by St. Teresa of Ávila in
+1977. For most recordings, the year of their acquisition is known.
+Figure~\ref{fig:teresa-year} shows how many hits for the name of the Saint are
+present in the transcription by recording year.
+
 \subsection{Acoustic Quality}
@@
 Sometimes the quality changes abruptly inside a file, when recording was stopped
-and resumed with different conditions.
+and resumed with different conditions. This prevents us from classifying
+individual recordings by their acoustic quality.
 
 Expectably, the age of a recording correlates with its quality but other factors
-play a significant role also. Apart of obvious ones like the relative position
-of the speaker to the microphone, the acoustics of the room and quality of the
-recording device, also the tape brand seems to be quite important.
+play a significant role also. Most are obvious like the relative position
+of the speaker to the microphone, the acoustics of the room, quality of the
+recording device and also the tape brand. Unfortunately, none of these can be
+said with certainty.
 

### 2018-04-12 2018-clarin-pisa/analysis/bin/get-humanic.pl

@@ -1,23 +1,0 @@
-#!/usr/bin/perl
-
-use 5.010;
-use strict;
-use warnings;
-use utf8;
-
-my $hum_fn = shift @ARGV;
-my $aut_fn = shift @ARGV;
-open my $hum_fh, '>', $hum_fn or die "Cannot open hum file $hum_fn: $!";
-open my $aut_fh, '>', $aut_fn or die "Cannot open aut file $aut_fn: $!";
-
-local $/ = '}';
-while (<>) {
-    if (/occurrence/) {
-        if (/humanic/) {
-            print {$hum_fh} $_;
-        }
-        else {
-            print {$aut_fh} $_;
-        }
-    }
-}

### 2018-04-13 2018-clarin-pisa/makon-data.tex
fix hrad nitra's year
add figure for teresa-year correlation
move things around
shortening and polishing clarin2018

@@
   briefly outlines the system for processing the corpus, which has been
-  published in a previous work, and elaborates on the corpus, the process of its
-  acquisition and its annotation.
+  published in a previous work, and elaborates on the corpus. A baseline for
+  topical analysis is presented.
 \end{abstract}
@@
 between 2010 and 2012. A complete transcription was obtained using a dedicated
-ASR system and the work of the community around K.M's legacy. The corpus is
+ASR system and the work of the community around K.M.'s legacy. The corpus is
 about 1000 hours in total length, of which about 66 have been transcribed
@@
 The actual content of the corpus with respect to topics, references and
-statements, is on one hand quite well known firstly because there are many books
-by the author with descriptive indexes like any other book, which largely
-correspond to what is said in the talks, secondly because the the domain stays
+statements, is on one hand quite well known
+because the the domain stays
 more or less consistent across the whole set. On the other hand, it is only
@@
 
-\section{Content of the Corpus}
-
-\subsection{Karel Mako\v{n}}
+\section{Author}
 
@@
 book that ``this life is the bridge to the eternal life''. The thought
-captivated him and gave him a new life direction, which also led to several
+captivated him and gave him a new life direction, which led to several
 ecstasies.
@@
 The next culmination happened in 1939, after nine years of constant prayer for a
-stronger love to God, when he was doported to the nazi concentration camp in
+stronger love to God, when he was deported to the nazi concentration camp in
 Sachsenhausen, as a university student. Here, he was promised certain death by a
 Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
-life completely to God and the attacker turned around in awe and fled.
+life completely to God and the attacker unexplainably turned around in awe and
+fled.
 
@@
 
-\subsection{Giving and Recording the Talks}
+\section{Giving and Recording the Talks}
 
@@
 One of the fortunate consequences of the seclusion of the talks is that there is
-very little background noise. Also, most of the recordings
+very little background noise on most of the recordings. Also, most of them
 % TODO: kolik je od Elgra?
@@
 
-\subsection{Topics}
+\section{Processing}
+
+We don't have the resources needed to manually annotate the corpus like
+e.g.~\ref{japsponta}, but there are lay
+volunteers willing to work on the project for the cause of tending Mako\v{n}'s
+legacy.
+
+A complete transcription is essential or at least helpful for any kind of
+further processing.  Therefore,
+I have focused on two major aspects of processing the material: 1. Making it
+accessible to the public and 2. acquiring a complete transcription.
+
+\subsection{Access to the Corpus}
+
+The speech recordings are available through two sources. First and foremost,
+it is in the Clarin repository, which guarantees it will be easily available no
+matter what fate this project will take.
+
+Secondly, I have created a web application for listening to the recordings
+directly in the browser. The web application is also capable of displaying the
+transcription synchronously. When there is an error in the transcription (which
+is frequently so in case of automatically acquired ones), the user can submit a
+correction.
+
+\subsection{Complete Transcription}
+
+Transcription of the whole corpus has been the main focus for the majority of
+the time spent on the project. The concept was to transcribe a few minutes from
+scratch, train a HMM-based ASR model on that, evaluate the bulk of the corpus
+and feed that to the web application, as a base for the users to correct.
+
+We only have a single speaker, a restricted domain, relatively consistent
+conditions, an extensive collection of the author's original books,
+translations, comments and letters to build a fitting language model. So
+even very little training data should give satisfactory results and the error
+rate should drop rapidly as the training data would increase.
+
+Well, the error rate did drop as the training data increased but the accuracy of
+the speech recognition was more of a challenge than I had thought. For the most
+part, the users were actually using the web interface to provide a new
+transcription from scratch, ignoring the way too erroneous output of the ASR,
+instead of merely correcting single errors, as was meant.
+
+Still, the community managed to provide over 66 hours of high-quality manual
+transcriptions, which have been used gradually as training data. Now, the
+acoustically intact recordings are automatically transcribed well enough to
+allow reading along and to do full-text search.
+
+The word error rate on our test set is currently 42\% but it is probably lower
+on clean recordings and also most mistakes are in word endings, so the meaning
+is comprehensible. I am currently working on a DNN acoustic model, so chances
+are the ASR score will soon rise.
+
+\subsubsection{Acoustic Quality}
+
+Despite of the relatively good recording technique and mostly quiet environment,
+the resulting acoustic quality varies wildly. Many recordings are clearly
+intelligible, while some suffer from various defects like
+1. background noise in varying intensity and bandwidth,
+2. altered speed,
+3. clipping or faitness,
+4. echo,
+5. overlap of two recordings.
+
+The varying acoustic quality is the most serious obstacle to the quality of the
+automatic transcription. But it is also detrimental to the manual transcription.
+
+To mitigate this, the web interface features since lately a user-controllable
+graphical equalizer, which can help especially in the case of narrow-band noise.
+
+\subsection{Secondary Processing Attempts}
+
+Some limited effort was invested into further processing beyond plain
+transcription. I have created a simple search engine over the transcription that
+yields deep links into relevant passages in the audio. Elasticsearch was used
+for this purpose, with Czech algorithmic stemmer.
+
+Another effort was processing manually written indexes that are available for a
+fraction of the recordings. These have been partly hand-written, partly typed
+and are aligned using the positions of the device counter. Most of these are
+scanned and some limited experiments were carried out to apply OCR. The explicit
+mapping of the indexes with corresponding recordings and time-positions is yet
+to be done.
+
+\section{Topics}
 
@@
 point of future work.
-
 We can take the current transcription %and the few notes some listeners made
@@
 like named entities. The ease about them is that we can look for their specific
-word form and if it is not found, then it means the transcription is not talking
-about it.
+word form and if it is not found, then it means the transcription is not
+covering it.
 
@@
 I have chosen the following example topics:
-
-\begin{enumerate}
-\item{Lazarus,}
-\item{Mithra, Mithraism,}
-\item{Satan,}
-\item{St. Teresa,}
-\item{fairy-tale.}
-\end{enumerate}
+1. \emph{Lazarus},
+2. \emph{Mithra, Mithraism},
+3. \emph{Satan},
+4. \emph{St. Teresa},
+5. \emph{fairy-tale}.
 
@@
 However, the hits were all clearly audible. I met none in an acoustically defect
-(see the next subsection) file, which leads to the hypothesis that those are not
-yet searchable to a distantly satisfying level.
+file, which leads to the hypothesis that those are not yet searchable to a
+distantly satisfying level.
 
@@
 
-The proclamation at the end of the introduction~\ref{intro}, that the
-topics of the talks correspond largely to that of the concurrently written
-books, deserves a kind of support. Mako\v{n} translated the book \emph{El
-Castillo Interior}, English \emph{The Interior Castle} by St. Teresa of Ávila in
-1977. For most recordings, the year of their acquisition is known.
-Figure~\ref{fig:teresa-year} shows how many hits for the name of the Saint are
-present in the transcription by recording year.
+K.M. was writing, translating and commenting books most of his life. We can
+assume that the book he was just working on influenced the topic of his talks. 
+For on, he translated the book \emph{El Castillo Interior}, English \emph{The
+Interior Castle} by St. Teresa of Ávila in 1988. For most recordings, the year
+of their acquisition is known.  Figure~\ref{fig:teresa-year} shows how many hits
+for the name of the Saint are present in the transcription by recording year.
 
-\subsection{Acoustic Quality}
+\begin{figure}[htpb]
+\includegraphics[scale=0.6]{rc/teresa-by-year.eps}
+\caption{Number of hits for the query \texttt{terez.*} in the transcription by
+alleged year of the recording}
+\label{fig:teresa-year}
+\end{figure}
 
-Despite of the relatively good recording technique and mostly quiet environment,
-the resulting acoustic quality varies wildly. Many recordings are clearly
-intelligible, while some suffer from various defects like
-\begin{itemize}
-\item{background noise in varying intensity and bandwidth,}
-\item{altered speed,}
-\item{clipping or faitness,}
-\item{echo,}
-\item{overlap of two recordings.}
-\end{itemize}
-
-Sometimes the quality changes abruptly inside a file, when recording was stopped
-and resumed with different conditions. This prevents us from classifying
-individual recordings by their acoustic quality.
-
-Expectably, the age of a recording correlates with its quality but other factors
-play a significant role also. Most are obvious like the relative position
-of the speaker to the microphone, the acoustics of the room, quality of the
-recording device and also the tape brand. Unfortunately, none of these can be
-said with certainty.
-
-\section{Processing}
-
-I have focused on two major aspects of processing the material: 1. Making it
-accessible to the public and 2. acquiring a complete transcription. To a lesser
-extent, I have also worked towards 1. overcoming the acoustic difficulties for
-listeners, 2. topically indexing some recordings and 3. enablind a full-text
-search over the whole corpus.
-
-\subsection{Access to the Corpus}
-
-The speech recordings are available through two sources. First and foremost,
-it is in the Clarin repository, which guarantees it will be easily available no
-matter what fate this project will take.
-
-Secondly, I have created a web application for listening to the recordings
-directly in the browser. The web application is also capable of displaying the
-transcription synchronously. When there is an error in the transcription (which
-is frequently so in case of automatically acquired ones), the user can submit a
-correction.
-
-\subsection{Complete Transcription}
-
-Transcription of the whole corpus has been the main focus for the majority of
-the time spent on the project. The concept was to transcribe a few minutes from
-scratch, train a HMM-based ASR model on that, evaluate the bulk of the corpus
-and feed that to the web application, as a base for the users to correct.
-
-We only have a single speaker, a restricted domain, relatively consistent
-conditions, an extensive collection of the author's original books,
-translations, comments and letters to build a well-fitting language model. So
-even very little training data should give satisfactory results and the error
-rate should drop rapidly as the training data would increase.
-
-Well, the error rate did drop as the training data increased but the accuracy of
-the speech recognition was more of a challenge than I had thought. Fot the most
-part, the users were actually using the web interface to provide a new
-transcription from scratch, ignoring the way too erroneous output of the ASR,
-instead of merely correcting single errors, as was meant.
-
-Still, the community managed to provide over 66 hours of high-quality manual
-transcriptions, which have been used gradually as training data. Now, the
-acoustically intact recordings are automatically transcribed well enough to
-allow reading along and to do full-text search.
-
-The word error rate on our test set is currently 42\% but it is probably lower
-on clean recordings and also most mistakes are in word endings, so the meaning
-is comprehensible. I am currently training a DNN acoustic model, so chances are
-the ASR score will soon rise noticeably.
-
-\subsection{Secondary Processing Attempts}
-
-\subsubsection{Mitigating Acoustic Defects}
-
-In a new version of the web interface (paper pending), I have implemented a
-user-controllable parametric graphical equalizer. This can help significantly in
-the case of narrow-band noise, which occurs on some recordings, predominantly in
-the bass range.
-
-This was a cheap solution for a low-hanging fruit but a more holistic approach
-is very desirable.
+The peak around the year 1988 supports the speculation that topics in the talks
+correlate with those of books written at the same time.
 

### 2018-04-13 2018-clarin-pisa/makon-data.tex
conclusion clarin2018

@@
 
+\section{Conclusion}
+
+With its length of over 1000 hours spoken by a single speaker over three
+decades, the corpus of Karel Mako\v{n} is a unique piece in the Lindat/Clarin
+repository. The consistent domain of Christian mystic is another of its
+distinguishing features. A basis for detailed topical analysis is presented
+exploiting the existing transcription and the author's books.
+
 % 7.6\% of the corpus is annotated with 620615 : 7574994 hum : auto

### 2018-04-13 2018-clarin-pisa/makon-data.tex
citations clarin2018

@@
 \usepackage{covington} % if needed, for linguistic examples
+
+\begin{filecontents}{citace.bib}
+@inproceedings{maekawa2000spontaneous,
+  title={Spontaneous Speech Corpus of Japanese.},
+  author={Maekawa, Kikuo and Koiso, Hanae and Furui, Sadaoki and Isahara, Hitoshi},
+  booktitle={LREC},
+  year={2000},
+  organization={Citeseer}
+}
+@inproceedings{kruuza2012making,
+  title={Making Community and ASR Join Forces in Web Environment},
+  author={Kr\r{u}za, Old{\v{r}}ich and Peterek, Nino},
+  booktitle={International Conference on Text, Speech and Dialogue},
+  pages={415--421},
+  year={2012},
+  organization={Springer}
+}
+@article{hajek2007cesky,
+  title={\v{C}esk\'{y} mystik Karel Mako\v{n}},
+  author={Jurik H\'{a}jek},
+  journal={Dingir},
+  year={2007},
+  volume={2007/4},
+  pages={142--143},
+  publisher={Dingir, s.r.o.},
+  ISSN={1212-1371},
+  editor={Zden\v{e}k Vojt\'{i}\v{s}ek}
+}
+\end{filecontents}
 
@@
 Most of the previous work was focused on acquiring the automatic transcription
-and development of a web interface to allow users to access the talks and
-provide corrections to the existing transcription, in order to have high-quality
-transcription as well as to gather more training data.
+and development of a web interface~\cite{kruuza2012making} to allow users to
+access the talks and provide corrections to the existing transcription, in order
+to have high-quality transcription as well as to gather more training data.
 
@@
 
-The author of the talks, Mr. Karel Mako\v{n} *1912 \textdagger{}1993, was giving
-talks in a strife to share his awareness of the proverbial meaning of life and
-to give a manual to eternal life after his release from a concentration camp
-during the WWII.
+The author of the talks, Mr. Karel Mako\v{n}~\cite{hajek2007cesky} *1912
+\textdagger{}1993, was giving talks in a strife to share his awareness of the
+proverbial meaning of life and to give a manual to eternal life after his
+release from a concentration camp during the WWII.
 
-His steep spiritual path began in early childhood when he was
-subject to repeated arm surgery without anesthesy. The pain the child suffered
-made him escape mentally from inside his body, which led him to what he calls
-``the yellow light''. As a result, he felt obliged always to {\em do the right
-thing}.
+His steep spiritual path began in early childhood when he was subject to
+repeated arm surgery without anesthesy. The pain the child suffered made him
+escape mentally from inside his body, which led him to what he calls ``the
+yellow light''. As a result, he felt obliged always to {\em do the right thing}.
 
@@
 We don't have the resources needed to manually annotate the corpus like
-e.g.~\ref{japsponta}, but there are lay
+e.g.~\cite{maekawa2000spontaneous}, but there are lay
 volunteers willing to work on the project for the cause of tending Mako\v{n}'s
@@
 on clean recordings and also most mistakes are in word endings, so the meaning
-is comprehensible. I am currently working on a DNN acoustic model, so chances
-are the ASR score will soon rise.
+is comprehensible. I am currently working on a DNN acoustic
+model~\cite{hannun2014deep}, so chances are the ASR score will soon rise.
 
@@
 A list of recurring topics and their identification in the sound files is a
-point of future work.
+point of future work with similarities to~\cite{skorkovska2011automatic}.
 We can take the current transcription %and the few notes some listeners made
@@
 
-\begin{thebibliography}{}
-
-\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
-Alfred~V. Aho and Jeffrey~D. Ullman.
-\newblock 1972.
-\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
-\newblock Prentice-{Hall}, Englewood Cliffs, NJ.
-
-\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
-{American Psychological Association}.
-\newblock 1983.
-\newblock {\em Publications Manual}.
-\newblock American Psychological Association, Washington, DC.
-
-\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
-{Association for Computing Machinery}.
-\newblock 1983.
-\newblock {\em Computing Reviews}, 24(11):503--512.
-
-\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
-Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
-\newblock 1981.
-\newblock Alternation.
-\newblock {\em Journal of the Association for Computing Machinery},
-  28(1):114--133.
-
-\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
-Dan Gusfield.
-\newblock 1997.
-\newblock {\em Algorithms on Strings, Trees and Sequences}.
-\newblock Cambridge University Press, Cambridge, UK.
-
-\end{thebibliography}
+\bibliography{citace}
 

### 2018-04-13 2018-clarin-pisa/makon-data.tex
more clarin text

@@
 \usepackage{covington} % if needed, for linguistic examples
+\usepackage{makecell}
 
@@
 \begin{abstract}
-  We present a detailed view on the spoken corpus of Karel Mako\v{n}. This paper
+  We present the spoken corpus of Karel Mako\v{n}. This paper
   briefly outlines the system for processing the corpus, which has been
@@
 escape mentally from inside his body, which led him to what he calls ``the
-yellow light''. As a result, he felt obliged always to {\em do the right thing}.
+yellow light''.
 
-This culminated in the age of 17, in a perceived life crissis, when he read in a
-book that ``this life is the bridge to the eternal life''. The thought
-captivated him and gave him a new life direction, which led to several
-ecstasies.
-
-The next culmination happened in 1939, after nine years of constant prayer for a
+His path culminated in 1939, after nine years of constant prayer for a
 stronger love to God, when he was deported to the nazi concentration camp in
@@
 direct antagonism with the communist regime. The opposition was strictly
-unidirectional though, as Mako\v{n} also understood the purpose of the regime
-and was far from wasting his energy on a fight against it. His way of dealing
+unidirectional though, as Mako\v{n}
+was far from wasting his energy on a fight against it. His way of dealing
 with the conflict was by staying out of wide popularity. Hence, most of the
@@
 have been taken by a single person who used the available state-of-the-art
-amateur recording technology and methods. He was also very systematic in
-labeling and archiving the recordings. As a result, the majority of the base for
-digitization was a set of reel-to-reel tapes labeled with ordinal indexes and a
-set of cassettes labeled by year and ordinal index. The reel-to-reel tapes have
-been acquired in a generous speed of 9cm/second, which granted them a very good
-shield against the inevitable aging.
+amateur recording technology and methods. This person was also very systematic
+in labeling and archiving the recordings. As a result, the majority of the base
+for digitization was a set of reel-to-reel tapes labeled with ordinal indexes
+and a set of cassettes labeled by year and ordinal index. The reel-to-reel tapes
+have been recorded in a generous speed of 9cm/second, which granted them a very
+good shield against the inevitable aging.
 
@@
 
-The speech recordings are available through two sources. First and foremost,
+The speech recordings are available through two sources. Firstly,
 it is in the Clarin repository, which guarantees it will be easily available no
@@
 directly in the browser. The web application is also capable of displaying the
-transcription synchronously. When there is an error in the transcription (which
-is frequently so in case of automatically acquired ones), the user can submit a
+transcription synchronously. When there is an error in the transcription (and
+the automatic transcription has plenty of errors), the user can submit a
 correction.
@@
 
-We only have a single speaker, a restricted domain, relatively consistent
-conditions, an extensive collection of the author's original books,
-translations, comments and letters to build a fitting language model. So
-even very little training data should give satisfactory results and the error
-rate should drop rapidly as the training data would increase.
-
-Well, the error rate did drop as the training data increased but the accuracy of
-the speech recognition was more of a challenge than I had thought. For the most
-part, the users were actually using the web interface to provide a new
-transcription from scratch, ignoring the way too erroneous output of the ASR,
-instead of merely correcting single errors, as was meant.
-
-Still, the community managed to provide over 66 hours of high-quality manual
+The community managed to provide over 66 hours of high-quality manual
 transcriptions, which have been used gradually as training data. Now, the
@@
 automatic transcription. But it is also detrimental to the manual transcription.
-
 To mitigate this, the web interface features since lately a user-controllable
@@
 precision, and we have no way of telling the recall given no annotated data
-for this task.\footnote{The 50-minute test set is way too small for this kind of
-task. Another source could be manual notes available to some of the recordings
-but preparing them for this purpose requires some more effort.}
+for this task anyway.\footnote{The 50-minute test set is way too small for this
+kind of task. Another source could be manual notes available to some of the
+recordings but preparing them for this purpose requires some more effort.}
 
 I have chosen the following example topics:
-1. \emph{Lazarus},
-2. \emph{Mithra, Mithraism},
-3. \emph{Satan},
-4. \emph{St. Teresa},
-5. \emph{fairy-tale}.
-
+1. \textbf{Lazarus},
+2. \textbf{Mithra, Mithraism},
+3. \textbf{Satan},
+4. \textbf{St. Teresa},
+5. \textbf{fairy-tale}.
 I chose fairy-tale (Czech: \emph{poh\'{a}dka}) as a recurring topic with a
@@
 
-The search for stems of these words in the automatic part of the transcription
+The search for stems of these words in the transcription
 yielded results summed up in Table~\ref{tab:topicsearch}. For terms where there
@@
 \begin{center}
-\begin{tabular}{|l|l|r|r|r|}
+\begin{tabular}{|l|l|r|r|r|r|}
 \hline
-Term & Query & hits in automatic transcription & hits in manual transcription & Precision \\
+Term &
+Query &
+\makecell{hits in automatic\\
+transcription} &
+\makecell{expected hits\\
+in automatic\\
+transcription} &
+\makecell{hits in manual\\
+transcription} &
+Precision \\
 \hline
-Lazarus & \texttt{lazar.*} & 16 & 14 & 8/10   \\
-Mithra & \texttt{mithra.*|mitra.*} & 0 & 0 & n/a   \\
-Satan & \texttt{satan.*} & 308 & 129 & 16/20   \\
-St. Teresa & \texttt{terez.*} & 882 & 91 & 15/20   \\
-fairy-tale & \texttt{pohádk.*} & 216 & 55 & 18/20   \\
+Lazarus & \texttt{lazar.*} & 16 & 171 & 14 & 8/10   \\
+Mithra & \texttt{mithra.*|mitra.*} & 0 & 0 & 0 & n/a   \\
+Satan & \texttt{satan.*} & 308 & 1575 & 129 & 16/20   \\
+St. Teresa & \texttt{terez.*} & 882 & 1111 & 91 & 15/20   \\
+fairy-tale & \texttt{pohádk.*} & 216 & 671 & 55 & 18/20   \\
 \hline
-\multicolumn{4}{|l|}{average precision} & 81.25\%\\
+\multicolumn{5}{|l|}{average precision} & 81.25\%\\
 \hline
 \end{tabular}
-\caption{Text search hits in the automatically-acquired part of the
-transcription}\label{tab:topicsearch}
+\caption{Text search hits in the transcription; precision only calculated from
+automatic transcription}\label{tab:topicsearch}
 \end{center}
@@
 distantly satisfying level.
+
+7.6\% of the total words in the corpus are manually transcribed.
 

### 2018-04-14 2018-clarin-pisa/makon-data.tex
clarin2018: expanded abstract

@@
 \begin{abstract}
-  We present the spoken corpus of Karel Mako\v{n}. This paper
-  briefly outlines the system for processing the corpus, which has been
-  published in a previous work, and elaborates on the corpus. A baseline for
-  topical analysis is presented.
+  We present the spoken corpus of Karel Mako\v{n}, a corpus of spontaneous
+  speech by a single speaker in Czech language within the domain of Christian
+  mystic, with partly automatic, partly manual transcription obtained through a
+  web application made for correcting the output of ASR.
+  This paper briefly outlines the system for processing the corpus and
+  elaborates on the corpus itself. A baseline for topical analysis is presented.
 \end{abstract}

### 2018-04-14 2018-clarin-pisa/makon-data.tex
clarin2018: polishing

@@
 
-The mystical, religious and spiritual nature of Mako\v{n}'s talks put K.M. in
+The mystical, religious and spiritual nature of his talks put K.M. in
 direct antagonism with the communist regime. The opposition was strictly
@@
 talks have been given during regular meetings in various places of
-Czechoslovakia and later the Czech Republic, in a close circle of friends.
+Czechoslovakia and later the Czech Republic, in a narrow circle of friends.
 
@@
 The community managed to provide over 66 hours of high-quality manual
-transcriptions, which have been used gradually as training data. Now, the
-acoustically intact recordings are automatically transcribed well enough to
-allow reading along and to do full-text search.
+transcriptions through the web application and which have been used gradually
+as training data. Now, the acoustically intact recordings are automatically
+transcribed well enough to allow reading along and to do full-text search.
 
@@
 \caption{Text search hits in the transcription; precision only calculated from
-automatic transcription}\label{tab:topicsearch}
+automatic part}\label{tab:topicsearch}
 \end{center}
@@
 
-7.6\% of the total words in the corpus are manually transcribed.
+7.6\% of the total words in the corpus are manually transcribed. The column
+``expected hits'' interpolates how many occurrences should appear given equal
+frequency. Oddly enough, the ratio of the expected and actual hits varies a lot
+among the search terms. It is almost certain that at least for some of them, the
+recall is very low.
 
@@
 assume that the book he was just working on influenced the topic of his talks. 
-For on, he translated the book \emph{El Castillo Interior}, English \emph{The
+For one, he translated the book \emph{El Castillo Interior}, English \emph{The
 Interior Castle} by St. Teresa of Ávila in 1988. For most recordings, the year

### 2018-05-21 2018-ftc-vancouver/webapp.tex
spelling errors as pointed out by reviewer

@@
 The best widespread match to our task is that of creating an application for
-transcribing speech recordings. Let's compare the two tasks, pointing out the
+transcribing speech recordings. Let us compare the two tasks, pointing out the
 main points of difference. For reference, we take (1)
@@
         occurrence: the word with contextual capitalization and
-        punctuation as it appeard in the text (currently being edited as the
+        punctuation as it appeared in the text (currently being edited as the
         selected initial letter reveals),
@@
     \begin{itemize}
-    \item{textarea prefilled with the current transcription,}
+    \item{text area prefilled with the current transcription,}
     \item{playback button that plays the corresponding segment,}
@@
 
-Although this may not seem like a big deal, it does make affect the
-responsiveness and memory footprint of the page.
+Although this may not seem like a big deal, it does affect the responsiveness
+and memory footprint of the page.
 
@@
 The performance gain for the majority of the usage time outweighs the possible
-errors in the corner cases, moreso since the eventual errors are not critical
+errors in the corner cases, more so since the eventual errors are not critical
 and mostly remedied by further playback.
@@
 setting with that of transcription apps in section~\ref{ssec:diff:trans},
-pointing out lessons to learn. Let's now look at some specific points and their
+pointing out lessons to learn. Let us now look at some specific points and their
 actual (lack of) implementation.
@@
 One of the features that Transcribe, the only commercial tool in our reference
-list, offers is setting up keyboard shortcuts for common words. We haven't
+list, offers is setting up keyboard shortcuts for common words. We have not
 implemented this because ideally, common words should be covered by the speech
@@
 to whole words. In this mode, the transcription to correct is shown in a
-textarea and the global playback controls are replaced by those that only allow
+text area and the global playback controls are replaced by those that only allow
 playback of audio corresponding to the selected transcription.
 
-Once the user believes that the content of the textarea corresponds precisely
+Once the user believes that the content of the text area corresponds precisely
 to the words uttered, she hits the {\em save} button or the \texttt{ctrl+enter}
@@
 and pushed to a CDN, on the other hand it is merged into the transcription word
-array in the JavaScript application. This redundancy warrants that we don't have
-to reload the whole transcription every time a segment is corrected.
+array in the JavaScript application. This redundancy warrants that we do not
+have to reload the whole transcription every time a segment is corrected.
 
@@
 This is because the transcriptions have a long cache time because normally, they
-don't change at all. At the page load, the versions of all transcriptions are
+do not change at all. At the page load, the versions of all transcriptions are
 loaded and used as cache busters. This enables us to use an external CDN and
@@
 basically it is what we know from database design as the normal
-representation.\cite{codd1970relational} It is the first time this approach
+representation\cite{codd1970relational}. It is the first time this approach
 reached the web front-end in such degree of popularity though.
@@
 for listening to recordings and correcting their transcription is nice but one
-that's really easy to use and inviting to people to submit corrections is nicer.
+that is really easy to use and inviting to people to submit corrections is nicer.
 
@@
 Another point of near-future endeavor is higher-level work with the contents. By
-this we mean that we'd like to use both automatic processing methods and the
+this we mean that we wouldd like to use both automatic processing methods and the
 users to do semantic analysis of the talks: What topic is covered where? What

### 2018-05-21 2018-ftc-vancouver/webapp.tex
zmínka o ostatních komunitních korpusových projektech (FTC2018)

@@
   editor={Zden\v{e}k Vojt\'{i}\v{s}ek}
+}
+@inproceedings{ide2010manually,
+  title={The manually annotated sub-corpus: A community resource for and by the people},
+  author={Ide, Nancy and Fellbaum, Christiane and Baker, Collin and Passonneau, Rebecca},
+  booktitle={Proceedings of the ACL 2010 conference short papers},
+  pages={68--73},
+  year={2010},
+  organization={Association for Computational Linguistics}
+}
+@article{bojar2008czeng,
+  title={CzEng 0.7: Parallel Corpus with Community-Supplied Translations},
+  author={Bojar, Ond{\v{r}}ej and Jan{\'\i}{\v{c}}ek, Miroslav and {\v{C}}e{\v{s}}ka, Pavel and Be{\v{n}}a, Peter and others},
+  journal={LREC 2008},
+  year={2008}
+}
+@inproceedings{5494979,
+  author={M. Marge and S. Banerjee and A. I. Rudnicky},
+  booktitle={2010 IEEE International Conference on Acoustics, Speech and Signal Processing},
+  title={Using the Amazon Mechanical Turk for transcription of spoken language},
+  year={2010},
+  volume={},
+  number={},
+  pages={5270-5273},
+  keywords={linguistics;natural language processing;speech processing;Amazon mechanical turk service;MTurk marketplace;Rover voting scheme;speaker demographics;spoken language transcription;transcription guideline;Computer science;Costs;Demography;Gold;Guidelines;Humans;Natural languages;Recruitment;Speech synthesis;Streaming media;crowd sourcing;speech transcription},
+  doi={10.1109/ICASSP.2010.5494979},
+  ISSN={1520-6149},
+  month={March},
+}
+@article{reese2010wikicorpus,
+  title={Wikicorpus: A word-sense disambiguated multilingual wikipedia corpus},
+  author={Reese, Samuel and Boleda, Gemma and Cuadros, Montse and Rigau, German},
+  year={2010},
+  publisher={Citeseer}
 }
@@
 
-% wreally: pause interval
-%          auto loop
-%          templates
-%          pedál
-%          Fka
-%          bez alignmentu
-
 The application we describe is a part of a larger system that deals with
@@
 However, we can compare single aspects found in other applications.
+
+\subsection{Corpora}
+
+Our project is not the first community-driven corpus annotation. We can mention
+the Manually annotated sub-corpus\cite{ide2010manually}, where annotations of
+various kinds are gathered from volunteers, or the
+Wikicorpus\cite{reese2010wikicorpus}, a corpus of Wikipedia articles with some
+linguistic annotation. Our project may reach profound similarities with these in
+the future, when we no longer focus on the transcription itself but rather on
+annotation. There is also CzEng\cite{bojar2008czeng}, the Czech-English Parallel
+Corpus, where a large part of the translation is provided by volunteers. The
+similarity in setting is considerable as both projects involve a
+machine-produced erroneous derivative of the original material (in our case
+audio transcriptions, in the case of CzEng Czech translations of English texts),
+and a community of volunteers corrects these. But the specifics of the projects
+bring different challenges and dictate different approaches. Finally, Marge
+2009\cite{5494979} investigates using The Mechanical Turk to obtain audio
+transcriptions.
 

### 2018-05-21 2018-ftc-vancouver/webapp.tex
more reviewer feedback adaptations

@@
   year={2013}
+}
+@article{mihalcea2004building,
+  title={Building sense tagged corpora with volunteer contributions over the web},
+  author={Mihalcea, Rada and Chklovski, Timothy},
+  journal={Recent Advances in Natural Language Processing III: Selected Papers from RANLP 2003},
+  volume={260},
+  pages={357},
+  year={2004},
+  publisher={John Benjamins Publishing}
 }
@@
 
-Basically the meat of his teachings is systematically encompassed in his written
-works, whilst the talks contain the sauce: talks tailored to the audience,
+Mako\v{n}'s written works present his teachings in a systematic, comprehensible
+fashion, while the recordings offer bonuses: talks tailored to the audience,
 answers to questions, personal experiences, behind-the-scenes to the books etc.
+The archive is freely accessible under the CC-BY license.
 
@@
 However, we can compare single aspects found in other applications.
-
-\subsection{Corpora}
-
-Our project is not the first community-driven corpus annotation. We can mention
-the Manually annotated sub-corpus\cite{ide2010manually}, where annotations of
-various kinds are gathered from volunteers, or the
-Wikicorpus\cite{reese2010wikicorpus}, a corpus of Wikipedia articles with some
-linguistic annotation. Our project may reach profound similarities with these in
-the future, when we no longer focus on the transcription itself but rather on
-annotation. There is also CzEng\cite{bojar2008czeng}, the Czech-English Parallel
-Corpus, where a large part of the translation is provided by volunteers. The
-similarity in setting is considerable as both projects involve a
-machine-produced erroneous derivative of the original material (in our case
-audio transcriptions, in the case of CzEng Czech translations of English texts),
-and a community of volunteers corrects these. But the specifics of the projects
-bring different challenges and dictate different approaches. Finally, Marge
-2009\cite{5494979} investigates using The Mechanical Turk to obtain audio
-transcriptions.
 
@@
 
+\subsection{Corpora}
+
+Our project is not the first involving community-driven corpus annotation. We
+can mention the Manually annotated sub-corpus\cite{ide2010manually}, where
+annotations of various kinds are gathered from volunteers, or the
+Wikicorpus\cite{reese2010wikicorpus}, a corpus of Wikipedia articles with some
+linguistic annotation. Our project may reach profound similarities with these in
+the future, when we no longer focus on the transcription itself but rather on
+annotation. There is also CzEng\cite{bojar2008czeng}, the Czech-English Parallel
+Corpus, where a large part of the translation is provided by volunteers. The
+similarity in setting is considerable as both projects involve a
+machine-produced erroneous derivative of the original material (in our case
+audio transcriptions, in the case of CzEng Czech translations of English texts),
+and a community of volunteers correct these. But the specifics of the projects
+bring different challenges and dictate different approaches. Marge
+(2009)\cite{5494979} investigates using The Mechanical Turk to obtain audio
+transcriptions. Mihalcea (2004)\cite{mihalcea2004building} offers a web
+interface for word-sense disambiguation but focuses mostly on annotator conflict
+resolution.
+
 \section{Description of the Web Application}
 
-We have pointed out the differences in our setting to others. Now we shall
-describe how we implemented the program.
+\subsection{Usage}
+
+We have no special assumption of the user beyond basic computer usage skills and
+understanding the audio. We assume no prior training. There is a manual for
+clearing common points of confusion. The main message in it is that anything
+that is to be transcribed, should be should be transcribed with respect to
+phonetic precision, even if it results in nonsensical character strings.
+
+Anything except words spoken by the one speaker of interest is to be left
+untranscribed, including noise or speech by other persons\footnote{We may later
+add a speaker annotation}. Incomprehensible words are to be left uncorrected
+(the ASR output kept) if the phones are unclear. If the phones uttered are clear
+but it is not clear what word was meant, the word may be transcribed
+phonetically.
+
+A submitted transcribed segment always overwrites the present version but we
+keep all the submissions in a database, so undo operations, clustering
+submissions by their author etc. are possible but we had little need for this so
+far.
+
+\subsection{Implementation}
 
@@
 transcription with the currently played word highlighted, commit corrections to
-the transcription, and inspect a word as well as a span of text and audio. The
-corrections are checked by a forced-alignment mechanism on the server side. Our
-solution overcomes performance challenges and is a serious improvement from the
-original version. More widespread usage as well as employing it on another data
-set is to be done.
+the transcription, and inspect a word. The corrections are checked by a
+forced-alignment mechanism on the server side. Our solution overcomes
+performance challenges and is a serious improvement from the original version.
+All our codebase is open-source, accessible on
+Github\footnote{https://github.com/sixtease/MakonReact} and we are actively
+looking for similar datasets with communities to employ the application on.
 

### 2018-05-22 2018-ftc-vancouver/webapp.tex
fine tuning for FTC 2018

@@
 
+The application we describe is a part of a larger system that deals with
+Mako\v{n}'s recordings. It consists roughly of (1) the corpus itself, (2) an ASR
+system trained specially for it and (3) a web interface for the users. These
+three parts form a whole where the ASR gives a baseline transcription, the users
+correct it and the corrections are fed as further training data to the acoustic
+and language models. In this paper, we focus on the web interface.
+
 \subsection{Motivation}
@@
 answers to questions, personal experiences, behind-the-scenes to the books etc.
-The archive is freely accessible under the CC-BY license.
+The archive is freely
+accessible\footnote{https://lindat.mff.cuni.cz/repository/xmlui/handle/11372/LRT-1455}
+under the CC-BY license.
 
 \section{Differences to Other Settings}
-
-The application we describe is a part of a larger system that deals with
-Mako\v{n}'s recordings. It consists roughly of (1) the corpus itself, (2) an ASR
-system trained specially for it and (3) a web interface for the users. These
-three parts form a whole where the ASR gives a baseline transcription, the users
-correct it and the corrections are fed as further training data to the acoustic
-and language models. In this paper, we focus on the web interface.
 
@@
 
+With regard to the transription as presented to the user, a submitted segment of
+transcription always overwrites the present version but we keep all the
+submissions in a database, so undo operations, clustering submissions by their
+author etc. are possible but we had little need for this so far.
+
 \subsection{Corpora}
 
-Our project is not the first involving community-driven corpus annotation. We
+Our project is not the first involving community-driven care of a corpus. We
 can mention the Manually annotated sub-corpus\cite{ide2010manually}, where
@@
 the future, when we no longer focus on the transcription itself but rather on
-annotation. There is also CzEng\cite{bojar2008czeng}, the Czech-English Parallel
+annotation.
+
+There is also CzEng\cite{bojar2008czeng}, the Czech-English Parallel
 Corpus, where a large part of the translation is provided by volunteers. The
@@
 and a community of volunteers correct these. But the specifics of the projects
-bring different challenges and dictate different approaches. Marge
-(2009)\cite{5494979} investigates using The Mechanical Turk to obtain audio
-transcriptions. Mihalcea (2004)\cite{mihalcea2004building} offers a web
-interface for word-sense disambiguation but focuses mostly on annotator conflict
+bring different challenges and dictate different approaches.
+
+Marge (2009)\cite{5494979} investigates using The Mechanical Turk to obtain
+audio transcriptions. Mihalcea (2004)\cite{mihalcea2004building} offers a web
+interface for word-sense disambiguation and focuses mostly on annotator conflict
 resolution.
@@
 clearing common points of confusion. The main message in it is that anything
-that is to be transcribed, should be should be transcribed with respect to
+that is to be transcribed, should be transcribed with respect to
 phonetic precision, even if it results in nonsensical character strings.
@@
 Anything except words spoken by the one speaker of interest is to be left
-untranscribed, including noise or speech by other persons\footnote{We may later
-add a speaker annotation}. Incomprehensible words are to be left uncorrected
+untranscribed, including noise or speech by other persons\footnote{In our data,
+other speakers represent a negligible fraction but we may later add support for
+speaker annotation.}. Incomprehensible words are to be left uncorrected
 (the ASR output kept) if the phones are unclear. If the phones uttered are clear
@@
 phonetically.
-
-A submitted transcribed segment always overwrites the present version but we
-keep all the submissions in a database, so undo operations, clustering
-submissions by their author etc. are possible but we had little need for this so
-far.
 
@@
 We shall only discuss the detail view as the others are not relevant to this
-article. Figure~\ref{fig:scn1lab} shows the interface during playback.
+paper. Figure~\ref{fig:scn1lab} shows the interface during playback.
 Figure~\ref{fig:scn2lab} shows the interface while a segment is being edited.
@@
 \includegraphics[scale=0.6]{rc/radio-makon-en-2-lab.png}
-\caption{Interface in the state of editing a segment, with labels}
+\caption{Interface in the state of editing a segment}
 \label{fig:scn2lab}
@@
 
-Most common tasks have keyboard shortcuts: \texttt{ctrl+space} for play/pause
-and \texttt{ctrl+enter} for submitting a correction.
+The most common tasks have keyboard shortcuts: \texttt{ctrl+space} for
+play/pause and \texttt{ctrl+enter} for submitting a correction.
 
@@
 list, offers is setting up keyboard shortcuts for common words. We have not
-implemented this because ideally, common words should be covered by the speech
+implemented this because ideally, common words should be covered by speech
 recognition. However, it could be sensible to implement it anyway. The reason is
@@
 Apart from this, the version of the transcription to the recording is updated.
-This is because the transcriptions have a long cache time because normally, they
-do not change at all. At the page load, the versions of all transcriptions are
-loaded and used as cache busters. This enables us to use an external CDN and
+This is because the transcription files have a long cache time because normally,
+they do not change at all. At the page load, the versions of all transcriptions
+are loaded and used as cache busters. This enables us to use an external CDN and
 cache effectively.
@@
 Among the states the app can enter is simple playback, inspecting a word and
-transcribing a segment. Now, the only relevant things we actually keep in the
+transcribing a segment. The only relevant things we actually keep in the
 Redux store are:
@@
 
+We shall also deploy the technology on a different data set once we find a good
+fit.
+
 \section{Conclusion}

### 2018-05-26 2018-itat-krompachy/paper.tex
add draft for itat 2018 paper

\documentclass{itatnew}
%% !!!dolezite: ak pisete po slovensky alebo po cesky pouzite
%% \documentclass[slovensky]{itatnew}
%% \documentclass[cesky]{itatnew}

\begin{document}

\title{Phonetic Transcription by Untrained Annotators}

\author{Oldřich Krůza\inst{1}}

\institute{ÚFAL MFF UK,\\
\email{kruza@ufal.mff.cuni.cz},\\ WWW home page:
\texttt{http://sixtease.net/}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The paper presents an application for lay, untrained users to generate
high-quality phonetic transcription of speech. The application has been in use
for several years and has served to transcribe over 600 thousand word forms over
two versions of a web interface. In this paper, I present measures for
compensating the lack of expert training.
\end{abstract}

\section{Introduction}

\subsection{Our Setting}

The work presented in this paper is a part of the project that tends the spoken
corpus of Karel Mako\v{n}\cite{hajek2007cesky}. The corpus is of the single
speaker and has been recorded in amateur conditions, while the author was
speaking to his friends about a novel way to interpret the teaching of Jesus and
of mystic and spirituality in general. Karel Makoň died in 1993 and a community
of favorers of his teachings has persevered since then.

There are over 1000 hours of digitized recordings of Karel Makoň, they are
accessible under the CC-BY license and the project aims at bringing the most
benefit from them. The first step was digitizing the recordings from the
original magnetic tapes, the second step was releasing all of them on the
world-wide web, the third step was developing a web-based system for human /
machine transcription of the bulk, allowing for search.

The transcription we do is both phonetic and ortographic\footnote{There is no
actual focus on ortography. Instead, we mean the natural way of transcribing the
speech to human-readable text. Where it matters, focus is directed at precise
correspondence with the utterances instead of language cleanliness}. Our users
are supposed provide ortographic transcription where the pronunciation is
standard and phonetic otherwise.

\section{Annotator Expertise}

Our case is on the edge of what can be called linguistic data annotation. In our
lucky part of the world where alphabetization nears 100\%, transcription of
speech is hardly expert work. On the other hand, ensuring that the transcription
exactly matches the audio
\begin{itemize}
\item{as a representation of the words uttered and their meaning,}
\item{on the phonetic level, phoneme for phoneme,}
\item{on the time axis}
\end{itemize}
is beyond what can be expected from an untrained user.

Linguistic data annotation in general requires trained personnel. If we only
look at the Prague Dependency Treebank, we can notice the annotators provided
such a degree of expertise they have become the
co-authors\cite{hajivc2005complex}.

Crowdsourcing, community-driven approach or engaging volunteers is an ever
stronger, popular way of obtaining assets that would otherwise be unbearably
costly. Let us mention for example Mihalcea (2004)\cite{mihalcea2004building}
who delegates word-sense disambiguation to volunteers. The
Wikicorpus\cite{reese2010wikicorpus} as well as the MASC\cite{ide2010manually}
gather annotation from volunteers.

In most cases, quality is very important for data annotation, so some kind of
control is essential, no matter how expert the annotators. Trivially, the less
expertise, the more control is needed.

\subsection{Quality Control}

A common way of dealing with quality control is to inspect annotator agreement.
This has the huge downside that every piece of data must be annotated at least
twice, which reduces the yield by 50+\%.

There is another reason not to use it in our case. Our application is
designed for people who want to listen to the recordings out of interest and
their contribution to the quality of the transcription is more of a by-product.
It would be hard to convince them to choose exactly a recording that another
user has already transcribed.

Luckily, we can implement automatic measures to aid the annotators to deliver
higher-quality transcriptions.

\subsection{Forced Alignment}

Since we always assume an existing transcription, we can see the user's
contribution as a correction. Every submission has
the form of replacing a text segment with another. Since the transcriptions are
time-aligned to the audio, we also know exactly what is the corresponding audio
segment to the text submitted.

This enables us to perform forced alignment on the submitted text and the audio.
With a well selected pruning threshold, we can distinguish false transcriptions
and reject them, providing feedback to the contibutor.
Since every segment of audio fits the acoustic model to a different degree, both
false positives and false negatives will inevitably occur.

False positives (when the system accepts a wrong transcription) present a
problem, since the error will enter the data set. But false negatives can often
be circumvented by submitting the transcription divided in different segments.
Of course, this method can also be used to force a wrong transcription but we
assume no malevolence on the part of the users.

Apart from catching wrong transcription, the forced alignment mechanism provides
exact synchronization on the time axis. This is a completely missing element in
the case of virtually all programs for aiding manual transcription. For example

\subsection{Autonomous Systems}
%
In this section, we will consider the case when the Hamiltonian $H(x)$
is autonomous. For the sake of simplicity, we shall also assume that it
is $C^{1}$.

We shall first consider the question of nontriviality, within the
general framework of
$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
the second subsection, we shall look into the special case when $H$ is
$\left(0,b_{\infty}\right)$-subquadratic,
and we shall try to derive additional information.
%
\subsubsection{The General Case: Nontriviality.}
%
We assume that $H$ is
$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
with $B_{\infty}-A_{\infty}$ positive definite. Set:
\begin{eqnarray}
\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
  J \frac{d}{dt} +A_{\infty}\ .
\end{eqnarray}

Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
boundary-value problem:
\begin{equation}
\begin{array}{rcl}
  \dot{x}&=&JH' (x)\\
  x(0)&=&x (T)
\end{array}
\end{equation}
has at least one solution
$\overline{x}$, which is found by minimizing the dual
action functional:
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
with finite codimension. Here
\begin{equation}
  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
\end{equation}
is a convex function, and
\begin{equation}
  N(x) \le \frac{1}{2}
  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
  + c\ \ \ \forall x\ .
\end{equation}

%
\begin{theorem}
Assume $H'(0)=0$ and $ H(0)=0$. Set:
\begin{equation}
  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
  \label{eq:one}
\end{equation}

If $\gamma < - \lambda < \delta$,
the solution $\overline{u}$ is non-zero:
\begin{equation}
  \overline{x} (t) \ne 0\ \ \ \forall t\ .
\end{equation}
\end{theorem}
%
\begin{proof}
Condition (\ref{eq:one}) means that, for every
$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
\begin{equation}
  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
\end{equation}

It is an exercise in convex analysis, into which we shall not go, to
show that this implies that there is an $\eta > 0$ such that
\begin{equation}
  f\left\|x\right\| \le \eta
  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
  \left\|y\right\|^{2}\ .
  \label{eq:two}
\end{equation}

\begin{figure}
\vspace{2.5cm}
\caption{This is the caption of the figure displaying a white eagle and
a white horse on a snow field}
\end{figure}

Since $u_{1}$ is a smooth function, we will have
$\left\|hu_{1}\right\|_\infty \le \eta$
for $h$ small enough, and inequality (\ref{eq:two}) will hold,
yielding thereby:
\begin{equation}
  \psi (hu_{1}) \le \frac{h^{2}}{2}
  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
\end{equation}

If we choose $\delta '$ close enough to $\delta$, the quantity
$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
will be negative, and we end up with
\begin{equation}
  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
\end{equation}

On the other hand, we check directly that $\psi (0) = 0$. This shows
that 0 cannot be a minimizer of $\psi$, not even a local one.
So $\overline{u} \ne 0$ and
$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
\end{proof}
%
\begin{corollary}
Assume $H$ is $C^{2}$ and
$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
equilibria, that is, the solutions of $H' (\xi ) = 0$.
Denote by $\omega_{k}$
the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
If:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{eq:three}
\end{equation}
then minimization of $\psi$ yields a non-constant $T$-periodic solution
$\overline{x}$.
\end{corollary}
%

We recall once more that by the integer part $E [\alpha ]$ of
$\alpha \in \bbbr$, we mean the $a\in \bbbz$
such that $a< \alpha \le a+1$. For instance,
if we take $a_{\infty} = 0$, Corollary 2 tells
us that $\overline{x}$ exists and is
non-constant provided that:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
or
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{eq:four}
\end{equation}

%
\begin{proof}
The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
largest negative eigenvalue $\lambda$ is given by
$\frac{2\pi}{T}k_{o} +a_{\infty}$,
where
\begin{equation}
  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
\end{equation}
Hence:
\begin{equation}
  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
\end{equation}

The condition $\gamma < -\lambda < \delta$ now becomes:
\begin{equation}
  b_{\infty} - a_{\infty} <
  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
\end{equation}
which is precisely condition (\ref{eq:three}).\qed
\end{proof}
%

\begin{figure*}
\vspace{2.5cm}
\caption{This is the caption of the figure displaying a white eagle and
a white horse on a snow field}
\end{figure*}

\begin{lemma}
Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
\end{lemma}
%
\begin{proof}
We know that $\widetilde{x}$, or
$\widetilde{x} + \xi$ for some constant $\xi
\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
\begin{equation}
  \dot{x} = JH' (x)\ .
\end{equation}

There is no loss of generality in taking $\xi = 0$. So
$\psi (x) \ge \psi (\widetilde{x} )$
for all $\widetilde{x}$ in some neighbourhood of $x$ in
$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.

But this index is precisely the index
$i_{T} (\widetilde{x} )$ of the $T$-periodic
solution $\widetilde{x}$ over the interval
$(0,T)$, as defined in Sect.~2.6. So
\begin{equation}
  i_{T} (\widetilde{x} ) = 0\ .
  \label{eq:five}
\end{equation}

Now if $\widetilde{x}$ has a lower period, $T/k$ say,
we would have, by Corollary 31:
\begin{equation}
  i_{T} (\widetilde{x} ) =
  i_{kT/k}(\widetilde{x} ) \ge
  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
\end{equation}

This would contradict (\ref{eq:five}), and thus cannot happen.\qed
\end{proof}
%
\paragraph{Notes and Comments.}
The results in this section are a
refined version of \cite{clar:eke};
the minimality result of Proposition
14 was the first of its kind.

To understand the nontriviality conditions, such as the one in formula
(\ref{eq:four}), one may think of a one-parameter family
$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
of periodic solutions, $x_{T} (0) = x_{T} (T)$,
with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
which is the period of the linearized system at 0.

\begin{table}
\caption{This is the example table taken out of {\it The
\TeX{}book,} p.\,246}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
Assume $H(t,x)$ is
$(0,\varepsilon )$-subquadratic at
infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
\begin{equation}
  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
\end{equation}
\begin{equation}
  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
\end{equation}
\begin{equation}
  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
\end{equation}
\begin{equation}
  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
\end{equation}

Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
$kT$-periodic solutions of the system
\begin{equation}
  \dot{x} = JH' (t,x)
\end{equation}
such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
\begin{equation}
  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
\end{equation}
\qed
\end{theorem}
%
\begin{example} [{{\rm External forcing}}]
Consider the system:
\begin{equation}
  \dot{x} = JH' (x) + f(t)
\end{equation}
where the Hamiltonian $H$ is
$\left(0,b_{\infty}\right)$-subquadratic, and the
forcing term is a distribution on the circle:
\begin{equation}
  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
\end{equation}
where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
\begin{equation}
  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
\end{equation}
where $\delta_{k}$ is the Dirac mass at $t= k$ and
$\xi \in \bbbr^{2n}$ is a
constant, fits the prescription. This means that the system
$\dot{x} = JH' (x)$ is being excited by a
series of identical shocks at interval $T$.
\end{example}
%
\begin{definition}
Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
operators in $\bbbr^{2n}$, depending continuously on
$t\in [0,T]$, such that
$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.

A Borelian function
$H: [0,T]\times \bbbr^{2n} \to \bbbr$
is called
$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
if there exists a function $N(t,x)$ such that:
\begin{equation}
  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
\end{equation}
\begin{equation}
  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
  {\rm is\ convex\ with\  respect\  to}\ \ x
\end{equation}
\begin{equation}
  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
\end{equation}
\begin{equation}
  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
\end{equation}

If $A_{\infty} (t) = a_{\infty} I$ and
$B_{\infty} (t) = b_{\infty} I$, with
$a_{\infty} \le b_{\infty} \in \bbbr$,
we shall say that $H$ is
$\left(a_{\infty},b_{\infty}\right)$-subquadratic
at infinity. As an example, the function
$\left\|x\right\|^{\alpha}$, with
$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
for every $\varepsilon > 0$. Similarly, the Hamiltonian
\begin{equation}
H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
\end{equation}
is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
Note that, if $k<0$, it is not convex.
\end{definition}
%

\paragraph{Notes and Comments.}
The first results on subharmonics were
obtained by Rabinowitz in \cite{rab}, who showed the existence of
infinitely many subharmonics both in the subquadratic and superquadratic
case, with suitable growth conditions on $H'$. Again the duality
approach enabled Clarke and Ekeland in \cite{clar:eke:2} to treat the
same problem in the convex-subquadratic case, with growth conditions on
$H$ only.

Recently, Michalek and Tarantello (see \cite{mich:tar} and \cite{tar})
have obtained lower bound on the number of subharmonics of period $kT$,
based on symmetry considerations and on pinching estimates, as in
Sect.~5.2 of this article.

%
% ---- Bibliography ----
%
\begin{thebibliography}{5}
%
\bibitem {clar:eke}
Clarke, F., Ekeland, I.:
Nonlinear oscillations and
boundary-value problems for Hamiltonian systems.
Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333

\bibitem {clar:eke:2}
Clarke, F., Ekeland, I.:
Solutions p\'{e}riodiques, du
p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
Note CRAS Paris {\bf 287} (1978) 1013--1015

\bibitem {mich:tar}
Michalek, R., Tarantello, G.:
Subharmonic solutions with prescribed minimal
period for nonautonomous Hamiltonian systems.
J. Diff. Eq. {\bf 72} (1988) 28--55

\bibitem {tar}
Tarantello, G.:
Subharmonic solutions for Hamiltonian
systems via a $\bbbz_{p}$ pseudoindex theory.
Annali di Matematica Pura (to appear)

\bibitem {rab}
Rabinowitz, P.:
On subharmonic solutions of a Hamiltonian system.
Comm. Pure Appl. Math. {\bf 33} (1980) 609--633

\end{thebibliography}
\end{document}

### 2018-05-26 2018-itat-krompachy/paper.tex
itat2018: remove sample paper text

@@
 exact synchronization on the time axis. This is a completely missing element in
-the case of virtually all programs for aiding manual transcription. For example
+the case of virtually all programs for aiding manual transcription. For some
+examples, Transcriber\footnote{http://trans.sourceforge.net/}, a veteran
+open-source transcribing program for Linux, expects the user to provide
+alignment on the level of phrases;
+Transcribe\footnote{http://transcribe.wreally.com/}, a commercial web-based
+transcribing tool, allows the user to add timestamps anywhere in the text. There
+is no acoustic model, hence nothing to match against.
 
-\subsection{Autonomous Systems}
-%
-In this section, we will consider the case when the Hamiltonian $H(x)$
-is autonomous. For the sake of simplicity, we shall also assume that it
-is $C^{1}$.
-
-We shall first consider the question of nontriviality, within the
-general framework of
-$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
-the second subsection, we shall look into the special case when $H$ is
-$\left(0,b_{\infty}\right)$-subquadratic,
-and we shall try to derive additional information.
-%
-\subsubsection{The General Case: Nontriviality.}
-%
-We assume that $H$ is
-$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
-for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
-with $B_{\infty}-A_{\infty}$ positive definite. Set:
-\begin{eqnarray}
-\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
-  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
-  J \frac{d}{dt} +A_{\infty}\ .
-\end{eqnarray}
-
-Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
-boundary-value problem:
-\begin{equation}
-\begin{array}{rcl}
-  \dot{x}&=&JH' (x)\\
-  x(0)&=&x (T)
-\end{array}
-\end{equation}
-has at least one solution
-$\overline{x}$, which is found by minimizing the dual
-action functional:
-\begin{equation}
-  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
-  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
-\end{equation}
-on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
-with finite codimension. Here
-\begin{equation}
-  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
-\end{equation}
-is a convex function, and
-\begin{equation}
-  N(x) \le \frac{1}{2}
-  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
-  + c\ \ \ \forall x\ .
-\end{equation}
-
-%
-\begin{theorem}
-Assume $H'(0)=0$ and $ H(0)=0$. Set:
-\begin{equation}
-  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
-  \label{eq:one}
-\end{equation}
-
-If $\gamma < - \lambda < \delta$,
-the solution $\overline{u}$ is non-zero:
-\begin{equation}
-  \overline{x} (t) \ne 0\ \ \ \forall t\ .
-\end{equation}
-\end{theorem}
-%
-\begin{proof}
-Condition (\ref{eq:one}) means that, for every
-$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
-\begin{equation}
-  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
-  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
-\end{equation}
-
-It is an exercise in convex analysis, into which we shall not go, to
-show that this implies that there is an $\eta > 0$ such that
-\begin{equation}
-  f\left\|x\right\| \le \eta
-  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
-  \left\|y\right\|^{2}\ .
-  \label{eq:two}
-\end{equation}
-
-\begin{figure}
-\vspace{2.5cm}
-\caption{This is the caption of the figure displaying a white eagle and
-a white horse on a snow field}
-\end{figure}
-
-Since $u_{1}$ is a smooth function, we will have
-$\left\|hu_{1}\right\|_\infty \le \eta$
-for $h$ small enough, and inequality (\ref{eq:two}) will hold,
-yielding thereby:
-\begin{equation}
-  \psi (hu_{1}) \le \frac{h^{2}}{2}
-  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
-  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
-\end{equation}
-
-If we choose $\delta '$ close enough to $\delta$, the quantity
-$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
-will be negative, and we end up with
-\begin{equation}
-  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
-\end{equation}
-
-On the other hand, we check directly that $\psi (0) = 0$. This shows
-that 0 cannot be a minimizer of $\psi$, not even a local one.
-So $\overline{u} \ne 0$ and
-$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
-\end{proof}
-%
-\begin{corollary}
-Assume $H$ is $C^{2}$ and
-$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
-$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
-equilibria, that is, the solutions of $H' (\xi ) = 0$.
-Denote by $\omega_{k}$
-the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
-\begin{equation}
-  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
-\end{equation}
-If:
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} <
-  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
-  \frac{T}{2\pi}\omega
-  \label{eq:three}
-\end{equation}
-then minimization of $\psi$ yields a non-constant $T$-periodic solution
-$\overline{x}$.
-\end{corollary}
-%
-
-We recall once more that by the integer part $E [\alpha ]$ of
-$\alpha \in \bbbr$, we mean the $a\in \bbbz$
-such that $a< \alpha \le a+1$. For instance,
-if we take $a_{\infty} = 0$, Corollary 2 tells
-us that $\overline{x}$ exists and is
-non-constant provided that:
-
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
-\end{equation}
-or
-\begin{equation}
-  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
-  \label{eq:four}
-\end{equation}
-
-%
-\begin{proof}
-The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
-largest negative eigenvalue $\lambda$ is given by
-$\frac{2\pi}{T}k_{o} +a_{\infty}$,
-where
-\begin{equation}
-  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
-  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
-\end{equation}
-Hence:
-\begin{equation}
-  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
-\end{equation}
-
-The condition $\gamma < -\lambda < \delta$ now becomes:
-\begin{equation}
-  b_{\infty} - a_{\infty} <
-  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
-\end{equation}
-which is precisely condition (\ref{eq:three}).\qed
-\end{proof}
-%
-
-\begin{figure*}
-\vspace{2.5cm}
-\caption{This is the caption of the figure displaying a white eagle and
-a white horse on a snow field}
-\end{figure*}
-
-\begin{lemma}
-Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
-that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
-minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
-\end{lemma}
-%
-\begin{proof}
-We know that $\widetilde{x}$, or
-$\widetilde{x} + \xi$ for some constant $\xi
-\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
-\begin{equation}
-  \dot{x} = JH' (x)\ .
-\end{equation}
-
-There is no loss of generality in taking $\xi = 0$. So
-$\psi (x) \ge \psi (\widetilde{x} )$
-for all $\widetilde{x}$ in some neighbourhood of $x$ in
-$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.
-
-But this index is precisely the index
-$i_{T} (\widetilde{x} )$ of the $T$-periodic
-solution $\widetilde{x}$ over the interval
-$(0,T)$, as defined in Sect.~2.6. So
-\begin{equation}
-  i_{T} (\widetilde{x} ) = 0\ .
-  \label{eq:five}
-\end{equation}
-
-Now if $\widetilde{x}$ has a lower period, $T/k$ say,
-we would have, by Corollary 31:
-\begin{equation}
-  i_{T} (\widetilde{x} ) =
-  i_{kT/k}(\widetilde{x} ) \ge
-  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
-\end{equation}
-
-This would contradict (\ref{eq:five}), and thus cannot happen.\qed
-\end{proof}
-%
-\paragraph{Notes and Comments.}
-The results in this section are a
-refined version of \cite{clar:eke};
-the minimality result of Proposition
-14 was the first of its kind.
-
-To understand the nontriviality conditions, such as the one in formula
-(\ref{eq:four}), one may think of a one-parameter family
-$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
-of periodic solutions, $x_{T} (0) = x_{T} (T)$,
-with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
-which is the period of the linearized system at 0.
-
-\begin{table}
-\caption{This is the example table taken out of {\it The
-\TeX{}book,} p.\,246}
-\begin{center}
-\begin{tabular}{r@{\quad}rl}
-\hline
-\multicolumn{1}{l}{\rule{0pt}{12pt}
-                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
-\hline\rule{0pt}{12pt}
-8000 B.C.  &     5,000,000& \\
-  50 A.D.  &   200,000,000& \\
-1650 A.D.  &   500,000,000& \\
-1945 A.D.  & 2,300,000,000& \\
-1980 A.D.  & 4,400,000,000& \\[2pt]
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
-Assume $H(t,x)$ is
-$(0,\varepsilon )$-subquadratic at
-infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
-\begin{equation}
-  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
-\end{equation}
-\begin{equation}
-  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
-\end{equation}
-\begin{equation}
-  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
-  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
-\end{equation}
-\begin{equation}
-  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
-  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
-\end{equation}
-
-Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
-everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
-$kT$-periodic solutions of the system
-\begin{equation}
-  \dot{x} = JH' (t,x)
-\end{equation}
-such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
-\begin{equation}
-  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
-\end{equation}
-\qed
-\end{theorem}
-%
-\begin{example} [{{\rm External forcing}}]
-Consider the system:
-\begin{equation}
-  \dot{x} = JH' (x) + f(t)
-\end{equation}
-where the Hamiltonian $H$ is
-$\left(0,b_{\infty}\right)$-subquadratic, and the
-forcing term is a distribution on the circle:
-\begin{equation}
-  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
-  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
-\end{equation}
-where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
-\begin{equation}
-  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
-\end{equation}
-where $\delta_{k}$ is the Dirac mass at $t= k$ and
-$\xi \in \bbbr^{2n}$ is a
-constant, fits the prescription. This means that the system
-$\dot{x} = JH' (x)$ is being excited by a
-series of identical shocks at interval $T$.
-\end{example}
-%
-\begin{definition}
-Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
-operators in $\bbbr^{2n}$, depending continuously on
-$t\in [0,T]$, such that
-$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.
-
-A Borelian function
-$H: [0,T]\times \bbbr^{2n} \to \bbbr$
-is called
-$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
-if there exists a function $N(t,x)$ such that:
-\begin{equation}
-  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
-\end{equation}
-\begin{equation}
-  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
-  {\rm is\ convex\ with\  respect\  to}\ \ x
-\end{equation}
-\begin{equation}
-  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
-  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
-\end{equation}
-\begin{equation}
-  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
-  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
-\end{equation}
-
-If $A_{\infty} (t) = a_{\infty} I$ and
-$B_{\infty} (t) = b_{\infty} I$, with
-$a_{\infty} \le b_{\infty} \in \bbbr$,
-we shall say that $H$ is
-$\left(a_{\infty},b_{\infty}\right)$-subquadratic
-at infinity. As an example, the function
-$\left\|x\right\|^{\alpha}$, with
-$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
-for every $\varepsilon > 0$. Similarly, the Hamiltonian
-\begin{equation}
-H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
-\end{equation}
-is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
-Note that, if $k<0$, it is not convex.
-\end{definition}
-%
-
-\paragraph{Notes and Comments.}
-The first results on subharmonics were
-obtained by Rabinowitz in \cite{rab}, who showed the existence of
-infinitely many subharmonics both in the subquadratic and superquadratic
-case, with suitable growth conditions on $H'$. Again the duality
-approach enabled Clarke and Ekeland in \cite{clar:eke:2} to treat the
-same problem in the convex-subquadratic case, with growth conditions on
-$H$ only.
-
-Recently, Michalek and Tarantello (see \cite{mich:tar} and \cite{tar})
-have obtained lower bound on the number of subharmonics of period $kT$,
-based on symmetry considerations and on pinching estimates, as in
-Sect.~5.2 of this article.
-
-%
-% ---- Bibliography ----
-%
-\begin{thebibliography}{5}
-%
-\bibitem {clar:eke}
-Clarke, F., Ekeland, I.:
-Nonlinear oscillations and
-boundary-value problems for Hamiltonian systems.
-Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
-
-\bibitem {clar:eke:2}
-Clarke, F., Ekeland, I.:
-Solutions p\'{e}riodiques, du
-p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
-Note CRAS Paris {\bf 287} (1978) 1013--1015
-
-\bibitem {mich:tar}
-Michalek, R., Tarantello, G.:
-Subharmonic solutions with prescribed minimal
-period for nonautonomous Hamiltonian systems.
-J. Diff. Eq. {\bf 72} (1988) 28--55
-
-\bibitem {tar}
-Tarantello, G.:
-Subharmonic solutions for Hamiltonian
-systems via a $\bbbz_{p}$ pseudoindex theory.
-Annali di Matematica Pura (to appear)
-
-\bibitem {rab}
-Rabinowitz, P.:
-On subharmonic solutions of a Hamiltonian system.
-Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
-
-\end{thebibliography}
 \end{document}

### 2018-06-02 2018-itat-krompachy/paper.tex
wip: ITAT 2018

@@
 The paper presents an application for lay, untrained users to generate
-high-quality phonetic transcription of speech. The application has been in use
+high-quality, aligned phonetic transcription of speech. The application has been in use
 for several years and has served to transcribe over 600 thousand word forms over
-two versions of a web interface. In this paper, I present measures for
-compensating the lack of expert training.
+two versions of a web interface. We present measures for compensating the lack of expert training.
 \end{abstract}
@@
 
-The transcription we do is both phonetic and ortographic\footnote{There is no
-actual focus on ortography. Instead, we mean the natural way of transcribing the
+The transcription we do is both phonetic and orthographic\footnote{There is no
+actual focus on orthography. Instead, we mean the natural way of transcribing the
 speech to human-readable text. Where it matters, focus is directed at precise
 correspondence with the utterances instead of language cleanliness}. Our users
-are supposed provide ortographic transcription where the pronunciation is
+are supposed provide orthographic transcription where the pronunciation is
 standard and phonetic otherwise.
@@
 exact synchronization on the time axis. This is a completely missing element in
-the case of virtually all programs for aiding manual transcription. For some
+the case of virtually all programs for computer-aided transcription. For some
 examples, Transcriber\footnote{http://trans.sourceforge.net/}, a veteran
@@
 alignment on the level of phrases;
-Transcribe\footnote{http://transcribe.wreally.com/}, a commercial web-based
+Transcribe\footnote{https://transcribe.wreally.com/}, a commercial web-based
 transcribing tool, allows the user to add timestamps anywhere in the text. There
@@
 
+\section{Phonetic Transcription}
+
+\subsection{Purpose}
+
+We have originally built the acoustic model using
+HTK\footnote{http://htk.eng.cam.ac.uk/}, the Hidden Markov Toolkit. Here,
+explicit phonetically labeled training data are necessary for training. We are
+switching to DNN, using Mozilla's
+DeepSpeech\footnote{https://github.com/mozilla/DeepSpeech}, where no explicit
+phonetic annotation is needed but for some purposes like forced alignment, the
+original HMM is still irreplaceable.
+
+Also, the phonetic labeling is valuable per se for research purposes.
+
+\subsection{Acquisition}
+
+The phonetic transcription is in normal case also a product of forced alignment,
+as in case of pronunciation variants, it selects the most fitting one. This
+requires a way to automatically obtain all pronunciation variants of any word.
+We use a combination of a rule-based system inspired by Psutka et
+al.\cite{psutka2004development}, in combination with a dynamic dictionary.
+The dynamic dictionary is a list of alternative pronunciations of a word, which
+expands as the app is being used.
+
+The users are instructed to transcribe any words with non-standard pronunciation
+phonetically and then correct their orthographical form. This is one of the few
+cases where we are coercing the users to something.
+
+When the orthographically broken, phonetic transcription of a word is submitted,
+if it passes the forced-alignment phase, it is integrated into the displayed
+transcription. The word's data representation consists of its
+\begin{enumerate}
+\item{occurrence:
+    the word as it appears in the text, including capitalization and
+    punctuation,
+}
+\item{wordform:
+    the word as it appears in the language model and phonetic dictionary
+    (computed as the occurrence in lowercase and stripped of non-alphabetic
+    characters\footnote{This implies that all non-alphabetic characters are
+    always a part of a token and never form a token on their own.}),
+}
+\item{pronunciation:
+    an array of phonemes,
+}
+\item{timestamp:
+    distance of the beginning of the word from the beginning of the file, in
+    seconds, in precision of 2 decimal digits,
+}
+\item{manual/automatic:
+    boolean flag denoting whether the word has been transcribed manually or not,
+}
+\item{confidence measure:
+    in case of automatically acquired words, the confidence-measure score of the
+    recognizer.
+}
+\end{enumerate}
+Once merged into the displayed transcription, each word's occurrence can be
+edited manually. Hence the user can enter the correct form deviating from Czech
+pronunciation rules.
+
+Doing so results in adding the wordform-pronunciation couple to the dynamic
+pronunciation dictionary and is also used at forced alignment. Thus, this
+operation must only be performed once per word and any subsequent time the word
+is entered in its standard orthographic form, the correct pronunciation is
+inferred.
+
+For example, let's examine the scenario of transcribing the sentence {\em Proč
+se toto nestalo Marii Markétě Alacoque?} Its phonetic representation is {\em p r
+o ch sil s e sil t o t o sil n e s t a l o sil m a r i j i sil m a r k ee tj e
+sil a l a k o k}.
+\begin{enumerate}
+\item{Suppose the user enters the correct ortographic transcription.}
+\item{
+    The phonetic transducer outputs {\em
+        p r o ch sil
+        s e sil
+        t o t o sil
+        n e s t a l o sil
+        m a r i j i sil
+        m a r k ee tj e sil
+        a l a c o k v u e
+    }.
+}
+\item{
+    With a bit of luck, the forced alignment fails because of the distiction of
+    the phone sequence {\em k o k} and {\em c o k v u e}.
+}
+\item{
+    The transcription is rejected, the user realizes that the word is pronounced
+    in a non-standard way and re-tries with
+    {\em Proč se toto nestalo Marii Markétě alakok?}
+}
+\item{
+    Forced alignment succeeds now and the entered transcription is merged into
+    the view.
+}
+\item{
+    The user selects the non-existent word {\em alakok} and edits its occurrence
+    to {\em Alacoque?}
+}
+\item{
+    Now the word is correctly stored and on any subsequent user inputs of {\em
+    Alacoque} with any punctuation or capitalization, the pronunciation
+    {\em a l a k o k} is considered at the forced alignment.
+}
+\end{enumerate}
+
+\subsection{Representation}
+
+As can be seen above, for the sake of simplicity and interoperability, we use
+the set of phonemes as defined by Nouza et al.\cite{nouza1997phonetic} With all
+advantages of this representation, it is clearly not the most natural way for
+lay Czechs to write down literal pronunciation. Thanks to the simple, mostly
+deterministic mapping between phonemes and graphemes, pronunciation respelling
+is a reliable, natural way. There's not even a need for explicit syllable
+separation as seen in English pronunciation respelling {\em
+(wikipedia\footnote{https://en.wikipedia.org/wiki/Pronunciation\_respelling}
+gives the example {\em ``Diarrhoea'' is pronounced DYE-uh-REE-a})}.
+
+The previous subsection gave an example of using pronunciation respelling in
+Czech with the example of {\em alakok} for {\em Alacoque}. The direction from
+the phonetic respelling to the phoneme array is covered by the
+ortographic-to-phonetic transducer. But we also need the opposite direction to
+provide the users a way to check whether the pronunciation selected by the
+forced alignment fits.
+
+For this purpose, we have created a JavaScript module for transduction between
+the array of phonemes and the pronunciation
+respelling\footnote{https://github.com/Sixtease/MakonReact/blob/master/src/lib/Phonet.js}.
+
+The module includes two-way transduction, although only the one from array of
+phonemes to human-readable phonetic respelling is needed in our application.
+Note that when encoding into the phonetic respelling, none of
+{\em di ti ni dě tě ně} is ever output. The palatal consonants are always
+explicitly spelled out.
+
+The ambiguity of diphtong / subsequent vocals, as in {\em poučit [p o u ch i t]}
+vs. {\em louže [l ow zh e]} is explicated by an apostrophe in the respelling
+{\em po'učit / louže}. Voiceless alveolar fricative trill is explicated as {\em
+r'} as in {\em tř'ít} vs. {\em dřít}.
+
 \end{document}

### 2018-06-03 2018-itat-krompachy/paper.tex
wip: itat 2018 text

@@
 \institute{ÚFAL MFF UK,\\
-\email{kruza@ufal.mff.cuni.cz},\\ WWW home page:
-\texttt{http://sixtease.net/}}
+\email{kruza@ufal.mff.cuni.cz}}
 
@@
 
-\subsection{Representation}
+\subsection{Phonetic Respelling}
+\label{subsec:respelling}
 
@@
 
+The algorithm is simple. In most cases, a phoneme corresponds uniquely to one
+character in the respelling. Exceptions are as follows:
+\begin{enumerate}
+\item{The phone {\em (x)} is spelled {\em ``ch''}.}
+\item{The phone {\em (dz) (dzh)} are spelled {\em ``dz'' ``dž''}.}
+\item{The diphtongs {\em (aw) (ew) (ow)} are spelled {\em ``au'' ``eu'' ``ou''}.}
+\item{
+    Sequences {\em (c) (h), (o) (u), (a) (u), (e) (u), (d) (z), (d zh) } are
+    spelled {\em ``c'h'', ``o'u'', ``a'u'', ``e'u'', ``d'z'', ``d'ž''}. Note
+    though, that the sequence {\em (c) (h)} is purely hypothetical, as it
+    contradicts voiced/voiceless assimilation.
+}
+\item{
+    Voiceless alveolar fricative trill is explicated as {\em ``r'{}''} as in {\em
+    tř'ít} vs. {\em dřít}.
+}
+\item{
+    Palatal nasal and labiodental nasal are spelled {\em ``n'{}'', ``m'{}''}.
+}
+\item{Trailing silence is not represented.}
+\end{enumerate}
+
 The module includes two-way transduction, although only the one from array of
 phonemes to human-readable phonetic respelling is needed in our application.
+Still, the user can mark up special-case pronunciation with the apostrophe, like
+the sequence of phonemes {\em (c)} and {\em (h)} with the string {\em c'h}. The
+need has never occurred during the six years' lifespan of the application.
+
 Note that when encoding into the phonetic respelling, none of
 {\em di ti ni dě tě ně} is ever output. The palatal consonants are always
-explicitly spelled out.
+explicitly spelled out and e.g. the sequence {\em (n) (i)} is always spelled
+{\em``ny''}
 
-The ambiguity of diphtong / subsequent vocals, as in {\em poučit [p o u ch i t]}
-vs. {\em louže [l ow zh e]} is explicated by an apostrophe in the respelling
-{\em po'učit / louže}. Voiceless alveolar fricative trill is explicated as {\em
-r'} as in {\em tř'ít} vs. {\em dřít}.
+A few examples of words and their phonetic respelling as output by the
+algorithm (given the corresponding pronunciation is input as phoneme list):
+\begin{itemize}
+\item{nic: {\em ňic},}
+\item{kdo: {\em gdo},}
+\item{disk: {\em dysk},}
+\item{dřít: {\em dřít},}
+\item{třít: {\em tř'ít},}
+\item{auto: {\em auto},}
+\item{nauka: {\em na'uka},}
+\item{džbán: {\em džbán},}
+\item{odželezněný: {\em od'železněný},}
+\item{odznak: {\em odznak},}
+\item{podzemí: {\em pod'zemí},}
+\item{noc: {\em noc},}
+\item{tento: {\em tento},}
+\item{hangár: {\em han'gár},}
+\item{samba: {\em samba},}
+\item{tonfa: {\em tom'fa}.}
+\end{itemize}
+
+We postulate that the phonetic respelling is natural to all alphabetized native
+Czech speakers as a fact without any supporting research, based on experience
+alone. The use of apostrophe for distinguishing ambiguities and special cases is
+not 100\% intuitive and presents another point where instruction is necessary
+for the user to use this feature properly.
+
+\section{Evaluation}
+
+We have presented our web application as a tool that enables gathering precisely
+aligned, phoneme-exact transcription from untrained casual visitors. We have
+presented measures for reaching this goal but the degree to which it was reached
+remains unclear.
+
+We have no gold standard data to measure the quality of our manual
+transcriptions. On the contrary, we use the manual transcriptions as gold
+standard for the automatic recognition. What we can to, however, is look at some
+random samples and try to get a rough idea of how the system performs.
+
+\subsection{Validation by Forced Alignment}
+
+One thing we can examine are the approvals / rejections of the forced alignment.
+Of 109640 forced alignment attempts, 3419 have failed, which makes for 3.12\%.
+We have manually inspected 20 random failed attempts and came to the following
+numbers:
+\begin{itemize}
+\item{
+    11 cases were false negatives, where the transcription was correct and
+    should have been accepted,
+}
+\item{
+    4 cases were caused by acoustic irregularities like noise,
+}
+\item{
+    4 cases were true negatives caused by wrongly chosen segment boundaries and
+}
+\item{
+    1 case was true negative caused by wrong transcription.
+}
+\end{itemize}
+
+Hence, in 25\% of the minimalistic sample, the forced alignment did its job of a
+validator and prevented a piece of broken training data from entering the
+dataset. In 55\% it was a nuisance and failure, and in the remaining 20\%, it
+rejected a valid transcription but prevented a bad training example from
+occurring, so we can see this in positive light.
+
+\subsection{Non-Standard Pronunciation}
+
+We can also track how the scenario described in
+subsection~\ref{subsec:respelling} is applied. We have looked up four promising
+example records in the dynamic dictionary and checked submitted transcriptions
+containing them. Table~\ref{tab:pronunc}
+
+\begin{table}[htpb]
+\begin{center}
+\begin{tabular}{|l r|l r|l r|l r|}
+\hline
+Correct written form        & \#
+	& wrong phonetic pronunc.   & \#
+		& correct pronunc.          & \#
+			& phon. respelling          & \# \\
+\hline
+Moody & 2 & {\texttt m o o d i} & 0 & {\texttt m uu d i} & 4 & {\em múdy}, {\em můdy} & 2 \\
+Descartes & 2 & {\texttt d e s c a r t e s} & 0 & {\texttt d e k aa r t} & 4 & {\em dekárt} & 2   \\
+Weinfurter & 18 & {\texttt v e j n f u r t e r} & 9 & {\texttt v a j n f u r t r} & 16 & {\em vajnfurtr} & 9 \\
+Michelangelo & 6 & {\texttt m i x e l a ng g e l o} & 9 & {\texttt m i k e l a n dzh e l o} & 9 & {\em mikelan'dželo} & 4 \\
+\hline
+\end{tabular}
+\caption{Examples of non-standard pronunciation in the manually transcribed data}\label{tab:pronunc}
+\end{center}
+\end{table}
 

### 2018-06-04 2018-itat-krompachy/paper.tex
more itat 2018 text

@@
 example records in the dynamic dictionary and checked submitted transcriptions
-containing them. Table~\ref{tab:pronunc}
+containing them. Table~\ref{tab:pronunc} lists for each of them the correct
+orthographic form, the wrong pronunciation obtained by the transducer, the
+correct pronunciation and finally the phonetic respelling. Each is followed by
+the number of occurrences in the manually transcribed data.
 
@@
 Descartes & 2 & {\texttt d e s c a r t e s} & 0 & {\texttt d e k aa r t} & 4 & {\em dekárt} & 2   \\
-Weinfurter & 18 & {\texttt v e j n f u r t e r} & 9 & {\texttt v a j n f u r t r} & 16 & {\em vajnfurtr} & 9 \\
-Michelangelo & 6 & {\texttt m i x e l a ng g e l o} & 9 & {\texttt m i k e l a n dzh e l o} & 9 & {\em mikelan'dželo} & 4 \\
+Weinfurter & 30 & {\texttt v e j n f u r t e r} & 13 & {\texttt v a j n f u r t r} & 19 & {\em vajnfurtr} & 2 \\
+Michelangelo & 6 & {\texttt m i x e l a ng g e l o} & 2 & {\texttt m i k e l a n dzh e l o} & 4 & {\em mikelandželo} & 0 \\
 \hline
@@
 
+\begin{table}[htpb]
+\begin{center}
+\begin{tabular}{|l|r|r|}
+\hline
+ & phonetically correct & phonetically incorrect \\
+orthographically correct & 25 & 15 \\
+orthographically incorrect & 6 & 0
+\hline
+\end{tabular}
+\caption{Success rate for phonetic and orthographic representation of foreign
+words}\label{tab:pronunc-rate}
+\end{center}
+\end{table}
+
+We can see in Table~\ref{tab:pronunc-rate} that the majority of cases results in
+both orthographic and phonetic forms being correct. Only in about 13\% cases,
+the orthographically incorrect form is kept. We attribute this to the fact that
+those who use the phonetic transcription are aware of the problematic and mostly
+go the whole way and clean up.
+
+On the other hand, nearly a third of the cases show the wrong phonetic
+representation. This is a serious problem on at least two levels: Firstly, it
+shows that the forced aligner failed to catch the error. Secondly
+
 \end{document}

### 2018-06-04 2018-itat-krompachy/paper.tex
more itat 2018 text

@@
 of favorers of his teachings has persevered since then.
+
+The talks can be seen as companions to Makoň's written works. Together they form
+a unique, extensive, consistent systematization of the spiritual path tailored
+to modern westerners and accessible primarily to Czech speakers. It draws
+heavily on traditional Christian mysticism as well as ancient tradition of India
+and China, adapting them for the present. The whole system can be seen as a
+manual for entering the eternal life prior to the physical death.
+
+Makoň points to the life of Jesus as depicted by the gospels as to perfect,
+detailed manual for the path to eternal life usable by anybody. He corrects many
+alleged mistakes in the official Catholic doctrine, although he sees its
+tradition as a very useful and underestimated tool. I, the author of this paper,
+would personally find it most interesting to be able to compare key Makoň's
+postulates with official stance of the Church and its evolution in the last
+decades.
 
@@
 Also, the phonetic labeling is valuable per se for research purposes.
+
+\subsection{Phoneme Set}
+
+We use a subset of PACal\cite{nouza1997phonetic}. We shall also refer to
+individual phonemes in this paper using the PACal notation. For reference,
+Table~\ref{tab:phones} lists the phonemes used with their IPA notation.
+
+\begin{table}[htpb]
+\fontspec{DoulosSIL}
+\begin{center}
+\begin{tabular}{|l|l|l|l||l|l|l|l|}
+\hline
+IPA & PACal & common grapheme & IPA & PACal &  common grapheme \\
+\hline
+% TODO: IPA
+a  & a   & a      &     ɱ  & mg  & tra\underline{m}vaj \\
+aː & aa  & á      &     n  & n   & \underline{n}e \\
+aʊ̯ & aw  & au     &     ŋ  & ng  & pa\underline{n}t \\
+b  & b   & b      &     ɲ  & nj  & \v{n} \\
+t͡s & c   & c      &     o  & o   & o \\
+t͡ʃ & ch  & č      &     oː & oo  & ó \\
+d  & d   & d      &     oʊ̯ & ow  & ou \\
+ɟ  & dj  & \v{d}  &     p  & p   & p \\
+d͡z & dz  & dz     &     r  & r   & r \\
+d͡ʒ & dzh & dž     &     r̝̊  & rsh & t\underline{\v{r}}i \\
+ɛ  & e   & e      &     r̝  & rzh & \underline{\v{r}}íz \\
+ɛː & ee  & é      &     s  & s   & s \\
+eʊ̯ & ew  & eu     &     ʃ  & sh  & š \\
+f  & f   & f      &     t  & t   & t \\
+g  & g   & g      &     c  & tj  & \v{t} \\
+ɦ  & h   & h      &     ʊ  & u   & u \\
+i  & i   & i      &     uː & uu  & ú, \r{u} \\
+iː & ii  & í      &     v  & v   & v \\
+j  & j   & j      &     x  & x   & ch \\
+k  & k   & k      &     z  & z   & z \\
+l  & l   & l      &     ʒ  & zh  & ž \\
+m  & m   & \underline{m}ák
+                  &        & sil & \\
+   &     &        &        & sp  & \\
+\hline
+\end{tabular}
+\caption{Phones used in transcription}\label{tab:phones}
+\end{center}
+\end{table}
+\normalfont
 
@@
 representation. This is a serious problem on at least two levels: Firstly, it
-shows that the forced aligner failed to catch the error. Secondly
+shows that the forced aligner failed to catch the error. Secondly, it lets bad
+examples into the training dataset.
+
+One of the apparent reasons for this to happen is that the dynamic dictionary
+only recognizes exact matches. We can see in one file, for example, all
+occurrences of the form {\em Weinfurter} to have correct pronunciation while
+{\em Weinfurterovi} a broken one.
+
+Other factors likely include user carelessness or and ignorance, which is
+exactly what our application is trying to compensate, but fails in these cases.
+
+The cases with false orthographic form don't pose much of a problem. It can
+harden searching for the term in question but performing a search for the
+phonetic respelling or even automatically searching the pronunciation would
+easily mitigate this.
+
+\section{Conclusion}
+
+\section*{Acknowledgments}
+
+The research was supported by SVV project number 260 453.\\
+\\
+This work has been using language resources stored
+and distributed  by the LINDAT/CLARIN project of the Ministry of Education,
+Youth and Sports of the Czech Republic (project LM2015071).
 

### 2018-06-04 2018-itat-krompachy/paper.tex
done writing itat 2018 text

--- 2018-06-04 2018-itat-krompachy/paper.tex
@@
 %% \documentclass[slovensky]{itatnew}
-%% \documentclass[cesky]{itatnew}
+% \documentclass[cesky]{itatnew}
+\usepackage{fontspec}
 
@@
 
-\author{Oldřich Krůza\inst{1}}
+\author{Oldřich Krůza}
 
-\institute{ÚFAL MFF UK,\\
-\email{kruza@ufal.mff.cuni.cz}}
+\institute{
+    Charles University,
+    Faculty of Mathematics and Physics,\\
+    Institute of Formal and Applied Linguistics\\
+    \email{kruza@ufal.mff.cuni.cz}
+}
 
@@
 
+The talks can be seen as companions to Makoň's written works. Together they form
+a unique, extensive, consistent systematization of the spiritual path tailored
+to modern westerners and accessible primarily to Czech speakers. It draws
+heavily on traditional Christian mysticism as well as ancient tradition of India
+and China, adapting them for the present. The whole system can be seen as a
+manual for entering the eternal life prior to the physical death.
+
 There are over 1000 hours of digitized recordings of Karel Makoň, they are
 accessible under the CC-BY license and the project aims at bringing the most
-benefit from them. The first step was digitizing the recordings from the
+benefit out of them. The first step was digitizing the recordings from the
 original magnetic tapes, the second step was releasing all of them on the
@@
 
-The transcription we do is both phonetic and orthographic\footnote{There is no
+The transcription we do is both phonetic and orthographic.\footnote{There is no
 actual focus on orthography. Instead, we mean the natural way of transcribing the
 speech to human-readable text. Where it matters, focus is directed at precise
-correspondence with the utterances instead of language cleanliness}. Our users
+correspondence with the utterances instead of language cleanliness.} Our users
 are supposed provide orthographic transcription where the pronunciation is
@@
 \begin{itemize}
-\item{as a representation of the words uttered and their meaning,}
-\item{on the phonetic level, phoneme for phoneme,}
+\item{as a representation of the words uttered and of their meaning,}
+\item{on the phonetic level, phone for phoneme,\footnote{In the sense that each
+written phoneme corresponds to exactly one uttered phone.}}
 \item{on the time axis}
@@
 Luckily, we can implement automatic measures to aid the annotators to deliver
-higher-quality transcriptions.
+higher-quality transcription.
 
@@
 
-Since we always assume an existing transcription, we can see the user's
-contribution as a correction. Every submission has
+We always assume an existing transcription,\footnote{The seed training data
+was of magnitude in order of minutes and has been annotated using Transcriber.}
+so we can see the user's contribution as a correction. Every submission has
 the form of replacing a text segment with another. Since the transcriptions are
@@
 the case of virtually all programs for computer-aided transcription. For some
-examples, Transcriber\footnote{http://trans.sourceforge.net/}, a veteran
+examples, Transcriber,\footnote{http://trans.sourceforge.net/} a veteran
 open-source transcribing program for Linux, expects the user to provide
 alignment on the level of phrases;
-Transcribe\footnote{https://transcribe.wreally.com/}, a commercial web-based
+Transcribe,\footnote{https://transcribe.wreally.com/} a commercial web-based
 transcribing tool, allows the user to add timestamps anywhere in the text. There
@@
 We have originally built the acoustic model using
-HTK\footnote{http://htk.eng.cam.ac.uk/}, the Hidden Markov Toolkit. Here,
+HTK,\footnote{http://htk.eng.cam.ac.uk/} the Hidden Markov Model Toolkit. Here,
 explicit phonetically labeled training data are necessary for training. We are
 switching to DNN, using Mozilla's
-DeepSpeech\footnote{https://github.com/mozilla/DeepSpeech}, where no explicit
+DeepSpeech,\footnote{https://github.com/mozilla/DeepSpeech} where no explicit
 phonetic annotation is needed but for some purposes like forced alignment, the
@@
 Also, the phonetic labeling is valuable per se for research purposes.
+
+\subsection{Phoneme Set}
+
+We use a subset of PACal\cite{nouza1997phonetic}. We shall also refer to
+individual phonemes in this paper using the PACal notation in \texttt{monospace
+font}. For reference, Table~\ref{tab:phones} lists the phonemes used with their
+IPA notation.
+
+\begin{table*}[htpb]
+\fontspec{DoulosSIL}
+\begin{center}
+\begin{tabular}{|l|l|l|l||l|l|l|l|}
+\hline
+IPA & PACal & common grapheme & IPA & PACal &  common grapheme \\
+\hline
+% TODO: IPA
+a  & a   & a      &     ɱ  & mg  & tra\underline{m}vaj \\
+aː & aa  & á      &     n  & n   & \underline{n}e \\
+aʊ̯ & aw  & au     &     ŋ  & ng  & ta\underline{n}k \\
+b  & b   & b      &     ɲ  & nj  & \v{n} \\
+t͡s & c   & c      &     o  & o   & o \\
+t͡ʃ & ch  & č      &     oː & oo  & ó \\
+d  & d   & d      &     oʊ̯ & ow  & ou \\
+ɟ  & dj  & \v{d}  &     p  & p   & p \\
+d͡z & dz  & dz     &     r  & r   & r \\
+d͡ʒ & dzh & dž     &     r̝̊  & rsh & t\underline{\v{r}}i \\
+ɛ  & e   & e      &     r̝  & rzh & \underline{\v{r}}íz \\
+ɛː & ee  & é      &     s  & s   & s \\
+eʊ̯ & ew  & eu     &     ʃ  & sh  & š \\
+f  & f   & f      &     t  & t   & t \\
+g  & g   & g      &     c  & tj  & \v{t} \\
+ɦ  & h   & h      &     ʊ  & u   & u \\
+i  & i   & i      &     uː & uu  & ú, \r{u} \\
+iː & ii  & í      &     v  & v   & v \\
+j  & j   & j      &     x  & x   & ch \\
+k  & k   & k      &     z  & z   & z \\
+l  & l   & l      &     ʒ  & zh  & ž \\
+m  & m   & \underline{m}ák
+                  &        & sil & \\
+   &     &        &        & sp  & \\
+\hline
+\end{tabular}
+\caption{Phonemes used in transcription}\label{tab:phones}
+\end{center}
+\end{table*}
+\normalfont
 
@@
 Once merged into the displayed transcription, each word's occurrence can be
-edited manually. Hence the user can enter the correct form deviating from Czech
+edited manually. Now the user can enter the correct form deviating from Czech
 pronunciation rules.
@@
 Doing so results in adding the wordform-pronunciation couple to the dynamic
-pronunciation dictionary and is also used at forced alignment. Thus, this
-operation must only be performed once per word and any subsequent time the word
+pronunciation dictionary and is also used for forced alignment. Thus, this
+operation need only be performed once per word and any subsequent time the word
 is entered in its standard orthographic form, the correct pronunciation is
@@
 For example, let's examine the scenario of transcribing the sentence {\em Proč
-se toto nestalo Marii Markétě Alacoque?} Its phonetic representation is {\em p r
-o ch sil s e sil t o t o sil n e s t a l o sil m a r i j i sil m a r k ee tj e
-sil a l a k o k}.
+se toto nestalo Marii Markétě Alacoque? (Why hasn't this happened to Mary
+Margaret Alacoque?)} Its phonetic representation is \texttt{
+    p r o ch sp
+    s e sp
+    t o t o sp
+    n e s t a l o sp
+    m a r i j i sp
+    m a r k ee tj e sp
+    a l a k o k sil
+}.
 \begin{enumerate}
@@
 \item{
-    The phonetic transducer outputs {\em
-        p r o ch sil
-        s e sil
-        t o t o sil
-        n e s t a l o sil
-        m a r i j i sil
-        m a r k ee tj e sil
-        a l a c o k v u e
+    The phonetic transducer outputs \texttt{
+        p r o ch sp
+        s e sp
+        t o t o sp
+        n e s t a l o sp
+        m a r i j i sp
+        m a r k ee tj e sp
+        a l a c o k v u e sil
     }.
@@
     With a bit of luck, the forced alignment fails because of the distiction of
-    the phone sequence {\em k o k} and {\em c o k v u e}.
+    the phone sequence\\ \texttt{k o k} and \texttt{c o k v u e}.
 }
@@
 \item{
-    The user selects the non-existent word {\em alakok} and edits its occurrence
+    The user selects the non-existent word {\em alakok?} and edits its occurrence
     to {\em Alacoque?}
@@
     Alacoque} with any punctuation or capitalization, the pronunciation
-    {\em a l a k o k} is considered at the forced alignment.
+    \texttt{a l a k o k} is inferred by the forced alignment.
 }
@@
 
-As can be seen above, for the sake of simplicity and interoperability, we use
-the set of phonemes as defined by Nouza et al.\cite{nouza1997phonetic} With all
-advantages of this representation, it is clearly not the most natural way for
-lay Czechs to write down literal pronunciation. Thanks to the simple, mostly
+With all advantages of using PACal as a representation for phonemes, it is
+clearly not the most natural way for lay Czechs to write down and read literal
+pronunciation. Thanks to the simple, mostly
 deterministic mapping between phonemes and graphemes, pronunciation respelling
 is a reliable, natural way. There's not even a need for explicit syllable
-separation as seen in English pronunciation respelling {\em
+separation as seen in English pronunciation respelling
 (wikipedia\footnote{https://en.wikipedia.org/wiki/Pronunciation\_respelling}
-gives the example {\em ``Diarrhoea'' is pronounced DYE-uh-REE-a})}.
+gives the example {\em ``Diarrhoea'' is pronounced DYE-uh-REE-a}).
+We postulate that the phonetic respelling is natural to all alphabetized native
+Czech speakers as a fact without any supporting research, based on experience
+alone.
 
@@
 the array of phonemes and the pronunciation
-respelling\footnote{https://github.com/Sixtease/MakonReact/blob/master/src/lib/Phonet.js}.
+respelling.\footnote{https://github.com/Sixtease/MakonReact/\allowbreak{}blob/master/src/lib/Phonet.js}
 
@@
 \begin{enumerate}
-\item{The phone {\em (x)} is spelled {\em ``ch''}.}
-\item{The phone {\em (dz) (dzh)} are spelled {\em ``dz'' ``dž''}.}
-\item{The diphtongs {\em (aw) (ew) (ow)} are spelled {\em ``au'' ``eu'' ``ou''}.}
+\item{The phoneme \texttt{x} is spelled {\em ch}.}
+\item{The phonemes \texttt{dz dzh} are spelled {\em dz dž}.}
+\item{The diphtongs \texttt{aw ew ow} are spelled {\em au eu ou}.}
 \item{
-    Sequences {\em (c) (h), (o) (u), (a) (u), (e) (u), (d) (z), (d zh) } are
-    spelled {\em ``c'h'', ``o'u'', ``a'u'', ``e'u'', ``d'z'', ``d'ž''}. Note
-    though, that the sequence {\em (c) (h)} is purely hypothetical, as it
+    Sequences \texttt{c h, o u, a u, e u, d z, d zh} are
+    spelled {\em c'h, o'u, a'u, e'u, d'z, d'ž}. Note
+    though, that the sequence \texttt{c h} is purely hypothetical, as it
     contradicts voiced/voiceless assimilation.
@@
 \item{
-    Voiceless alveolar fricative trill is explicated as {\em ``r'{}''} as in {\em
-    tř'ít} vs. {\em dřít}.
+    Voiceless alveolar fricative trill is explicated as {\em r'}.
 }
 \item{
-    Palatal nasal and labiodental nasal are spelled {\em ``n'{}'', ``m'{}''}.
+    Palatal nasal and labiodental nasal are spelled {\em n', m'}.
 }
@@
 Still, the user can mark up special-case pronunciation with the apostrophe, like
-the sequence of phonemes {\em (c)} and {\em (h)} with the string {\em c'h}. The
-need has never occurred during the six years' lifespan of the application.
+the sequence of phonemes \texttt{c} and \texttt{h} with the string \texttt{c'h}.
+The need has never occurred during the six years' lifespan of the application.
 
@@
 {\em di ti ni dě tě ně} is ever output. The palatal consonants are always
-explicitly spelled out and e.g. the sequence {\em (n) (i)} is always spelled
-{\em``ny''}
+explicitly spelled out and e.g. the sequence \texttt{n i} is always spelled
+{\em ny}
 
-A few examples of words and their phonetic respelling as output by the
+A few examples of words, pronunciation and phonetic respelling as output by the
 algorithm (given the corresponding pronunciation is input as phoneme list):
 \begin{itemize}
-\item{nic: {\em ňic},}
-\item{kdo: {\em gdo},}
-\item{disk: {\em dysk},}
-\item{dřít: {\em dřít},}
-\item{třít: {\em tř'ít},}
-\item{auto: {\em auto},}
-\item{nauka: {\em na'uka},}
-\item{džbán: {\em džbán},}
-\item{odželezněný: {\em od'železněný},}
-\item{odznak: {\em odznak},}
-\item{podzemí: {\em pod'zemí},}
-\item{noc: {\em noc},}
-\item{tento: {\em tento},}
-\item{hangár: {\em han'gár},}
-\item{samba: {\em samba},}
-\item{tonfa: {\em tom'fa}.}
+\item{nic /\texttt{nj i c}/: {\em ňic},}
+\item{kdo /\texttt{g d o}/: {\em gdo},}
+\item{disk /\texttt{d i s k}/: {\em dysk},}
+\item{dřít /\texttt{d rzh ii t}/: {\em dřít},}
+\item{třít /\texttt{t rsh ii t}/: {\em tř'ít},}
+\item{auto /\texttt{aw t o}/: {\em auto},}
+\item{nauka /\texttt{n a u k a}/: {\em na'uka},}
+\item{džbán /\texttt{dzh b aa n}/: {\em džbán},}
+\item{odžít /\texttt{o d dz ii t}/: {\em od'žít},}
+\item{odznak /\texttt{o dz n a k}/: {\em odznak},}
+\item{podzemí /\texttt{p o d z e m ii}/: {\em pod'zemí},}
+\item{noc /\texttt{n o c}/: {\em noc},}
+\item{tento /\texttt{t e n t o}/: {\em tento},}
+\item{hangár /\texttt{h a ng g aa r}/: {\em han'gár},}
+\item{samba /\texttt{s a m b a}/: {\em samba},}
+\item{tonfa /\texttt{t o mg f a}/: {\em tom'fa}.}
 \end{itemize}
 
-We postulate that the phonetic respelling is natural to all alphabetized native
-Czech speakers as a fact without any supporting research, based on experience
-alone. The use of apostrophe for distinguishing ambiguities and special cases is
+The use of apostrophe for distinguishing ambiguities and special cases is
 not 100\% intuitive and presents another point where instruction is necessary
@@
 One thing we can examine are the approvals / rejections of the forced alignment.
-Of 109640 forced alignment attempts, 3419 have failed, which makes for 3.12\%.
-We have manually inspected 20 random failed attempts and came to the following
-numbers:
+Of 109640 forced alignment attempts, 3419 have failed, which makes for 3.12\%
+rejection rate.  We have manually inspected 20 random failed attempts and came
+to the following numbers:
 \begin{itemize}
@@
 
-\begin{table}[htpb]
+\begin{table*}[htpb]
 \begin{center}
@@
 \hline
-Correct written form        & \#
+Correct spelling            & \#
 	& wrong phonetic pronunc.   & \#
-		& correct pronunc.          & \#
-			& phon. respelling          & \# \\
+		& correct pronunciation     & \#
+			& phon. respel.             & \# \\
 \hline
-Moody & 2 & {\texttt m o o d i} & 0 & {\texttt m uu d i} & 4 & {\em múdy}, {\em můdy} & 2 \\
-Descartes & 2 & {\texttt d e s c a r t e s} & 0 & {\texttt d e k aa r t} & 4 & {\em dekárt} & 2   \\
-Weinfurter & 30 & {\texttt v e j n f u r t e r} & 13 & {\texttt v a j n f u r t r} & 19 & {\em vajnfurtr} & 2 \\
-Michelangelo & 6 & {\texttt m i x e l a ng g e l o} & 2 & {\texttt m i k e l a n dzh e l o} & 4 & {\em mikelandželo} & 0 \\
+Moody & 2 & \texttt{m o o d i} & 0 & \texttt{m uu d i} & 4 & {\em múdy}, {\em můdy} & 2 \\
+Descartes & 2 & \texttt{d e s c a r t e s} & 0 & \texttt{d e k aa r t} & 4 & {\em dekárt} & 2   \\
+Weinfurter & 30 & \texttt{v e j n f u r t e r} & 13 & \texttt{v a j n f u r t r} & 19 & {\em vajnfurtr} & 2 \\
+Michelangelo & 6 & \texttt{m i x e l a ng g e l o} & 2 & \texttt{m i k e l a n dzh e l o} & 4 & {\em mikelandželo} & 0 \\
 \hline
@@
 \end{center}
-\end{table}
+\end{table*}
 
-\begin{table}[htpb]
+\begin{table*}[htpb]
 \begin{center}
@@
  & phonetically correct & phonetically incorrect \\
+\hline
 orthographically correct & 25 & 15 \\
-orthographically incorrect & 6 & 0
+\hline
+orthographically incorrect & 6 & 0 \\
 \hline
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
 the orthographically incorrect form is kept. We attribute this to the fact that
-those who use the phonetic transcription are aware of the problematic and mostly
+those who use the phonetic respelling are aware of the problematic and mostly
 go the whole way and clean up.
@@
 representation. This is a serious problem on at least two levels: Firstly, it
-shows that the forced aligner failed to catch the error. Secondly
+shows that the forced aligner failed to catch the error. Secondly, it lets bad
+examples into the training dataset.
+
+One of the apparent reasons for this to happen is that the dynamic dictionary
+only recognizes exact matches. We can see in one file, for example, all
+occurrences of the form {\em Weinfurter} to have correct pronunciation while
+{\em Weinfurterovi} to have a broken one.
+
+Other factors likely include user carelessness or ignorance, which is
+exactly what our application is trying to compensate, but fails in these cases.
+
+The cases with false orthographic form don't pose much of a problem. It can
+harden searching for the term in question but performing a search for the
+phonetic respelling or even automatically searching the pronunciation would
+easily mitigate this.
+
+The fourth combination of phonetic respelling and false pronunciation is of
+course not occuring.
+
+\section{Conclusion}
+
+We have presented an application that has been providing access to the
+extensive corpus of Karel Makoň and to acquire an almost complete transcription
+thereof. Nearly 70 hours corresponding to over 600,000 word forms have been
+transcribed manually with minimal financial\footnote{In early stages, we kept a paid
+annotator to test the application.} as well as development\footnote{The system
+has been written by a single developer.} costs. Only some of the volunteer have
+indulged instruction time in order of minutes. The rest of the corpus has been
+transcribed using an ASR system trained on these ever-growing data.
+
+We have presented the ways we use to aid the untrained users to provide a
+high-quality orthographic and phonetic time-aligned transcription. We have
+attempted a rough evaluation of the success rate of the measures presented.
+Though clearly far from perfect, they do serve the purpose and set a baseline
+for improvements or novel approaches.
+
+The system has been built with the motivation of spreading the message contained
+in Karel Makoň's talks. However, to make the technology more useful, we are
+actively looking for similar settings where it could be deployed.
+
+\section*{Acknowledgments}
+
+The research was supported by SVV project number 260 453.\\
+\\
+This work has been using language resources stored
+and distributed  by the LINDAT/CLARIN project of the Ministry of Education,
+Youth and Sports of the Czech Republic (project LM2015071).
+
+\bibliographystyle{unsrt}
+\bibliography{citace}
 

### 2018-11-21 2019-cl-cardiff/paper.tex
tracking 2019-cl-cardiff

\begin{document}
\begin{abstract}
\end{abstract}

\section{Introduction}

The way a volume of data is stored impacts all subsequential processing and
exploitation. A decision on storing the data has to be made as soon as the data
is being acquired. Obviously, the difficulty of changing the way of storing the
data increases with the volume of the data stored and with parties relying on
it. The relying parties can be in-house software tools or even public
interfaces\footnote{I hesitate to say API because we're talking about access to
data, not to applications}.

In our case, the data at hand is a collection of speech recordings of total
magnitude of about one thousand hours. The decisions to be made span many
levels and the highest levels dictate further questions. Will the data be stored
as files on a traditional file system or in a database or in a yet another way?
Will the whole data set be contained on one physical storage device or will it
be distributed over several?

I have chosen the traditional way of using files on a single directory tree, so
the next questions are: How will the audio material be divided into individual
files, what will be the directory structure, what will be the file naming and
optionally what additional metadata will be maintained.

In the following chapters, I shall list the original structuring for the Spoken
corpus of Karel Makoň, provide reasons that led to that structuring, explain why
a change was necessary, and describe the journey to realizing it.

\section{Original Structuring}

\subsection{Prior to Digitization}

The corpus arose as a set of amateur recordings on magnetic tapes. The author of
the talks, Mr. Karel Makoň, a Czech mystic, started giving talks in private
groups after his deportation into a concentration camp in 1939, and stopped in
1991, two years before his death. His regular listeners were recording his
talks. I don't know exactly when the recording started but the earliest denoted
date is in 1973.

A part of the recordings (30\% of total length) was taken onto reels, the rest
onto cassettes. Mostly, there was an identifier for a reel tape or for a
cassette. Some pieces were unlabeled and some shared a label.

The majority of the recordings were cassettes taken by a single person with
identifiers of the format YY-NN where YY are the two digits of the year and NN
is the ordinal number in that year, so for example {\texttt 85-05} is the fifth
recording taken in the year 1985. These occupy 686 resulting files out of
total 802 files originating in cassettes.

\subsection{Digitization}

Digitization of the tapes is naturally done one side of the tape into one file.
Any merging or splitting is extra work. Thus, most of the files resulting from
the digitization process correspond to one side of a magnetophone cassette
(worth mostly 45 minutes) or one pass from reel to reel in 9cm/s (90min).

During the two years of digitization, some experiments aiming at saving time and
effort were done. For instance, using a device that can auto-reverse and play
two cassettes in a row.

\section{Need for Change}

\section{New Structuring}

\subsection{Where It Is and Isn't Used}

\subsection{}

\end{document}

### 2018-11-21 2018-clarin-pisa/makon-data.tex
2018 papers

@@
 
-\author{First Author \\
-  Department (optional)\\
-  University of City, Country \\
-  {\tt email@domain} \\\And % if needed: this makes a second column
-  Second Author \\
-  Department (optional)\\
-  University Name without city \\
-  City, Country \\
- {\tt email@domain} \\
+\author{Old\v{r}ich Kr\r{u}za \\
+  Institute of Formal and Applied Linguistics\\
+  Faculty of Mathematics and Physics, Charles University in Prague, Czech
+  Republic \\
+  {\tt kruza@ufal.mff.cuni.cz} %\\\And % if needed: this makes a second column
+%  Second Author \\
+%  Department (optional)\\
+%  University Name without city \\
+%  City, Country \\
+% {\tt email@domain} \\
 % \AND % if needed: this makes a second row
@@
 
+\section*{Acknowledgments}
+
+The research was supported by SVV project number 260 453.\\
+\\
+This work has been using language resources stored
+and distributed  by the LINDAT/CLARIN project of the Ministry of Education,
+Youth and Sports of the Czech Republic (project LM2015071).
+
 % 7.6\% of the corpus is annotated with 620615 : 7574994 hum : auto

### 2018-11-21 2018-ftc-vancouver/webapp.tex
2018 papers

@@
 
-Mako\v{n}'s written works present his teachings in a systematic, comprehensible
+Mako\v{n}'s written works present his teachings in a systematic, comprehensive
 fashion, while the recordings offer bonuses: talks tailored to the audience,
@@
 
-The most common tasks have keyboard shortcuts: \texttt{ctrl+space} for
+The commonest tasks have keyboard shortcuts: \texttt{ctrl+space} for
 play/pause and \texttt{ctrl+enter} for submitting a correction.
@@
 segment is merged into the transcription, in which case we only need to
-recalculate for the words further in the document.
+recalculate for the words further in the document.\footnote{We could even stop
+the recalculation as soon as we find that the new horizontal coordinate of a
+word is left untouched, and add the difference in the vertical coordinate to all
+subsequent words, i.e. when a line stays the same, so do all below it.}
 
@@
 We chose \texttt{ctrl+space} combination. We argue that \texttt{esc} is not the
-best of options because the distance the fingers have to travel from the
-alphanumeric keys causes a noticeable delay. This can lead to missing a pause
-between words. The \texttt{tab} key as chosen by Transcriber is a splendid choice
-from the ergonomy point of view and there is no reason not to use it in a
+best of options for desktops because the distance the fingers have to travel
+from the alphanumeric keys causes a noticeable delay. This can lead to missing a
+pause between words. The \texttt{tab} key as chosen by Transcriber is a splendid
+choice from the ergonomy point of view and there is no reason not to use it in a
 dedicated user interface. However, in the browser, where the \texttt{tab} key
@@
 Another point of near-future endeavor is higher-level work with the contents. By
-this we mean that we wouldd like to use both automatic processing methods and the
+this we mean that we would like to use both automatic processing methods and the
 users to do semantic analysis of the talks: What topic is covered where? What

### 2018-11-21 2018-itat-krompachy/paper.tex
2018 papers

@@
 
+\subsection{Architecture Overview}
+
+The system consists of
+
+\begin{enumerate}
+\item{
+    The corpus in compressed audio format. We use mp3 and ogg/vorbis to
+    accomodate most browsers.  These data are hosted on an external CDN.
+}
+\item{
+    The exact copy of the corpus in parametrized (MFCC) format. These data
+    reside on the back-end server.
+}
+\item{
+    A complete, aligned transcription of the recordings, hosted on the back-end
+    server and mirrored on a CDN.
+}
+\item{
+    Acoustic model trained on the human-transcribed part of the corpus.
+}
+\item{
+    Language model trained using Srilm\cite{stolcke2002srilm} on a combination
+    of publicly available Czech texts, Karel Makoň's written works, and both the
+    human-submitted and automatically-acquired transcription.
+}
+\item{
+    Back-end API for collecting correcions to the transcription, serving the
+    transcription and allowing full-text search with
+    elasticsearch\footnote{https://www.elastic.co/products/elasticsearch}.
+}
+\item{
+    Separately hosted front-end web application serving as an interface for
+    playing the recordings, synchronously displaying the transcriptions and
+    collecting the corrections from users.
+}
+\end{enumerate}
+
+To get the initial transcription, we have manually transcribed some 10 minutes
+of the material using Transcriber\footnote{http://trans.sourceforge.net/},
+trained an acoustic model on it and recognized the whole data using it.
+
 \section{Annotator Expertise}
@@
 
-We always assume an existing transcription,\footnote{The seed training data
-was of magnitude in order of minutes and has been annotated using Transcriber.}
-so we can see the user's contribution as a correction. Every submission has
-the form of replacing a text segment with another. Since the transcriptions are
-time-aligned to the audio, we also know exactly what is the corresponding audio
-segment to the text submitted.
+We always assume an existing transcription, so we can see the user's
+contribution as a correction. Every submission has the form of replacing a text
+segment with another. Since the transcriptions are time-aligned to the audio, we
+also know exactly what is the corresponding audio segment to the text submitted.
 
@@
 False positives (when the system accepts a wrong transcription) present a
-problem, since the error will enter the data set. But false negatives can often
-be circumvented by submitting the transcription divided in different segments.
-Of course, this method can also be used to force a wrong transcription but we
-assume no malevolence on the part of the users.
+problem, since the error will enter the training data set. But users can often
+circumvent false negatives by submitting the transcription divided in different
+segments.  Of course, this method can also be used to force a wrong
+transcription but we assume no malevolence on the part of the users.
 
@@
 the case of virtually all programs for computer-aided transcription. For some
-examples, Transcriber,\footnote{http://trans.sourceforge.net/} a veteran
-open-source transcribing program for Linux, expects the user to provide
-alignment on the level of phrases;
+examples, Transcriber, a veteran open-source transcribing program for Linux,
+expects the user to provide alignment on the level of phrases;
 Transcribe,\footnote{https://transcribe.wreally.com/} a commercial web-based
@@
 \begin{center}
-\begin{tabular}{|l|l|l|l||l|l|l|l|}
+\begin{tabular}{|l|l|l||l|l|l|}
 \hline
@@
 Still, the user can mark up special-case pronunciation with the apostrophe, like
-the sequence of phonemes \texttt{c} and \texttt{h} with the string \texttt{c'h}.
+the sequence of phonemes \texttt{o} and \texttt{u} with the string \texttt{o'u}.
 The need has never occurred during the six years' lifespan of the application.
@@
 \end{tabular}
-\caption{Examples of non-standard pronunciation in the manually transcribed data}\label{tab:pronunc}
+\caption{Examples of non-standard pronunciation in the manually transcribed data}
+\label{tab:pronunc}
 \end{center}
@@
 \caption{Success rate for phonetic and orthographic representation of foreign
-words}\label{tab:pronunc-rate}
+words based on data from table\ref{tab:pronunc}}
+\label{tab:pronunc-rate}
 \end{center}
@@
 annotator to test the application.} as well as development\footnote{The system
-has been written by a single developer.} costs. Only some of the volunteer have
+has been written by a single developer.} costs. Only some of the volunteers have
 indulged instruction time in order of minutes. The rest of the corpus has been

### 2019-01-04 2019-cicling-la-rochelle/kmgram.tex
add draft for cicling 2019

\documentclass{llncs}
\usepackage{graphicx}

\begin{document}
\newtheorem{Definition}{Definition}
\title{Dynamic Grammars for Embedded ASR}

%\author{Old\v{r}ich Kr\r{u}za}
\author{Anonymous}

%\institute{Charles University in Prague\\
%           Faculty of Mathematics and Physics\\
%           Institute of Formal and Applied Linguistics\\
%           Malostransk\'{e} n\'{a}m\v{e}st\'{i} 25, Prague\\
%           Czech Republic\\
%           kruza@ufal.mff.cuni.cz, vk@ufal.mff.cuni.cz}
\institute{Anonymous university\\
           Anonymous institute \\
           Address\\
           Country\\
           E-mail}

\maketitle

\begin{abstract}
This paper presents a speech recognition system embedded in a wearable device
where many applications can run and each defines its own set of recognized
commands. The command set for each application can be fixed or dynamically
created from available data.
\end{abstract}

\section{Introduction}
Speech recognition has undergone a massive improvement since the adpotion of
deep neural networks in the recent years\cite{hinton2012deep}, although the
practice is much older\cite{morgan1995neural}. Modern hardware enables ASR
systems to operate with dictionaries in the magnitude of hundreds of thousands
of words and even to recognize words out of vocabulary. However, when embedding
an ASR system\cite{huggins2006pocketsphinx} to an outdated device with limited
CPU and memory, and with battery life concerns, it can be desirable or necessary
to sacrifice some of the power and flexibility.

We present one such setting, where voice control for a wearable device running
Ubuntu 12.04 on an ARM architecture is developed.

\section{Device}

The device in our concern is the {\em AIR\textsuperscript{e} Lens}, a smart glass by
Konica Minolta designed for use in industry. There
is a head-mount display attachable to a protective glass or worn
directly with a detachable nose support. Connected by a cable is a
handheld-sized control box with the computing unit that features an ARM CPU, 1GB RAM and
runs Ubuntu 12.04\footnote{Newer versions of the operating system are not
supported.}

The head-mount display has a gyroscopical accelerometer, a compass, a camera, a
microphone, and a monocular optical element. The control box has a spare USB
slot and 9 bulky rubber buttons suitable for use in protective gloves.

\subsection{Control Methods}

The glass can be most reliably controlled by the nine buttons but their use
requires dedicating a hand and reaching out for the control box, which is
normally worn on a belt. This can be impractical in situations where the user is
performing some task using both hands. The small number of buttons further
limits the comfort in performing complex interactions.

Further input methods include gestures, head motion and pressing a wrist-worn
bluetooth button. All of these input methods, however, have a very limited
symbol set. Possible solutions are offered by image (video) recognition and
speech recognition. Indeed, a QR reader app is used for input of text strings,
like URLs or WiFi configurations.

\section{Voice Control}

For implementing speech recognition for the smart glass, the first fundamental
question is whether to embed the engine to the glass or run it remotely. Since
offline functionality is a highly valued feature in the intended use cases, the
embedded variant is clearly desirable.

The second fundamental question is the type of the acoustic model. The number
one choice are of course the deep neural networks. The only thinkable competitor
could be the old school hidden Markov models. The major pros of the neural
networks are better fundamental suitability for acoustic modeling and thus
implied higher accuracy, and also the option of skipping explicit dealing with
the phonetic transcription.
The arguments for HMMs are lesser computational intensity and maturity of tools.

As a compromise, we have chosen the excellent free software speech recognition
tool Julius\cite{lee2001julius}, \cite{lee2009recent}, which is very stable,
feature-rich and allows HMM as well as GMM-DNN-hybrid models.

\section*{Acknowledgments}

This work has been using language resources developed, stored and distributed by the LINDAT/CLARIN project of the Ministry of Education, Youth and Sports of the Czech Republic (project LM2010013).
\\
This work was supported by the Grant of Czech Science Foundation (GA\v{C}R) No. P202/10/1333.
\\
This work was supported by the Charles University Grant Agency (GAUK) Grant No. 920913.
\\
Participation in the conference was supported by Foundation of Vilem Mathesius.
\\
This research was partially supported by SVV project number 260 104.

\bibliographystyle{splncs}

\begin{thebibliography}{}

\end{thebibliography}
%\bibliography{paper}
\end{document}

### 2019-01-05 2019-cicling-la-rochelle/kmgram.tex
wip: cicling 2019

@@
 As a compromise, we have chosen the excellent free software speech recognition
-tool Julius\cite{lee2001julius}, \cite{lee2009recent}, which is very stable,
-feature-rich and allows HMM as well as GMM-DNN-hybrid models.
+tool Julius\cite{lee2001julius}, \cite{lee2009recent}, which is very mature,
+feature-rich and allows HMM as well as DNN and GMM-DNN-hybrid models.
+
+\subsection{Acoustic Modeling}
+
+We have used the Mozilla Common Voice\footnote{voice.mozilla.org} corpus for
+training, the CMU Sphinx\cite{lamere2003cmu} dictionary and their
+Tensorflow\cite{abadi2016tensorflow}-based pronunciation tool named {\em
+g2p-seq2seq} for phonetic transcription. For training of the models, we have
+used HTK\cite{young2002htk}. We use tied-state triphones clustered by decision
+trees and Gaussian mixture models with 8 mixtures.
+
+The model in binary form has 23MB and takes about 2 seconds to load on the
+device.
+
+\subsection{Usage}
+
+The glass runs dedicated applications to perform its various tasks. All of those
+applications are designed to be controlled by four directional arrow keys and a
+confirm key. The remaining four keys present on the control box are for
+operations outside the scope of an application. For example, turning the display
+on and off, exiting to the main screen etc.
+
+By default, the voice control uses a grammar that copies the buttons, i.e. it is
+in an isolated-word mode with vocabulary of {\em next, back, up, down, OK}.
+Some synonyms are also recognized, like {\em left} for {\em back}, {\em upward}
+for {\em up} or {\em enter} for {\em OK}.
+
+Whenever one of these commands is recognized, pressing the corresponding key
+is triggered.
+
+Every application can either rely on this behavior or capture the recognition
+output, handling it as it sees fit. The recognition grammar can be redefined, so
+the set of commands is changed as long as the app is running.
+
+For example, an application that offers a list of items for selection, can
+define the recognition grammar to accept commands {\em next}, {\em back}, {\em
+OK} and the names of the offered items. Of course, this requires that the
+pronounciation of each item is known. We do not run g2p-seq2seq on the device
+for performance reasons, so this has to be taken care of in advance.
+
+\section{Architecture}
+
+The glass uses the web platform internally. Firefox (version 52) serves web applications
+locally via the \texttt{file://} protocol. There is a Java-based middleware
+layer called the SG server, which provides endpoints for camera, logs and other
+device functionality. Julius can run either as a command-line application
+outputting to standard output, or as a TCP server capable of handling one single
+connection at a time.
+
+The browser supports websockets, which is the most suitable technology for
+accepting a continuous data stream over extended time periods without the
+necessity of polling on the client side. To bridge the
+websocket-protocol-speaking client with the TCP-speaking server, and to mitigate
+the limitation to a single connection, we have created a {\em broker}
 
@@
 
-\begin{thebibliography}{}
+\bibliography{citace}
 
-\end{thebibliography}
-%\bibliography{paper}
 \end{document}

### 2019-02-15 2019-cicling-la-rochelle/cardiff.tex
adapt rejected cardiff paper for cicling

\documentclass{llncs}
\usepackage{graphicx}

\begin{document}
\newtheorem{Definition}{Definition}
\title{Restructuring Spoken Corpus for Streaming Emulation}

%\author{Old\v{r}ich Kr\r{u}za}
\author{Anonymous}

%\institute{Charles University in Prague\\
%           Faculty of Mathematics and Physics\\
%           Institute of Formal and Applied Linguistics\\
%           Malostransk\'{e} n\'{a}m\v{e}st\'{i} 25, Prague\\
%           Czech Republic\\
%           kruza@ufal.mff.cuni.cz, vk@ufal.mff.cuni.cz}
\institute{Anonymous university\\
           Anonymous institute \\
           Address\\
           Country\\
           E-mail}

\maketitle

\begin{abstract}
This paper shares the experience of transforming the way of storing a spoken
corpus into more numerous smaller files. Discussed is the motivation: a change
in the usage of the data; the corpus itself: a 1000-hour set of recordings of a
single speaker; and decisions to be made for the restructuring, reasons for them
and their impact.
\end{abstract}

\section{Introduction}

The way a volume of data is stored impacts all subsequent processing and
exploitation\cite{reppen2010building}. A decision on storing the data has to be made as soon as the
data is being acquired. Obviously, the difficulty of changing the way of storing
the data increases with the volume of the data stored and with more parties relying
on it. The relying parties can be in-house software tools or even public
interfaces (I hesitate to say API because we're talking about access to data,
not to applications).

In my case, the data at hand is a collection of speech recordings of total
magnitude of about one thousand hours. The decisions to be made span many levels
and the highest levels dictate further questions. Will the data be stored as
files on a traditional file system or in a database or in a yet another way?
Will the whole data set be contained on one physical storage device or will it
be distributed over several? How will the audio material be divided into
individual files and directories? What additional metadata will be
maintained?\cite{crowdy1993spoken}

In the following chapters, I shall list the original structuring for the Spoken corpus of Karel Makoň, provide reasons that led to that structuring, explain why a change was necessary, and describe the journey to realizing it.

\section{The Spoken Corpus of Karel Makoň}

The corpus arose as a set of amateur recordings on magnetic tapes. The author of
the talks, Mr. Karel Makoň\cite{hajek2007cesky}, a Czech mystic, started giving talks in private
groups after his deportation into a concentration camp in 1939, and stopped in
1991, two years before his death. His regular listeners were recording his
talks. I don't know exactly when the recording started but the earliest denoted
year is 1973. The earliest year I could make out from analyzing the contents is
1970 on an undated recording.

Makoň had groups of attendants in various places of then Czechoslovakia. There
was a group in Pilsen, a group in Prague and a group in Gottwaldov, today's
Zlín. Some of the recordings are from what we would now call retreats -- meetings
in a group in detached places that lasted several days. One such place was in
Kaly near Brno in south Moravia. Another was in the cottage Čeřínek in the
Bohemian-Moravian Highlands. Yet another was in Žiar, Slovakia. Other recordings are mostly from meetings hosted by one of
the members.

\subsection{Corpus Content}

The talks can be seen as accompaniment to the author's written opus. Karel Makoň
wrote twenty seven books on his own and translated and commented twenty eight
books of other authors. Since the culmination of his spiritual path in the
concentration camp in Sachsenhausen, where he acquired profound understanding of
symbolism in Christianism and other traditions, he kept translating his
experience to words in order to enable others to find enlightenment.

His work can be compared to that of other modern spiritual masters, though there
are pecularities.

Makoň bases his teachings mainly on Christian mystic, drawing
heavily from the catholic tradition. He repeatedly uses the parable of the
prodigal son and the parable of the talents in his arguments. He stresses that
the words of Jesus must always be considered within the context of his deeds and
always as a whole. He refers very often to Saint Teresa of
Ávila, Augustine of Hippo and Padre Pio among others. He criticizes the catholic
tradition for some points that he claims form an obstacle on the path to God and
are in conflict with Jesus' intent. One such example is the belief that the
eternal life can only be reached after the physical death and that a life of
virtues leads to it. Makoň claims that the eternal life -- being consciously
eternal -- can only be reached within the physical body and that virtues alone
never suffice to this end.

He also draws from ancient Indian tradition, notably referring to
Sat-Chit-Ananda as the three qualities of God. He translated and commented a
book by Sri Aurobindo Ghos and draws from it. He also often refers to the life
of Siddhartha Gautama Buddha, notably for his enlightenment in the moment of
discarding the path of asceticism. He suggests that the traditional hinduistic
concept of re-incarnation should be revised firstly because a better
understanding of what happens with the soul is at hand, secondly because, as he
claims, correctly used Christianity leads to redemption within one incarnation.

Aside from the symbolism in the New Testament, Makoň often uses milestones in
his own life for explaining the general spiritual path. The most important of
these milestones are repeated surgery without anaesthesia in the age of six, and
confrontation with his own certain death in the concentration camp.

\subsection{Original Recording}

The vast majority of the recordings have been taken by one of the members
of the Gottwaldov group. Fortunately, these recordings were taken using the best
amateur technology available at that time and stored very carefully. Recordings
from other places also exist but rather as rarities, with inferior acoustic
quality and lacking systematic identification.

A part of the recordings (30\% of total length) was taken onto reels, the rest onto cassettes. Mostly, there was an identifier for a reel tape or for a cassette. Some pieces were unlabeled and some shared a label.

The majority of the recordings are cassettes with
identifiers of the format \texttt{YY-NN}, for example \texttt{85-05} is the fifth recording taken in the year 1985. These occupy 686 resulting files out of total 802 files originating in cassettes.

Of the 39 self-digitized reel-to-reel tapes, 36 were recorded at 9.53cm per
second. The remaining three in 2.38cm per second. The latter took six hours to
play back in one direction and the quality was heavily impaired. Of the
reel-to-reel tapes, 24 are identified by a letter. The sequence
is roughly alphabetical, though there
are three distinct ones identified by the letter I, and there is none with the
letter G, which could have been lost. 85 are identified with year spanning from
1973 to 1988. There are may duplicities, however, so the actual number of
distinct recordings is likely much lower in this category. 10 have a numerical
identifier, 29 have a textual identifier, and 2 have no identifier whatsoever.

There is also a three-hour video on youtube: \\
\texttt{https://www.youtube.com/watch?v=UaNm9jnnJiA}

\section{Digitization}

For cassettes, the digitization has mostly been done one side of a cassette to
one file. For reel-to-reel, it was one channel of one pass from reel to reel to one file. Notable exceptions are cassettes digitized in auto-reverse mode, which has been experimented with during the two years of the digitization.

An exhaustive list of the volumes corresponding to each digitized file follows:
\begin{enumerate}
    \item{sides of cassettes: 615 files,}
    \item{whole cassettes: 140 files,}
    \item{reel-to-reel passes: 112 files,}
    \item{imported, uncertain: 222 files,}
    \item{two concatenated cassettes: 1 file.}
\end{enumerate}

The imported files are those that have been digitized by other parties prior to my own  effort. The format for digitization was 48kHz, 16bit, real time. An exception to real-time digitization were the three reels recorded in 2.38cm/s. These have been digitized in the standard speed of 9.53cm/s and had the sample rate set to one quarter of the nominal value.

\section{Usage of Digitized Material}

The digitized audio files have been used for two purposes. On one hand for
direct distribution over physical media and on the other hand in a dedicated web
application that has been serving for direct playback in the browser and for
acquisition of manual corrections to ASR-generated\cite{ircing2001large} transcription. There were
two generations of said web application, see Krůza 2012\cite{kruuza2012making} for the first one,
and Krůza 2018\cite{biblio:KrKuSecondGenerationWeb2018} for the second one.

The original version used the HTML audio element for playback. This suffers from
poor timing precision when playing back specific words but it handles streaming
very well. The next generation of the web application uses the Web Audio API,
which remedies the precision issues at the cost of streaming capability.

Initially, the second generation web application would load the whole recording,
decode it in-memory and only then enable playback and other operation. Since 90
minutes is a very common length of a recording (264 of the total 1090 files are
over 80 minutes long), the load on the user computer was very heavy. A raw,
decoded 90-minute monaural recording in 24KHz occupies about 240MB. However,
after decoding a recording like that, the browser would often occupy triple that
amount of memory. Even when the user's computer could handle that, the waiting
time was in order of minutes, which ensured a very poor user experience.

To remedy this issue, a change was necessary.

\section{New Structuring}

To enable the web application to access random segments of a long recording, while keeping the precision provided by Web Audio API, these options emerge: 
\begin{enumerate}
    \item{pre-load and decode the whole recording,}
    \item{serve ranges by cutting the audio on the server side,}
    \item{store the recordings in smaller chunks,}
    \item{wait for streaming to be supported.}
\end{enumerate}

Option 1 was our baseline as discussed above.

Option 2 - serving arbitrary ranges cut on demand by the server seems quite
viable. It is flexible and not very hard to implement. One downside is that the
server must be able to script as well as access the whole recordings. Since the
amount is in order of tens of gigabytes even using compressed audio formats, it
limits the options for hosting and increases costs. Another downside is with
caching. Repeated sessions with the same recording would likely lead to
different ranges being requested, thus disabling the advantages of caching.
Lastly, computation-heavy operations on the server side can impair site
reliability or require costly cloud solutions.

Option 3 - storing the corpus in smaller chunks has the downside of serving unnecessary context when requesting a specific range. Also, if we want to retain full download capability, we must either stitch the recording together or host both split and integral versions of all recordings.

Option 4 - waiting for streaming support in Web Audio API might seem silly but the issue has been discussed by the Web Audio API developer team, see https://github.com/WebAudio/web-audio-api/issues/1305 .

I opted for storing the corpus in smaller chunks. The downside of unnecessary context is a minor one, especially if we choose a fitting length of the chunks. Having to host two versions of the corpus raises the costs in my case by less than one US Dollar per month.

\section{The Process of Restructuring}

There are four choices to be made for the restructuring: 
\begin{enumerate}
    \item{how long will the chunks be,}
    \item{how to choose individual split points,}
    \item{what will be the directory structure,}
    \item{the file names.}
\end{enumerate}

\subsection{Segment Length}

The length of the chunks should be no more than 5 minutes. With 24kHz
compressed audio, a 5-minute chunk takes up about 1.5MB. Theoretically, with
modern 3G connections, even larger downloads should be instantaneous. However,
the reality diverts from table speeds and we also have to wait for decoding,
which takes about 4 seconds for a 5-minute chunk with an Intel Core2 @ 2.5GHz.
According to the magazine UXMovement\cite{foursecondrule}, four seconds are the
threshold above which the user starts to abandon the original intent. The Nielsen
Norman Group\cite{websiteresponsetimes} claims the same for ten seconds.

The lower boundary is much freer but there always is a chance of an artifact at
the glue point during playback, so the less of them the better. Also, each chunk
means an extra HTTP request, which itself has considerable overhead.

A reasonable compromise seems to be a span of 30 to 120 seconds for a chunk.

\subsection{Finding Split Points}

By choosing split points well, the impact of occasional artifacts caused by
imprecisely switching segments during playback can be reduced. Ideally,
splitting should be performed in pauses between sentences or at least between
words. It is practically impossible to speak for two minutes without taking a
breath, so there should always be a pause to be found for a segment's
boundaries.

There are numerous ways to find a moment of silence in an audio recording. The
most requiring and most reliable way is manual annotation.

Another reliable method is by looking for predicted silences in
phone-level-aligned transcription. This method could largely be used because there are
manual or automatically-acquired transcriptions to most of the recordings in the
corpus.

Where either speech recognition or its forced alignment failed, and thus there
is no transcription, another method comes to question: Voice-activity detection
(VAD).  I have used the perl module Audio::FindChunks for this purpose. This
method, unfortunately, is not very reliable in case of audio input with low
signal-to-noise ratio, which is an abounding phenomenon in the Spoken Corpus of
Karel Makoň.

Where not even VAD helps, which can be identified so that the detected chunks
are too long, there only remains the blunt method of creating fixed-size
segments, disregarding the frequent case of splitting inside of a word.

Two of these methods have been eliminated by experiments: Manual annotation was
way too inefficient. Six volunteers attempted this task, losing their patience
after zero to ten minutes of transcribed material. Furthermore, detection by aligned
transcription could be used for some recordings where initially speech
recognition failed. By splitting the recordings into smaller chunks of fixed
size and then
recognizing each segment apart, the problematic recordings could be
speech-recognized. The resulting accuracy was catastrophic given the bad
acoustic quality of these recordings but the
longer silences were mostly precisely detected\footnote{There are no gold
standard data, so there's no way of evaluating the result exactly but so far, I
found no split in the middle of a word using this technique.} The chunks where
the ASR failed again, were also beyond what VAD could handle, so there was no reason
to resort to it.

Table~\ref{tab:splitpoints} summarizes the number of split points acquired by
each method. The significant number of split points acquired by fix-size
spliting is due to the fact that there are still recordings for which the
split-and-recognize procedure is yet to be performed.

\begin{table}[htpb]
\begin{center}
\begin{tabular}{|l|l|l|l||l|l|l|l|}
\hline
acquisition method & number of uses \\
\hline
manually & 0 \\
by aligned transcription & 60424 \\
by voice activity detection & 0 \\
fix size & 22043 \\
total & 82467 \\
\hline
\end{tabular}
\caption{number of split points by their method of
acquisition}\label{tab:splitpoints}
\end{center}
\end{table}

\subsection{Split Point Selection}

For fix-size segment splitting, I chose a segment length of 60 seconds. For
splitting by aligned transcription, the task at hand is one where the input is a
sequence of silences in the recordings, each identified by start and end, and
the output is an optimal subsequence of these silences. The weight used is
length of the silence (the longer the pause, the better it is as a split point).
A constraint is the distance of the selected silences to be between 30 and 120
seconds.

Notice the task has no solution if the recording is shorter than 30 seconds.
However, this never occurs in the Spoken Corpus of Karel Makoň and even if it
did, we could simply leave the short recording in one segment.

Another case where there is no solution is when the distance between adjacent
silences (strictly speaking between the mid-points of adjacent silences) is
greater than 120 seconds. Such a case occurs when the silences themselves are
very long. I have solved such cases manually but an automated approach would
also be very simple.

The algorithm sought seems to be a typical demonstration of dynamic programming:
Find an optimal split of a part of the recording that only consists of the first
word, then add next word and find the optimal split based on the current one.

There is also a simpler variant, however:
\begin{enumerate}
\item{start with the set of all
silences;}
\item{iterate over them from shortest to longest;}
\item{remove the silence
from the set if joining its adjacent segments doesn't yield one that's longer
than 60 seconds;}
\item{end iteration;}
\item{iterate again over chosen silences;}
\item{remove a silence from the
set if one of its adjacent segments is shorter than 30 seconds;}
\item{end iteration;}
\end{enumerate}

This algorithm is very easy to program and it has linear complexity.

\subsection{File Naming}

Naming of the chunks is arbitrary as long as the file names are unique. But a
good naming scheme can help with further processing of the data. A good naming
scheme can be kept when the split points are changed for a recording, without
collisions. This excludes simple ordinal numbering. Another criterion is ease of
filtering and parsing of the file names. Lastly, the if the file name is
descriptive and human-readable, it can save time and effort. To accommodate
these criteria, I chose to name the chunks \\
\texttt{{\em{}recording-identifier}--from-{\em{}start-time}--to-{\em{}end-time}.{\em{}format}} \\
for example \\
\texttt{87-25B--from-2186.45--to-2239.63.ogg}.

To ensure smooth transitions, the segments are encoded 500 milliseconds longer
than declared. This is to compensate for rounding of the length of the split
file when encoded in a compressed format. Of course, this commands for the
algorithm used in the web application to switch playback from one segment to
another before the actual last sample of the expiring segment.

\section{Reproduction}

The corpus can be accessed through its web interface at
\texttt{http://radio.makon.cz/}. The playback and other operations like playing
back or downloading an arbitrary span (which can span a segment boundary) can be
tested there. The source code for splitting the audio can be accessed at
GitHub: \\
\texttt{https://github.com/Sixtease/CorpusMakoni/tree/master/scripts}.
Similarly, the source code for the JavaScript part that works with the segments
can be found at its GitHub repository:
\texttt{https://github.com/Sixtease/MakonReact}, specifically in the files
\texttt{src/store/audio.js} and \texttt{src/store/AudioChunks.js}.

\section{Conclusion}

The new segmented structuring is used in the web application to emulate streaming capability while exploiting the advantages offered by Web Audio API. Other uses of the corpus, like direct downloads, archiving and on-demand forced alignment still rely on the original structuring where one recording of length in order of tens of minutes corresponds to a single file.

The redundant storage needs impose a negligible financial overhead but the work necessary for creation and maintenance of these extra files bring about an even stronger wish for the Web Audio API to support streaming. However, storing the corpus in files of a smaller size and devising the mechanism to seamlessly glue the chunks together during playback and transcription was a worthy lesson with potential use in other settings.

%
%\section*{Bibliography}
%    1. Reppen, Randi: Building a corpus: what are the key considerations?
%in collection The Routledge handbook of corpus linguistics, pages 59-65, 2010, Routledge
%    2. Crowdy: Spoken corpus design
%in journal Literary and Linguistic Computing 1993/4, pages 259-265, 1993, Oxford University Press
%    3. Hájek: Český mystik Karel Makoň
%in journal Dingir 2007/4, ISSN 1212-1371, pages 142-143, ed. Vojtíšek, 2007, Dingir s.r.o.
%    4. Ircing, Krbec, Hajič, Psutka, Khudanpur, Jelinek, Byrne: On large vocabulary continuous speech recognition of highly inflectional language-Czech
%in proceedings of Seventh European Conference on Speech Communication and Technology, 2001
%    5. Krůza, Peterek: Making Community and ASR Join Forces in Web Environment
%in proceedings of International Conference on Text, Speech and Dialogue, pages 415-421, 2012, Springer
%    6. Krůza, Kuboň: Second-Generation Web Interface to Correcting ASR Output
%in Proceedings of the Future Technologies Conference (FTC) 2018, ISBN 978-3-030-02685-1, ISSN 2194-5357, volume 1, pages 749-762, 2018, Springer
%    7. Psutka, Ircing, Psutka, Hajič, Byrne, Mírovský: Automatic transcription of Czech, Russian, and Slovak spontaneous speech in the MALACH project
%in proceedings of Ninth European Conference on Speech Communication and Technology, ISBN 1018-4074, pages 1349-1352, 2005, ISCA
%

\section*{Acknowledgments}

This work has been using language resources developed, stored and distributed by the LINDAT/CLARIN project of the Ministry of Education, Youth and Sports of the Czech Republic (project LM2010013).
\\
This work was supported by the Grant of Czech Science Foundation (GA\v{C}R) No. P202/10/1333.
\\
This work was supported by the Charles University Grant Agency (GAUK) Grant No. 920913.
\\
Participation in the conference was supported by Foundation of Vilem Mathesius.
\\
This research was partially supported by SVV project number 260 104.

\bibliographystyle{splncs}

\bibliography{citace}

\end{document}

### 2019-04-19 2019-cicling-la-rochelle/cardiff.tex
stuff

@@
 
-%\author{Old\v{r}ich Kr\r{u}za}
-\author{Anonymous}
+\author{Jan Old\v{r}ich Kr\r{u}za}
+%\author{Anonymous}
 
-%\institute{Charles University in Prague\\
-%           Faculty of Mathematics and Physics\\
-%           Institute of Formal and Applied Linguistics\\
-%           Malostransk\'{e} n\'{a}m\v{e}st\'{i} 25, Prague\\
-%           Czech Republic\\
-%           kruza@ufal.mff.cuni.cz, vk@ufal.mff.cuni.cz}
-\institute{Anonymous university\\
-           Anonymous institute \\
-           Address\\
-           Country\\
-           E-mail}
+\institute{Charles University in Prague\\
+           Faculty of Mathematics and Physics\\
+           Institute of Formal and Applied Linguistics\\
+           Malostransk\'{e} n\'{a}m\v{e}st\'{i} 25, Prague\\
+           Czech Republic\\
+           kruza@ufal.mff.cuni.cz}
+%\institute{Anonymous university\\
+%           Anonymous institute \\
+%           Address\\
+%           Country\\
+%           E-mail}
 
@@
 \\
-This work was supported by the Grant of Czech Science Foundation (GA\v{C}R) No. P202/10/1333.
-\\
-This work was supported by the Charles University Grant Agency (GAUK) Grant No. 920913.
-\\
-Participation in the conference was supported by Foundation of Vilem Mathesius.
+Participation in the conference was supported by COST action IS1404.
 \\

### 2019-04-19 2019-cicling-la-rochelle/kmgram.tex
stuff

@@
 the limitation to a single connection, we have created a {\em broker}
+application that acts on one hand as the sole client for Julius' TCP server and
+on the other hand as a websocket server for any number of connections,
+forwarding the traffic from Julius to all clients and vice versa. Whenever there
+is no active client connection, the default recognition grammar is restored and
+button keypresses emulated.
 

### 2019-04-19 2019-cl-cardiff/paper.tex
stuff

@@
 The way a volume of data is stored impacts all subsequential processing and
-exploitation. A decision on storing the data has to be made as soon as the data
+exploitation, see Reppen 2010\cite{reppen2010building}.
+A decision on storing the data has to be made as soon as the data
 is being acquired. Obviously, the difficulty of changing the way of storing the
@@
 
+Of the 39 self-digitized reel-to-reel tapes, 36 were recorded at 9.53cm per
+second. The remaining three in 2.38cm per second. The latter took six hours to
+play back and the quality was heavily impaired.
+
 \subsection{Digitization}
 
-Digitization of the tapes is naturally done one side of the tape into one file.
-Any merging or splitting is extra work. Thus, most of the files resulting from
-the digitization process correspond to one side of a magnetophone cassette
-(worth mostly 45 minutes) or one pass from reel to reel in 9cm/s (90min).
+The digitization has mostly been done one side of a cassette to one file and one
+channel of one pass from reel to reel to one file. Notable exceptions are
+cassettes digitized in auto-reverse mode, which has been experimented with during
+the two years of the digitization.
 
-During the two years of digitization, some experiments aiming at saving time and
-effort were done. For instance, using a device that can auto-reverse and play
-two cassettes in a row.
+An exhaustive list of the volumes corresponding to each digitized file follows:
 
-\section{Need for Change}
+\begin{enumerate}
+\item{sides of cassettes: 615 files;}
+\item{whole cassettes: 140 files;}
+\item{reel-to-reel passes: 112 files;}
+\item{imported, uncertain: 222 files;}
+\item{two concatenated cassettes: 1 file;}
+\end{enumerate}
+
+The imported files are those that have been digitized by other parties prior to
+my own  effort. The format for digitization was 48kHz, 16bit, real time. An
+exception to real-time digitization were the three reels recorded in 2.38cm/s.
+These have been digitized in the standard speed of 9.53cm/s and had the sample
+rate set to one quarter of the nominal value.
+
+\section{Usage of Digitized Material}
+
+The digitized audio files have been used for two purposes. On one hand for
+direct distribution over physical media and on the other hand in a dedicated web
+application that has been serving for direct playback in the browser and for
+acquisition of manual transcription. There were two generations of said web
+application, see Krůza 2012\cite{kruuza2012making} for the first one, and Krůza
+2018\cite{biblio:KrKuSecondGenerationWeb2018} for the second one.
+
+The original version used the HTML audio element for playback. This suffers from
+poor timing precision when playing back specific words but it handles streaming
+very well. The next generation of the web application uses the Web Audio API,
+which remedies the precision issues at the cost of streaming capability.
+
+To prevent the necessity to load the whole recording, which is often 90 minutes
+long, a change was inevitable.
 
 \section{New Structuring}
+
+To enable the web application to access random segments of a long recording,
+while keeping the precision provided by Web Audio API, these options emerge:
+
+\begin{enumerate}
+\item{pre-load and decode the whole recording;}
+\item{serve ranges by cutting the audio on the server side;}
+\item{store the recordings in smaller chunks;}
+\item{wait for streaming to be supported;}
+\end{enumerate}
+
+Option 1 - pre-loading the whole recording is the easiest one to implement but
+it has unbearable downsides for the user. The recoding not only must be
+completely downloaded, it also must be decoded. This takes minutes to complete.
+Worse yet, it can freeze the computer by using up all memory.
+
+Option 2 - serving arbitrary ranges cut on the server side seems quite viable.
+It is flexible and not very hard to implement. One downside is that the server
+must be able to script as well as access the whole recordings. Since the amount
+is in order of tens of gigabytes even using compressed audio formats, it limits
+the options for hosting and increases costs. Another downside is with caching.
+Repeated sessions with the same recording would likely lead to different ranges
+being requested, thus disabling the advatages of caching.
+
+Option 3 - storing the corpus in smaller chunks has the downside of serving
+unnecessary context when requesting a specific range. Also, if we want to retain
+full download capability, we must either stitch the recording together or host
+both split and integral versions of all recordings.
+
+Option 4 - waiting for streaming support in Web Audio API might seem silly but
+the issue has been discussed by the Web Audio API developer team.\footnote{see
+https://github.com/WebAudio/web-audio-api/issues/1305}
+
+In the end I opted for storing the corpus in smaller chunks. The downside of
+unnecessary context is a minor one, especially if we choose a fitting length of
+the chunks. Having to host two versions of the corpus raises the costs in my
+case by less than a US Dollar per month, which is bearable.
+
+\subsection{The Process of Restructuring}
+
+There are three choices to be made for the restructuring:
+\begin{enumerate}
+\item{how long will the chunks be;}
+\item{how to choose individual split points;}
+\item{what will be the directory structure;}
+\end{enumerate}
+
+The length of the chunks should be no more than 2 minutes. With 16kHz compressed
+audio, a 2-minute chunk takes up about 500kB. This
 

### 2019-05-15 2012-tsd-brno/MakonFM-final/tsd-makonfm.tex
add 2012 TSD article

\documentclass{llncs}
%\usepackage[cp1250]{inputenc}
\usepackage[utf8]{inputenc}
%\usepackage[czech]{babel}
\usepackage{graphicx}
\usepackage{tipa}

\begin{document}
\newtheorem{Definition}{Definition}
\title{Making Community and ASR Join Forces in Web Environment}

\author{Old\v{r}ich Krůza and Nino Peterek}
\institute{Charles University in Prague\\
           Faculty of Mathematics and Physics\\
           Institute of Formal and Applied Linguistics\\
           Malostranské nám. 25, Prague, Czech Republic\\
           kruza@ufal.mff.cuni.cz, peterek@ufal.mff.cuni.cz}

\maketitle

\begin{abstract}

The paper presents a system for combining human transcriptions with automated
speech recognition to create a quality transcription of a large corpus in good
time. The system uses the web as interface for playing back audio, displaying
the automatically-acquired transcription synchronously, and enabling the visitor
to correct errors in the transcription. The human-submitted corrections are then
used in the statistical ASR to improve the acoustic as well as language model
and re-generate the bulk of transcription. The system is currently under
development.\\
The paper presents the system design, the corpus processed as well as
considerations for using the system in other settings.

\end{abstract}

\section{Introduction}

For a setting where recorded speech data are available, it is often desirable to
obtain a transcribed version of the spoken words because processing digital text
is much simpler than processing digital audio. Human transcription is costly and
no automatically acquired transcription is perfect, so human revision of ASR
output is a common scenario.

In our setting, where an uncatalogized spoken corpus and a community of lay
volunteers are available, we feel a need for a system that would employ the
modern speech-recognition technology and thoroughly exploit the precious brain
cycles of the involved humans. The web with its ever-growing possibilities
offers a neat platform for creating such a system.

\subsection{Spoken corpus of Karel Mako\v{n}}

Karel Mako\v{n} (1912-1993) was a Czech mystic who authored 27 books on
spirituality and Christian symbolism, and translated and commented 28
others\footnote{http://www.makon.cz/}.
Aside from his writing, he was giving talks in a close, private circle of
friends as the topics he was covering were not safe to express openly under the
socialist regime. His friends and apprentices recorded most of his lectures on
magnetic tapes. The recording started in late 60's using reels, then switched to
cassettes in the course of the 70's and went on until Mako\v{n}
ceased to give lectures in 1992, one year before his death.

The recordings were archived in the homes of their makers, losing the quality
as the magnetic signal was slowly fading over time. Attempts have been made to
digitize the material, but using little to no automation only a
fraction has been converted, leaving the rest of the material at time's mercy. A
systematic digitization took place only from 2010 to 2012 when most of the material
was converted to wav files.

The complete corpus, as of the time of writing, comprises about 1000 hours and
is available on-line\footnote{http://www.makon.fm/}. Our work deals
with processing this corpus, although the system under development is intended to
work on any data and we hope it will be useful in a broader spectrum of
settings.

\section{Motivation for transcription}

The digitized corpus is largely unstructured. The only metadata we have, are
years and sequential numbers in case of the cassettes. For some reels, there are
handmade indexes listing the topics covered and the corresponding time positions on the
counter of the device. These are typed on paper and have been photographed. The
indexes, however, cover only 76 of the 1096 total files.

This is a similar setting to the Malach\cite{malach} project. Malach was
approached by doing automatic speech recognition and building a search engine on
the ASR output. Our setting differs from that of Malach significantly, as we
merely have to deal with material by one single speaker, and we just have to handle about one thousand hours
in comparison with Malach's stunning one hundred sixteen thousand hours. This
plus the fact that there is an active community around the legacy of Karel
Mako\v{n} motivates us to attempt a full, high-quality, verbatim transcription
of Mako\v{n}'s talks.

\section{Architecture of the system}

Obviously, there are two basic ways to do speech-to-text transcription:
\begin{itemize}
\item{manually,}
\item{automatically.}
\end{itemize}
Manual transcription usually delivers very high quality, but it is unbearably
time-consuming. Automatic transcription can be quite fast, but the amount of
errors leaves much to be desired. Hence, we propose and are developing a system
to draw on the power of both to converge quickly to a good transcription.

\begin{figure}[htpb]
\includegraphics[scale=0.6]{arch.eps}
\caption{Schema of the system architecture}
\label{fig:arch}
\end{figure}

Figure~\ref{fig:arch} outlines the system architecture:
\begin{enumerate}
\item{An initial transcription is obtained by means of an automatic speech recognition
system.\label{enum:arch:init}}
\item{The resulting transcription is transformed into a JavaScript data
structure suitable for synchronous displaying on the web
interface.\label{enum:arch:htk2sub}}
\item{Visitors use the web interface to listen to the recordings,
at the same time seeing the synchronized transcription.}
\item{When encountering an error in the transcription, the visitor selects the
erroneous text and edits it, entering the correct transcription.}
\item{The corrected subtitle is immediately saved, so that it is shown on
subsequent requests of that passage.}
\item{The corrected subtitle is added to the training data of the ASR system.}
\item{The ASR model is re-trained with the new training data.}
\item{The re-trained ASR model is applied and the bulk of the transcription that
has not yet been human-corrected, is re-generated.}
\end{enumerate}

\subsection{First transcription}

We did some first experiments to obtain the initial transcription. We tried
to train an acoustic model on 15 minutes of manually transcribed material from
one of the recordings. The training was executed using HTK\cite{htk}. We also
adapted the acoustic model from CUCFN\cite{cucfn} to Makoň's speech using the
mentioned 15 minutes of manual transcription.

For language modeling, we trained a bigram model on Karel Makoň's books and one
on PDT\cite{pdt}. We're searching for a fitting ASR configuration to use the
models. The current results on ASR for Czech\cite{czasr} give us certain
expectations with respect to accuracy.

\subsection{Web interface}

The web interface where people can collaborate on speech transcription is a
cornerstone of the whole project. The audio is played back by browser's native
HTML5 \texttt{\textless{}audio\textgreater} element, falling back to
flash\footnote{http://www.adobe.com/products/flashplayer.html} where HTML5 is
not supported\footnote{jPlayer was used for this; http://jplayer.org/}. The
subtitles are shown as three lines of text with the current word being
highlighted. The lines scroll down as the recording is played back. This was
chosen for ergonomic reasons. Two other possible implementations that came into
mind were rejected:
\begin{enumerate}
\item{a horizontally-scrolling line of text, like a
\texttt{\textless{}marquee\textgreater} HTML element,\label{enum:subs:marquee}}
\item{subtitles like in a movie: one phrase shown at a time, which disappears
and is replaced by a new one.\label{enum:subs:srt}}
\end{enumerate}
Using a marquee-like solution (point \ref{enum:subs:marquee}) would cause
constant scrolling in a non-constant speed, since the rate of the speech varies.
This would quickly become annoying.

Using a movie-subtitle-like solution (point \ref{enum:subs:srt}) has the
disadvantage of making it impossible to select (and thus correct) a span of
text reaching across the given phrases.

In our solution, several lines are shown, they are horizontally static, and the
current word is highlighted, always being kept on the middle line (unless the
beginning or end of the recording is being played back). This delivers the reader enough
context and feels similar to reading static text. A drawback of
this method is that showing many lines gets computationally intense. We wrap
each word into a dedicated HTML element to be able to keep track of the
currently played word and to be able to track down the corresponding subtitles
when the user selects and corrects some text chunk. A good trade-off seems to be using
three lines. There is always enough context, it is never hard to spot the
highlighted word, and the browser stays responsive.

The highlighting of the current word still has a small difficulty. Since the
time-update event, which triggers re-drawing of the highlighted word, is quite
coarsely granular -- i.e. it fires only about 4 times per second --, it happens that
a word is highlighted with a noticeable delay. This reduces the comfort slightly
but can be avoided using look-ahead methods.

As hinted at in point \ref{enum:arch:htk2sub} of the architecture outline, the
subtitles are stored in a JavaScript structure. To be more precise, we're
storing it in the JSONP\cite{jsonp} format.
This allows us to have the subtitles on an external
CDN\footnote{Content-delivery network} and dedicate the web server to processing
the corrections. Another obvious advantage is that the browser can parse the
subtitles rapidly, not losing responsivity for a long period.

\section{Dealing with submitted corrections}

Processing the submissions from the visitors is probably the most important point
of the work. Working collaboratively on large data over web is a very popular
approach, most notably exploited by
Wikipedia\footnote{http://www.wikipedia.org/}. Our setting shares some notions
with that of Wikipedia, but is obviously quite different. Not only because our
data are acoustic but mainly because in our case, there is one correct output
that we hope the users to provide, whereas in case of Wikipedia, the users are
creative. This makes our task much easier in respect of conflict resolution.
When two people edit the same Wikipedia article at the same time, an interactive
edit-conflict procedure has to be
employed\footnote{http://en.wikipedia.org/wiki/Help:Edit\_conflict}.

To explain our situation, we'll first outline what happens when a correction is
sent by a visitor.
\begin{enumerate}
\item{The visitor selects the subtitles where the error is contained. The
selection is automatically padded to whole words.}
\item{The selected subtitles are replaced by a textarea where the words are
copied.}
\item{When the textarea loses focus, and if it had changed, then the new text is
sent to the server using an
AJAX\cite{ajax} request. Along with the text, the timestamp of the first word
and of the one after the last word is sent.}
\item{On the server, the correction is saved into the database.}
\item{The corresponding audio sequence is extracted and matched against the
provided text.

If the forced alignment fails, the process ends
and no further modifications take place; an error response is sent to the
client.

If, on the other hand, the forced alignment of the provided subtitles to
the audio succeeds, then the individual words get timestamps
and the process continues.\label{enum:corr:align}}
\item{The corrected span of subtitles is replaced with the submitted aligned
text and the transcription is saved.}
\item{The submitted transcription is added to training data for the acoustic
model.}
\item{The submitted transcription is added to the training data for the language
model.}
\item{Re-training of the model is scheduled.\footnote{Actually re-training the
model and re-recognizing the whole corpus after every submissions is, of course,
infeasible.}}
\end{enumerate}

Notice that % 1) all submissions are saved into the database and 2)
if two people
send overlapping corrections, they are simply pasted onto each other. The
intersection is saved from the latter of the concurring submissions.
Assuming they were both correct, there is no conflict.

If we give up the assumption of the correctness of the submissions, we have two
ways of dealing with that. Firstly, notice point \ref{enum:corr:align} of the
correction sequence: If the alignment fails, then the subtitle is not saved.
This protects the system from outright vandalism and spam, since transcriptions
that are too far from the audio will be automatically rejected.

Other, more subtle forms of errors introduced by the visitors will naturally be
harder to compensate. One thing we might do, is to track which correction comes from
which user. If a user tends to consistently make a certain kind of mistake, it
may be possible to revert it, or to set up a group of proof-readers. This is
pure theory though. We'll have to deal with these problems when they come.

We were making a simplification up to this point: Upon submission of a corrected
subtitle, we cannot just add this to the training data. We also have to remove
previous corrections of the same passage, so that previously introduced errors
can be compensated.

\subsection{Foreign words}

Another scenario that brings in complications is, when foreign words or any
words with non-standard pronunciation are a part of the corrected subtitle.
Suppose that the name \emph{George} appears in the audio and is not correctly
recognized. Also suppose the name is not present in the dictionary of the speech
recognition system. Since the mechanism for accepting or rejecting a submission
is based on acoustic fit and we use a simple rule-based word-to-phonemes
converter\footnote{We use phonetic alphapet designed by Psutka et al. 1997
\cite{czphabc} based on Czech phonology as described by Palková 1992 \cite{czphon}},
if the user enters \emph{George}, the word-to-phonemes
converter will assume pronunciation \textipa{/gEorgE/} instead of the correct
\textipa{/dZO:\*rdZ/} -- and thus will likely be rejected as non-matching.

We approach the problem by instructing the users to type in foreign words in
their phonetic transcription. In this case, \emph{džordž}. After the subtitle is
saved, the words become data structures with independent word forms and
pronunciations. Then, the proper spelling \emph{George} can be introduced.

%\subsection{Selecting data to correct}
%
%One interesting aspect of our setting is that we can influence which part of the
%corpus will be subject to corrections. We can recommend to visitors, which parts
%they should listen to right now.
%
%With enough training data, the system reaches a state where more training data
%no longer increases the accuracy\cite{
%    % TODO: někoho citovat
%}. Therefore, it is good to choose such samples for manual transcription that
%will globally maximize the precision of the model in the point of training-data
%saturation. This set can be to a certain degree approximated\cite{
%    % TODO: citovat ty, co popisujou výběr vhodných trénovacích dat
%}. % TODO: popsat, jak se to přibližně dělá

\section{Using the system in other settings}

As much as our primary aim is to process this specific corpus, we hope the
functionality developed to be applicable in a wide range of settings. Applying
the system for a different corpus should be straight-forward. Complications
could arise if the corpus were stored in smaller files since we don't take care
of playing one file after another: we assume that they are separate and the user
is only ever interested in one specific file at a time.

Using the system for a corpus in another language would also make modifications
necessary. Mainly the word-to-phonemes conversion would have to be implemented
in a way that is suitable for the given language. Of course the whole ASR can be
replaced -- and we plan to do this to try out different ones.

One part that lends itself neatly for modifications and re-use is the web
interface. For data, where audio, or video for that matter, and aligned
transcription are available, the interface can be used as a generic subtitle
displayer.

In music, the audio-text synchronous display could be used for collaborative
transcription of lyrics or even for karaoke. One possible serious drawback in
this would be the difficult matching of lyrics to audio. The music mixed into
the words, artistic interpretation of the vocalist and a variety of speakers
(singers actually) would likely make the alignment much less robust. In a
setting though, where the aligned lyrics with potential errors are on input, the
system could still be well used to for display and gathering corrections.

In language teaching, students could use the system to match spoken words with
written ones to learn pronunciation of their language of interest. Introducing
deliberate errors and expecting corrections could be a way of examination.

\section{Conclusion}

The described system is under development. Much of the functionality
has already been implemented, although the current form of the user interface is
crude at best. Our work on this project has merely begun. We are thrilled to
develop a system that on one hand will systematize and make available maybe the
most comprehensive opus on Christian mystic in Czech, on the other hand will
open new possibilities in collaborative transcription of speech.

\section*{Acknowledgments}

The research was supported by SVV project number 265 314.\\
\\
This work has been using language resources stored
by the LINDAT-Clarin project of the Ministry of
Education of the Czech Republic (project LM2010013).

\bibliographystyle{splncs}

\begin{thebibliography}{}

\bibitem{malach}
W Byrne, D Doermann, M Franz, S Gustman, J Hajic, D Oard, M Picheny, J Psutka, B
Ramabhadran, D Soergel, T Ward, Wei-Jing Zhu
\emph{Automatic recognition of spontaneous speech for access to multilingual
oral history archives}
in Proc. Eurospeech 2003, Geneva, Switzerland
http://www.ee.umd.edu/~oard/pdf/tsap04.pdf

\bibitem{htk}
S Young, G Evermann, M Gales, T Hain, D Kershaw, X A Liu, G Moore, J Odell, D
Ollason, D Povey, V Valtchev, P Woodland
\emph{The HTK Book}
2006\\
http://htk.eng.cam.ac.uk/

\bibitem{cucfn}
W Byrne, J Hajič, P Ircing, F Jelinek, S Khudanpur, J McDonough, N Peterek
and J Psutka
\emph{Large Vocabulary Speech Recognition for Read and Broadcast Czech}
Lecture Notes in Computer Science; Springer Berlin / Heidelberg; 1999\\
http://www.springerlink.com/content/8f82hg9y13tylthc/fulltext.pdf

\bibitem{pdt}
A B\"{o}hmová, J Haji\v{c}, E Haji\v{c}ová, B Hladká
\emph{The Prague Dependency Treebank: A Three-Level Annotation Scenario}
2007\\
http://www.scientificcommons.org/43211198

\bibitem{czasr}
J Psutka, J Hajic, W Byrne
\emph{The development of ASR for Slavic languages in the MALACH project}
in Proc. ICASSP 2004\\
http://svr-www.eng.cam.ac.uk/~wjb31/ppubs/icassp04-malach-final.pdf

\bibitem{jsonp}
B Ippolito
\emph{Remote JSON - JSONP}
2005\\
http://bob.pythonmac.org/archives/2005/12/05/remote-json-jsonp/

\bibitem{ajax}
J J Garrett
\emph{Ajax: A new approach to web applications}
2005\\
http://mmccauley.net/com601/material/ajax\_garrett.pdf

\bibitem{czphabc}
J Nouza, J Psutka, J Uhlíř
\emph{Phonetic alphabet for speech recognition of Czech}
Radioengineering Vol.6, pp. 16-20, 1997\\
http://www.radioeng.cz/fulltexts/1997/97\_04\_04.pdf

\bibitem{czphon}
Z Palková
\emph{Fonetika a fonologie Češtiny}
Univerzita Karlova, Praha 1992

\end{thebibliography}
%\bibliography{paper}
\end{document}

### 2020-04-04 2020-tsd-brno-svolocz/svolocz.tex
add svolocz paper

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{rotating}

\usepackage{url}
\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Czech parliament meeting recordings as ASR training data}

% a short form should be given in case it is too long for the running head
\titlerunning{Czech parliament meeting recordings as ASR training data}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Jan Oldřich Krůza}
%
\authorrunning{Jan Oldřich Krůza}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Institute of Formal and Applied Linguistics,\\
Faculty of Mathematics and Physics,\\
Charles University
\mailsa\\
\url{ufal.mff.cuni.cz}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Czech parliament meeting recordings as ASR training data}
\tocauthor{Jan Oldřich Krůza}
\maketitle

\begin{abstract}
I present a way to leverage the stenographed recordings of the Czech parliament
meetings for purposes of training a speech-to-text system. The article presents
a method for scraping the data, acquiring word-level alignment and selecting
reliable parts of the imprecise transcript. Finally, I present an ASR system
trained on these data.
\keywords{speech corpus, speech recognition, czech}
\end{abstract}

\section{Introduction}

Training data for speech recognition is always a demanded commodity, especially
if it is free. There are for sure already some free Czech corpora fit for speech
recognition training:
\begin{itemize}
\item{
    Vystadial\cite{vystadialarticle} with its 77 hours of VoIP
    calls\cite{vystadialdata}, 
}
\item{
    The Prague Database of Spoken Czech\cite{pdtscarticle} with its 122 hours
    of richly annotated spontaneous dialogues\cite{pdtscdata},
}
\item{
    The Czech Senior COMPANION Expressive Speech Corpus with its 5 hours
    of professionally spoken utterances by a single speaker\cite{companiondata},
}
\item{
    Otázky Václava Moravce: 35 hours of transcribed recordings of the
    Czech TV talk show\cite{ovmdata},
}
\item{
    STAZKA, a set of speech recording from vehicles with its 35 hours of
    background noise and utterances\cite{stazkadata},
}
\item{
    Spoken Corpus of Karel Makoň\cite{kruuza2012making} with its 70 hours of
    manually transcribed spontaneous speech by a single speaker\cite{makondata},
}
\item{and possibly others that I am not aware of.}
\end{itemize}

The Czech parliament meeting recordings represent a publicly available dataset
of high-quality audio recordings of contemporary Czech in consistent low-noise
audio quality worth almost 4000 hours of downloadable material, about 2800 hours
with the overlaps subtracted. Extracting
training data for speech recognition systems would provide a corpus at least
one order greater in length than those so far publicly available.

Verily, I am not the first person to attempt using these recordings for speech
recognition. The Department of Cybernetics of University of West Bohemia
developed an automatic online subtitling system for the meetings in
2006\cite{pspsubs} and as a result, an 88-hour subset annotated by high-quality
automatic transcript has been released for speech recognition training
purposes\cite{pspdata}.

I attempt to use the official stenographic transcripts available for all the
talks so that it can be a new entry in the above list, on par in quality and
excelling in size.

\section{Data Preparation}

Since the source data is publicly available and in the public domain, I merely
provide the scripts for downloading and building the corpus. The algorithms and
parameters used are described in this section.

\subsection{Scraping}

Regrettably, the data are to my best knowledge only available in human-readable
form. The transcript is not clearly distinguished in the markup and is
interlaced with metainformation. My method of isolating the transcript is quite
crude but it covers the vast majority of cases. The criterion is to extract the
subtree of all nodes with HTML attribute \texttt{[align=justify]}, except HTML
elements \texttt{<b>}, which contain speaker identification.

The known shortcomings of this method are that 1) it discards the speaker
annotations, although it is valuable metainformation and 2) it skips some short
passages, e.g. references to other meetings, as can be seen in the meeting
from Feb. 12th 2020 10:10 -
10:20\footnote{https://www.psp.cz/eknih/2017ps/stenprot/040schuz/s040372.htm}.

\subsection{Alignment}

One of the obstacles in using the stenographic transcripts for training an ASR
system is the very loose alignment available. The recordings are all 14 minutes
long and have a 4-minute overlap. The corresponding transcript is thus aligned
in 10-minute blocks with a roughly 2-minute padding on each side of the audio.
Figure~\ref{fig:overlap} schematically shows the alignment of the stenographic
transcript to the audio and the overlap of the recordings.

\begin{figure}[htpb]
\includegraphics[scale=0.25]{rc/overlap.eps}
\caption{Alignment and overlap of audio files and transcript. The examples are
from Feb. 12th 2020 around 10 o'clock. The transcript corresponding to the
recording in the upper left covers audio positions 01:34 - 11:24. The one in the
lower right from 01:24 to 12:00.}
\label{fig:overlap}
\end{figure}

Systems for aligning long audio segments to their transcripts already exist,
like that of Moreno et al. 1998\cite{moreno1998recursive} or Hazen
2006\cite{hazen2006automatic}. They are both based on an already existing
automatically acquired transcript. I use this technique as well, though much
simplified.

I have used the dataset mentioned above\cite{pspdata} to train a GMM-based ASR
system, using the stenographs as training data for a language model. Using these
models, a word-level-aligned transcript of the whole set of recordings has been
acquired.

The predicted transcript and the stenographic one have then been compared for
Levenshtein distance, determining the edit operations needed to transform one
into the other. For each predicted word, an un-reliability score is then
computed as the number of edit operations taken on it divided by its length.
A reliability score is actually used, as 1 - un-reliability.
Figure~\ref{fig:align} shows how the stenographic transcript is aligned with the
audio on word level.

\begin{figure}[htpb]
\includegraphics[scale=0.4]{rc/align.eps}
\caption{Schema of aligning the audio to the stenographic transcript on word
level.}
\label{fig:align}
\end{figure}

\subsection{Audio Segmentation}

To create a usable dataset for training a speech-to-text system, it is not
necessary to perfectly align the whole transcript. It is fine to just select the
reliable segments.

The criteria for good training samples are:
\begin{enumerate}
\item{100\% precise transcript,}
\item{roughly sentence-level length,}
\item{consistent length.}
\end{enumerate}

To ensure precise transcript, it is good to have the samples padded by some
silence, since the alignment may be a bit imprecise. The need to split at longer
pauses goes against the need to split at consistent, none-too-great lengths.

The problem of selecting the optimal set of silences so that the longest ones
are kept and split the recording in chunks of set length boundaries looks like
one for dynamic programming but I use a simpler approach: Start with a set of
all silences predicted by the forced alignment. Iterate over the silences
shortest-first and remove it if it doesn't break the constraints.

I have experimentally set the length boundaries to 12 - 30 seconds. The maximum
length could be decreased at the cost of more frequent splits in the middle of a
word.

\subsection{Training Samples Selection}

With the audio segmented and corresponding manual transcripts extracted, the
last step remaining is selecting which segments to include in the traning data.
Indeed, since the recordings have a 2-minute padding on each side for 10 middle
minutes, we must discard at the very least 40\% of the segments. I use the
following criteria for including a segment in the data:

\begin{enumerate}
\item{The first and last token have reliability at least 50\%,}
\item{The mean reliability of all tokens is at least 70\%,}
\item{The number of words is no less than five.}
\end{enumerate}

Minimum reliability of border tokens is considered to minimize the danger of
false alignment. Mean reliability is considered because it is OK for some
words to have reliability zero: there are enough errors in the prediction,
that's why we use the manual transcript after all. But if too many tokens have
too low reliability, then it is a sign of a suspicious segment. The number of
words has a minimum because with only a few words, the probability of
misalignment with good score is much greater than when there are enough words.

Why use mean reliability and not median? The way the reliability is computed,
considers the number of edit operations on one line in the automatic transcript.
In the case where there are many insertions, the reliability of one line can go
arbitrarily deep sub zero. So it can happen that there are several lines
extra words in the (mis)aligned chunk. The mean taps these.

\subsection{Data Extraction Summary}

All the constants and criteria are to be considered a baseline solution. They
all could be tweaked much more rigorously and solved much more soundly. However,
this simple solution readily yields a high-quality training dataset of 1058
hours. Of the total 542,610 segments, 166,458 (31\%) have been accepted to the
training dataset. Of the total 376,152 discarded segments, 316,030 (84\%) were
discarded because of the criterion of unreliable start or end.

Setting the minimum reliability of the boundary words from 50\% to 70\% reduces
the number of accepted chunks to by 17\%. It strips 5\% segments
of the total number from the dataset. But if we consider that 40\% of the total
number of segments must be discarded because of audio padding, the loss is
acually 9\%. It is an option to raise the matching reliability with a cost of
training data volume.

\section{ASR Based on the Dataset}

I have trained a standard DeepSpeech\cite{hannun2014deep} model with training :
dev : test ratio of 48 : 1 : 1; batch size 50; learning rate 0.0001. The
training took TODO epochs and the final WER on testing data from the corpus
itself was TODO.

\section{Conclusion}

I have presented a new corpus of spoken Czech suitable for training speech
recognition systems based on data in the public domain. The corpus size exceed
by an order the size of other freely available such corpora.

\section*{Acknowledgments}

This work has been using language resources developed, stored and distributed by
the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
Czech Republic (project LM2018101).

\bibliographystyle{splncs}

\bibliography{citace}

\end{document}

### 2020-04-07 2020-tsd-brno-svolocz/svolocz.tex
new minimum start/end reliability 70%

@@
 \begin{enumerate}
-\item{The first and last token have reliability at least 50\%,}
+\item{The first and last token have reliability at least 70\%,}
 \item{The mean reliability of all tokens is at least 70\%,}
@@
 this simple solution readily yields a high-quality training dataset of 1058
-hours. Of the total 542,610 segments, 166,458 (31\%) have been accepted to the
-training dataset. Of the total 376,152 discarded segments, 316,030 (84\%) were
+hours. Of the total 539,057 segments, 142,530 (26\%) have been accepted to the
+training dataset. Of the total 396,527 discarded segments, 350,258 (88\%) were
 discarded because of the criterion of unreliable start or end.
 
-Setting the minimum reliability of the boundary words from 50\% to 70\% reduces
-the number of accepted chunks to by 17\%. It strips 5\% segments
-of the total number from the dataset. But if we consider that 40\% of the total
-number of segments must be discarded because of audio padding, the loss is
-acually 9\%. It is an option to raise the matching reliability with a cost of
-training data volume.
+Reducing the minimum reliability of the boundary words from 70\% to 50\% increases
+the number of accepted chunks by 17\%. It adds 5\% segments
+of the total number to the dataset. But if we consider that 40\% of the total
+number of segments must be discarded because of audio padding, the gain is
+acually 9\%. It is an option to increase the training data volume at the cost
+of matching precision.
 

### 2020-04-12 2020-tsd-brno-denoise/denoise.tex
add denoise tsd 2020 paper

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{rotating}

\usepackage{url}
\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Mitigating Low Audio Quality by Domain Transfer for Better ASR}

% a short form should be given in case it is too long for the running head
\titlerunning{Audio Domain Transfer for Better ASR}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Jan Oldřich Krůza}
%
\authorrunning{Jan Oldřich Krůza}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Institute of Formal and Applied Linguistics,\\
Faculty of Mathematics and Physics,\\
Charles University
\mailsa\\
\url{ufal.mff.cuni.cz}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Mitigating Low Audio Quality by Domain Transfer for Better ASR}
\tocauthor{Jan Oldřich Krůza}
\maketitle

\begin{abstract}
The paper presents an attempt to solve the problem of bad audio quality
affecting the accuracy of speech-to-text systems. I
describe an experiment with Gycle-GAN domain transfer from the defect to the
pure as a way to reduce the impact of noise, overdrive, reverberation and other
defects.
\keywords{speech recognition, GycleGAN, domain transfer, signal processing}
\end{abstract}

\section{Introduction}

Automatic speech recognition is generally approached by means of machine
learning: a statistical model is inferred based on some training data. Its
application bases on the assumption that whatever was learned about the training
data will also apply for the input data. It is clear then, that if the input
data have different properties than the training data, the basic assumption is
violated and the performance will suffer.
Also if the training data themselves are damaged in a way that makes the desired
information harder to extract, the learning is going to be less successful.

For speech recognition, the problem of noise, reverberation and other defects
has been a pain point and also a point of research. To mention some, Gillespie
and Atlas (2002)\cite{gillespie2002diversity} deal with reverberation for
GMM-based ASR, Yoshioka et al. (2012)\cite{reverbmagazine} summarize various
dereverberation techniques, Ko et al. (2017)\cite{reverbaugment} attempt to
simlulate reverberation in training data digitally. Seltzer, Yu and Wang
(2013)\cite{dnnnoiserobust} deal with noise in DNN-based ASR.

There are in principle two ways to deal with the negative effect of acoustic
deficiencies in speech recognition: 1) to adapt the model, which includes
training on augmented data, and 2) to adapt the input data. This article
presents an experiment in the latter approach.

\section{Spoken Corpus of Karel Makoň}

I am looking into the problem of acoustic inconsistencies in terms of varying
overall quality, reverberation, noise, overdrive and other phenomena in a
dataset that is to be automatically transcribed and listened to by humans.

The application is at the spoken corpus of Karel Makoň\cite{makondata}, for which
a speech recognition system has been developed\cite{kruuza2012making} but where
the heterogenous acoustic quality affects both the speech recognition
performance and the listening experience. Whereas the word error rate on the
overall test set is 21\%, it is 45\% on a sample of recordings suffering from
overdrive, and 68\% on recordings taken on a magnetophone tape with a slow
recording speed of 2 centimeters per second after some 30 years of living-room
storage.

I am conducting an experiment to make the input data similar to the training
data instead of the more usual other way around because 1) it is not easy to
make the volunteer annotators transcribe acoustically flawed data, 2) a
success would not only benefit speech recognition results but also listening
experience and 3) the model could stay leaner.

What acoustic defects are actually present? Since the corpus was digitized from
magnetophone tapes of spontaneous talks by a single speaker recorded in various
amateur conditions and with consumer devices, we can broadly distinguist these
types of interefering signal:
\begin{enumerate}
\item{additive noise -- hum or hiss,}
\item{
    stationary interference like screeching added by low-quality magnetophone
    parts or the erasing head signal,
}
\item{non-stationary interference like like background speech or door slams,}
\item{
    room echo or ill-equalized microphone boosting or cutting certain
    frequencies,
}
\item{non-linear distortion of the magnetophone}
\item{speed fluctuations.}
\end{enumerate}

The individual interference types affect each other and can occur several times
in several stages. For instance, echo is like a convolution with a certain
sample vector. It is followed by non-linear distortion and after that
convolution again with a different vector in the imperfect magnetophone
circuits. Additive noise comes on top of that. Another batch of distortions
comes during the playback. Clearly, modeling, let alone reverting these
interferences is a difficult task.

\section{Neural Domain Transfer}

The revolutionary article of Zhu et al. (2017)\cite{cyclegan} presenting
cycle-consistent generative adversarial networks gave mankind a mighty tool and
a hilarious tool that was used for de-hazing
photographs\cite{Engin_2018_CVPR_Workshops}, giving people other's facial
expressions\cite{jin2017faceoff}, in biomedicine\cite{yang2018biogan} and also
in speech processing: Kaneko and Tameoka (2017)\cite{kaneko2017parallel} present
speech domain transfer and Hosseini-Asl et al.
(2018)\cite{hosseini2018malevoicegan} do the same for the purpose of speech
recognition.

Most closely related to the work presented in this article is probably
SEGAN\cite{pascual2017segan}: CycleGAN employed to enhance speech.

CycleGAN assumes two datasets in consistent domains: day and night photos,
schematic and photographic maps, male and female speech etc. In this case,
we have a clear consistent domain of clean, high-quality recordings and the
rest, which suffers from any combination of a number of defects. The damaged
recordings are very unlike each other, they don't form a consistent domain.

There are two ways of dealing with this situation: 1) adapt CycleGAN so that it
can deal with a ``compact'' and a ``scattered'' domain or 2) cluster the
data and apply CycleGAN on the individual clusters. While adapting CycleGAN may be a
point of future work, I have explored the simpler way of clustering the data.

\section{Clustering}

In order to perform clustering on any dataset, a metric on the data is needed.
I have used that proposed by Mandel \& Ellis (2005)\cite{mandel2005song}, which
is based on mel-frequency cepstra. A distance matrix has been created using the
tool musly\cite{schnitzer2011using} and the clusters on top of that using
hierarchical clustering\cite{johnson1967hierarchical}.

A manual... or rather aureal check on the clusters confirms that files falling
into a cluster are indeed acoustically similar. So we arrived at a cluster of
overdriven recordings, a cluster of heavily hummed ones etc.

For curiosity, I looked at the distances of the files among each other and
compared it to that between other familiar sounds. For instance, two adjacent
segments of the same speaker from a meeting of the Czech parliament has a
distance of 1.4. Different speakers have a distance of 6.5. One of these
segments compared to a black metal song chorus has the distance of 519. On the
other hand, the median distance among the Spoken corpus of Karel Makoň, where
there is one single speaker, amounts to 55.9.

The experiment presented below is conducted on three clusters: 1) A cluster
of the cleanest recordings, used as the destination domain, 2) a cluster of
overdriven recordings and 3) one of those recorded with low tape speed. The
latter recordings are the hardest to understand also for humans. All clusters
have a maximum internal distance of 25.

\section{Results}

I have used the voice transfer method as proposed by Kaneko \& Kameoka
(2017)\cite{kaneko2017parallel} and implemented by
Mao\footnote{github.com/leimao/Voice\_Converter\_CycleGAN}. I have trained
transfer 1) between the clean and the overdriven cluster and 2) between the
clean and the low-speed cluster.

After 200 epochs, I have performed speech recognition on the transferred data.
The result is summed up in Table~\ref{tab:results}.

\begin{table}[htpb]
\begin{center}
\begin{tabular}{|l||r|r|}
\hline
           & original & transferred \\
\hline
overdriven & 0.450 & 0.441 \\
low-speed  & 0.685 & 0.939 \\
\hline
\end{tabular}
\caption{Word error rate for the two damaged clusters before and after CycleGAN
transfer.}\label{tab:results}
\end{center}
\end{table}

The small but apparent reduction in word error rate in case of the overdriven
cluster is the positive, promising result. The improvement is rather small, and
as such it does not seem worthwhile applying it on the entire corpus. It
proves, however, that there is some potential. An important immediate benefit is
that the transferred overdriven (or should I say de-overdriven) sound files are
much easier on the ear and it can be a help for the human listener.

The wrecking word error rate increase from 69\% to 94\% in case of the
recordings taken at low tape speed requires further investigation.
Figure~\ref{fig:plzen} shows the waveform and spectrum of a sample from the low-speed
cluster before and after the transfer. For comparison,
Figure~\ref{fig:overdrive} shows the same for the overdriven cluster.
Notice how parts of the signal are completely missing in the converted low-speed
sample. It seems that when the signal is too hard to separate from the noise,
the covertor prefers to generate silence.

\begin{figure}[htpb]
\includegraphics[scale=0.4]{rc/plzen.eps}
\caption{Wave form (above) and spectrogram (below) of a recording taken in low
tape speed before the CycleGAN transfer (left) and afterwards (right).}
\label{fig:plzen}
\end{figure}

\begin{figure}[htpb]
\includegraphics[scale=0.4]{rc/overdrive.eps}
\caption{Wave form (above) and spectrogram (below) of an overdriven recording
before the CycleGAN transfer (left) and afterwards (right).}
\label{fig:overdrive}
\end{figure}

%\subsection{Adapting CycleGAN for scattered domain}
%
%The way CycleGAN works given two datasets, one representing a domain $A$, the
%other representing a domain $B$, it learns two functions $f(A) -> B$ and
%$g(B) -> A$ that map the data between the two domains. In my case, where domain
%$B$ is not really a domain but more like a linear combination of several
%domains, the function $f$ that maps clean data to damaged data is not well
%defined. Indeed, how can we define a function that maps a clean recording to a
%damaged one when ``damaged'' can mean anything from overdrive, reverberation,
%noise, a frequency filter, and a number of other phenomena, and their
%combinations?
%
%A way could be to leverage the knowledge of the partially above listed possible
%effects that can turn a clean recording to a damaged one. If we defined
%elementary transformations simulating the individual ways a recording can be
%damaged, we could be looking for a linear combination of these transformations
%such that applying them on the clean sample $a \in A$ yields a result as similar
%to the damaged sample $b \in B$ as possible.
%
%Like this, we could make instead of $f(a \in A) -> B$ a function
%$f(a \in A, b \in B) -> B$ without the danger that it would degenerate to just
%returning the sample $b$.
%
%Aside of a metric for similarity between two sound files, we would need the list
%of transformations that take a substantial part in the acoustic damaging at
%hand.
%

\section{Conclusion}

Neural-network domain transfer offers a universally applicable method to
mitigate any kind of and combination of acoustic defects without the necessity
of modeling the specific effects.
The method was demonstrated to help to a small extent with speech recognition
results. However, where the damage is too severe, it discards the only
information at hand instead of improving its quality.

The benefit for human listeners seems quite promising although a quantitative
evaluation would be needed to assess it properly.

\section*{Acknowledgments}

This work has been using language resources developed, stored and distributed by
the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
Czech Republic (project LM2018101).

\bibliographystyle{splncs}

\bibliography{citace}

\end{document}

### 2020-04-12 2020-tsd-brno-denoise/denoise.tex
tsd 2020 denoise: typos

@@
 data have different properties than the training data, the basic assumption is
-violated and the performance will suffer.
+violated and the inference performance will suffer.
 Also if the training data themselves are damaged in a way that makes the desired
@@
 }
-\item{non-stationary interference like like background speech or door slams,}
+\item{non-stationary interference like background speech or door slams,}
 \item{
@@
 cycle-consistent generative adversarial networks gave mankind a mighty tool and
-a hilarious tool that was used for de-hazing
+a hilarious toy that was used for de-hazing
 photographs\cite{Engin_2018_CVPR_Workshops}, giving people other's facial

### 2020-04-12 2020-tsd-brno-denoise/denoise.tex
tsd 2020 denoise: remove years from inline citations

@@
 has been a pain point and also a point of research. To mention some, Gillespie
-and Atlas (2002)\cite{gillespie2002diversity} deal with reverberation for
-GMM-based ASR, Yoshioka et al. (2012)\cite{reverbmagazine} summarize various
-dereverberation techniques, Ko et al. (2017)\cite{reverbaugment} attempt to
-simlulate reverberation in training data digitally. Seltzer, Yu and Wang
-(2013)\cite{dnnnoiserobust} deal with noise in DNN-based ASR.
+and Atlas\cite{gillespie2002diversity} deal with reverberation for
+GMM-based ASR, Yoshioka et al.\cite{reverbmagazine} summarize various
+dereverberation techniques, Ko et al.\cite{reverbaugment} attempt to
+simlulate reverberation in training data digitally. Seltzer, Yu and
+Wang\cite{dnnnoiserobust} deal with noise in DNN-based ASR.
 
@@
 
-The revolutionary article of Zhu et al. (2017)\cite{cyclegan} presenting
+The revolutionary article of Zhu et al.\cite{cyclegan} presenting
 cycle-consistent generative adversarial networks gave mankind a mighty tool and
@@
 expressions\cite{jin2017faceoff}, in biomedicine\cite{yang2018biogan} and also
-in speech processing: Kaneko and Tameoka (2017)\cite{kaneko2017parallel} present
-speech domain transfer and Hosseini-Asl et al.
-(2018)\cite{hosseini2018malevoicegan} do the same for the purpose of speech
-recognition.
+in speech processing: Kaneko and Tameoka\cite{kaneko2017parallel} present
+speech domain transfer and Hosseini-Asl et al.\cite{hosseini2018malevoicegan} do
+the same for the purpose of speech recognition.
 
@@
 In order to perform clustering on any dataset, a metric on the data is needed.
-I have used that proposed by Mandel \& Ellis (2005)\cite{mandel2005song}, which
+I have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
 is based on mel-frequency cepstra. A distance matrix has been created using the
@@
 
-I have used the voice transfer method as proposed by Kaneko \& Kameoka
-(2017)\cite{kaneko2017parallel} and implemented by
+I have used the voice transfer method as proposed by Kaneko \&
+Kameoka\cite{kaneko2017parallel} and implemented by
 Mao\footnote{github.com/leimao/Voice\_Converter\_CycleGAN}. I have trained

### 2020-04-13 2020-tsd-brno-svolocz/svolocz.tex
proof reading of svolocz

@@
 Levenshtein distance, determining the edit operations needed to transform one
-into the other. For each predicted word, an un-reliability score is then
-computed as the number of edit operations taken on it divided by its length.
-A reliability score is actually used, as 1 - un-reliability.
+into the other. For each predicted word, a reliability score is then
+computed as 1 - unreliability where unreliability is the number of edit
+operations taken on it divided by its length.
 Figure~\ref{fig:align} shows how the stenographic transcript is aligned with the
@@
 are kept and split the recording in chunks of set length boundaries looks like
-one for dynamic programming but I use a simpler approach: Start with a set of
+one for dynamic programming but a simpler approach is also possible: Start with a set of
 all silences predicted by the forced alignment. Iterate over the silences
@@
 Minimum reliability of border tokens is considered to minimize the danger of
-false alignment. Mean reliability is considered because it is OK for some
+shifted alignment boundaries. Mean reliability is considered because it is OK for some
 words to have reliability zero: there are enough errors in the prediction,
@@
 
-Why use mean reliability and not median? The way the reliability is computed,
+Why use mean reliability and not median? The way the reliability is computed
 considers the number of edit operations on one line in the automatic transcript.
 In the case where there are many insertions, the reliability of one line can go
-arbitrarily deep sub zero. So it can happen that there are several lines
-extra words in the (mis)aligned chunk. The mean taps these.
+arbitrarily deep sub zero. So it can happen that there are several inserted
+words in a (mis)aligned chunk that only affect the reliability score of a single
+word. The mean taps these while the median doesn't.
 
@@
 
+Source code for scraping and building the corpus is in public domain and
+available on \texttt{GitHub.com/Sixtease/cz-parliament-speech-corpus}.
+
 \section*{Acknowledgments}

### 2020-04-15 2020-tsd-brno-denoise/denoise.tex
redaction from DKl

@@
 learning: a statistical model is inferred based on some training data. Its
-application bases on the assumption that whatever was learned about the training
-data will also apply for the input data. It is clear then, that if the input
+application is based on the assumption that whatever was learned about the training
+data will also apply to the input data. It is clear then, that if the input
 data have different properties than the training data, the basic assumption is
 violated and the inference performance will suffer.
-Also if the training data themselves are damaged in a way that makes the desired
+Also, if the training data themselves are damaged in a way that makes the desired
 information harder to extract, the learning is going to be less successful.
@@
 data instead of the more usual other way around because 1) it is not easy to
-make the volunteer annotators transcribe acoustically flawed data, 2) a
-success would not only benefit speech recognition results but also listening
-experience and 3) the model could stay leaner.
+make the volunteer annotators transcribe acoustically flawed data, 2) not only
+speech recognition would benefit from successful denoising, the listening
+experience would improve too, and 3) the model could stay leaner.
 
 What acoustic defects are actually present? Since the corpus was digitized from
-magnetophone tapes of spontaneous talks by a single speaker recorded in various
+magnetophone tapes of spontaneous talks of a single speaker recorded in various
 amateur conditions and with consumer devices, we can broadly distinguist these
-types of interefering signal:
+types of quality degradation:
 \begin{enumerate}
@@
 The individual interference types affect each other and can occur several times
-in several stages. For instance, echo is like a convolution with a certain
-sample vector. It is followed by non-linear distortion and after that
-convolution again with a different vector in the imperfect magnetophone
-circuits. Additive noise comes on top of that. Another batch of distortions
+in several stages. For instance, the echo is essentially convolution with a certain vector of
+amplitude values, known as the impulse response. It even might not be
+constant if there are moving objects in the room or if the speaker or
+microphone is moving. This convolution is followed by a non-linear
+distortion in the microphone and its preamplifier and convolved again with
+another impulse response, this time representing a frequency and phase
+response of recorder's amplifiers. Finally it is non-linearly distorted in
+the process of making a magnetic recording on the tape. Another batch of distortions
 comes during the playback. Clearly, modeling, let alone reverting these
@@
 other hand, the median distance among the Spoken corpus of Karel Makoň, where
-there is one single speaker, amounts to 55.9.
+there is a single speaker, amounts to 55.9.
 
-The experiment presented below is conducted on three clusters: 1) A cluster
-of the cleanest recordings, used as the destination domain, 2) a cluster of
-overdriven recordings and 3) one of those recorded with low tape speed. The
-latter recordings are the hardest to understand also for humans. All clusters
-have a maximum internal distance of 25.
+The experiment presented below is conducted on three clusters:
+\begin{enumerate}
+\item{A cluster of the cleanest recordings, used as the destination domain,}
+\item{a cluster of overdriven recordings and}
+\item{a cluster of recordings taken with low tape speed.}
+\end{enumerate}
+Recordings in cluster 3 are the hardest to understand even for humans. All these
+clusters have a maximum internal distance of 25.
 

### 2020-04-15 2020-tsd-brno-denoise/denoise.tex
s/I/we/

@@
 The paper presents an attempt to solve the problem of bad audio quality
-affecting the accuracy of speech-to-text systems. I
+affecting the accuracy of speech-to-text systems. We
 describe an experiment with Gycle-GAN domain transfer from the defect to the
@@
 
-I am looking into the problem of acoustic inconsistencies in terms of varying
+We are looking into the problem of acoustic inconsistencies in terms of varying
 overall quality, reverberation, noise, overdrive and other phenomena in a
@@
 
-I am conducting an experiment to make the input data similar to the training
+We are conducting an experiment to make the input data similar to the training
 data instead of the more usual other way around because 1) it is not easy to
@@
 data and apply CycleGAN on the individual clusters. While adapting CycleGAN may be a
-point of future work, I have explored the simpler way of clustering the data.
+point of future work, we have explored the simpler way of clustering the data.
 
@@
 In order to perform clustering on any dataset, a metric on the data is needed.
-I have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
+We have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
 is based on mel-frequency cepstra. A distance matrix has been created using the
@@
 
-For curiosity, I looked at the distances of the files among each other and
+For curiosity, we looked at the distances of the files among each other and
 compared it to that between other familiar sounds. For instance, two adjacent
@@
 
-I have used the voice transfer method as proposed by Kaneko \&
+We have used the voice transfer method as proposed by Kaneko \&
 Kameoka\cite{kaneko2017parallel} and implemented by
-Mao\footnote{github.com/leimao/Voice\_Converter\_CycleGAN}. I have trained
+Mao\footnote{github.com/leimao/Voice\_Converter\_CycleGAN}. We have trained
 transfer 1) between the clean and the overdriven cluster and 2) between the
@@
 
-After 200 epochs, I have performed speech recognition on the transferred data.
+After 200 epochs, we have performed speech recognition on the transferred data.
 The result is summed up in Table~\ref{tab:results}.
@@
 proves, however, that there is some potential. An important immediate benefit is
-that the transferred overdriven (or should I say de-overdriven) sound files are
-much easier on the ear and it can be a help for the human listener.
+that the transferred overdriven (or should we say de-overdriven) sound files are
+much easier on the ear and it can be a help for human listeners.
 

### 2020-04-24 2020-tsd-brno-svolocz/svolocz.tex
ASR results and proof reading for svolocz

@@
 \item{
-    Spoken Corpus of Karel Makoň\cite{kruuza2012making} with its 70 hours of
+    Spoken Corpus of Karel Makoň\cite{kruuza2012making} with its 100 hours of
     manually transcribed spontaneous speech by a single speaker\cite{makondata},
@@
 audio quality worth almost 4000 hours of downloadable material, about 2800 hours
-with the overlaps subtracted. Extracting
+after subtraction of the overlaps. Extracting
 training data for speech recognition systems would provide a corpus at least
@@
 Systems for aligning long audio segments to their transcripts already exist,
-like that of Moreno et al. 1998\cite{moreno1998recursive} or Hazen
-2006\cite{hazen2006automatic}. They are both based on an already existing
-automatically acquired transcript. I use this technique as well, though much
-simplified.
+like that of Moreno et al.\cite{moreno1998recursive} or
+Hazen\cite{hazen2006automatic}. They are both based on an already existing
+automatically acquired transcript. I use this technique as well, though 
+simplified and adapted to the task.
 
@@
 To create a usable dataset for training a speech-to-text system, it is not
-necessary to perfectly align the whole transcript. It is fine to just select the
-reliable segments.
+necessary to perfectly align the whole transcript. On the contrary, it is
+desirable to align what is reliably precise and discard the rest.
 
@@
 To ensure precise transcript, it is good to have the samples padded by some
-silence, since the alignment may be a bit imprecise. The need to split at longer
+silence, since the alignment obtained from the initial ASR may be a bit
+imprecise. We thus want to split at pauses, the longer the better, up to a
+certain limit (about 1 second). The need to split at longer
 pauses goes against the need to split at consistent, none-too-great lengths.
 
-The problem of selecting the optimal set of silences so that the longest ones
-are kept and split the recording in chunks of set length boundaries looks like
-one for dynamic programming but a simpler approach is also possible: Start with a set of
-all silences predicted by the forced alignment. Iterate over the silences
-shortest-first and remove it if it doesn't break the constraints.
+So the problem is to select an optimal set of silences so that the longest ones
+are used and so that they split the recording into chunks of length in a given
+range. This looks like a problem for dynamic programming but a simpler approach
+is also possible: Start with a set of all silences predicted by the forced
+alignment.  Iterate over the silences shortest-first and remove each if it
+doesn't break the constraints.
 
 I have experimentally set the length boundaries to 12 - 30 seconds. The maximum
-length could be decreased at the cost of more frequent splits in the middle of a
-word.
+length could be decreased at the cost of available pauses to choose from, which
+would lead to more frequent splits in the middle of a word.
 
@@
 Minimum reliability of border tokens is considered to minimize the danger of
-shifted alignment boundaries. Mean reliability is considered because it is OK for some
-words to have reliability zero: there are enough errors in the prediction,
-that's why we use the manual transcript after all. But if too many tokens have
-too low reliability, then it is a sign of a suspicious segment. The number of
-words has a minimum because with only a few words, the probability of
+shifted alignment boundaries. Mean reliability is considered because it is OK
+for some words to have very low reliability: there are enough errors in the
+prediction, that's why we use the manual transcript after all. But if too many
+tokens have too low reliability, then it is a sign of a suspicious segment. The
+number of words has a minimum because with only a few words, the probability of
 misalignment with good score is much greater than when there are enough words.
@@
 training dataset. Of the total 396,527 discarded segments, 350,258 (88\%) were
-discarded because of the criterion of unreliable start or end.
+discarded because of the criterion of unreliable start or end. It should be
+noted however, that the start / end reliability criterion is applied first, so
+it catches segments that would be discarded for other reasons also.
 
@@
 
-I have trained a standard DeepSpeech\cite{hannun2014deep} model with training :
-dev : test ratio of 48 : 1 : 1; batch size 50; learning rate 0.0001. The
-training took TODO epochs and the final WER on testing data from the corpus
-itself was TODO.
+I have trained a standard DeepSpeech\cite{hannun2014deep} model on the 1058
+hours with training :
+development : test ratio of 48 : 1 : 1; batch size 50; learning rate 0.0001; dropout
+rate 0.2. The training took 12 epochs to reach optimal dev fit and the final
+word error rate on testing data from the corpus itself is 8.40\%.
+
+The language model used was a pentagram model with pruned singleton trigrams,
+tetragrams and pentagrams. The bulk of scraped transcriptions, including those
+with no downloadable corresponding audio, was used as training data for the
+language model.
 
@@
 recognition systems based on data in the public domain. The corpus size exceed
-by an order the size of other freely available such corpora.
+by an order the size of other freely available such corpora. A speech
+recognition system with competitive performance was made to show the fitness
+of the dataset to the purpose.
 
-Source code for scraping and building the corpus is in public domain and
+Source code for scraping and building the corpus is in the public domain and
 available on \texttt{GitHub.com/Sixtease/cz-parliament-speech-corpus}.

### 2020-04-25 2020-tsd-brno-restructuring/splits-anonymous.tex
add cicling2019 republication

--- 2019-04-19 2019-cicling-la-rochelle/cardiff.tex
@@
-\documentclass{llncs}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
+%
+\documentclass[runningheads]{llncs}
+%
 \usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-\newtheorem{Definition}{Definition}
+%
 \title{Restructuring Spoken Corpus for Streaming Emulation}
-
-\author{Jan Old\v{r}ich Kr\r{u}za}
-%\author{Anonymous}
-
-\institute{Charles University in Prague\\
-           Faculty of Mathematics and Physics\\
-           Institute of Formal and Applied Linguistics\\
-           Malostransk\'{e} n\'{a}m\v{e}st\'{i} 25, Prague\\
-           Czech Republic\\
-           kruza@ufal.mff.cuni.cz}
-%\institute{Anonymous university\\
-%           Anonymous institute \\
-%           Address\\
-%           Country\\
-%           E-mail}
-
-\maketitle
+%
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{Anonymous Author\inst{1}}
+%
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Anonymous Institute,\\
+Anonymous University,\\
+Adress Undisclosed\\
+\email{anonymous@example.org}}
+%
+\maketitle              % typeset the header of the contribution
+%
 
@@
 
-In the following chapters, I shall list the original structuring for the Spoken corpus of Karel Makoň, provide reasons that led to that structuring, explain why a change was necessary, and describe the journey to realizing it.
+In the following chapters, I shall list the original structuring for the Spoken
+corpus of Karel Makoň, provide reasons that led to that structuring, explain why
+a change was necessary, and describe the journey to realizing it.
 
@@
 
-The algorithm sought seems to be a typical demonstration of dynamic programming:
-Find an optimal split of a part of the recording that only consists of the first
-word, then add next word and find the optimal split based on the current one.
-
-There is also a simpler variant, however:
-\begin{enumerate}
-\item{start with the set of all
-silences;}
-\item{iterate over them from shortest to longest;}
-\item{remove the silence
-from the set if joining its adjacent segments doesn't yield one that's longer
-than 60 seconds;}
-\item{end iteration;}
-\item{iterate again over chosen silences;}
-\item{remove a silence from the
-set if one of its adjacent segments is shorter than 30 seconds;}
-\item{end iteration;}
-\end{enumerate}
-
-This algorithm is very easy to program and it has linear complexity.
+The silences are selected by starting with the complete set and removing
+shortest first while constraints remain kept.
 
@@
 
-%
-%\section*{Bibliography}
-%    1. Reppen, Randi: Building a corpus: what are the key considerations?
-%in collection The Routledge handbook of corpus linguistics, pages 59-65, 2010, Routledge
-%    2. Crowdy: Spoken corpus design
-%in journal Literary and Linguistic Computing 1993/4, pages 259-265, 1993, Oxford University Press
-%    3. Hájek: Český mystik Karel Makoň
-%in journal Dingir 2007/4, ISSN 1212-1371, pages 142-143, ed. Vojtíšek, 2007, Dingir s.r.o.
-%    4. Ircing, Krbec, Hajič, Psutka, Khudanpur, Jelinek, Byrne: On large vocabulary continuous speech recognition of highly inflectional language-Czech
-%in proceedings of Seventh European Conference on Speech Communication and Technology, 2001
-%    5. Krůza, Peterek: Making Community and ASR Join Forces in Web Environment
-%in proceedings of International Conference on Text, Speech and Dialogue, pages 415-421, 2012, Springer
-%    6. Krůza, Kuboň: Second-Generation Web Interface to Correcting ASR Output
-%in Proceedings of the Future Technologies Conference (FTC) 2018, ISBN 978-3-030-02685-1, ISSN 2194-5357, volume 1, pages 749-762, 2018, Springer
-%    7. Psutka, Ircing, Psutka, Hajič, Byrne, Mírovský: Automatic transcription of Czech, Russian, and Slovak spontaneous speech in the MALACH project
-%in proceedings of Ninth European Conference on Speech Communication and Technology, ISBN 1018-4074, pages 1349-1352, 2005, ISCA
-%
-
-\section*{Acknowledgments}
-
-This work has been using language resources developed, stored and distributed by the LINDAT/CLARIN project of the Ministry of Education, Youth and Sports of the Czech Republic (project LM2010013).
-\\
-Participation in the conference was supported by COST action IS1404.
-\\
-This research was partially supported by SVV project number 260 104.
-
 \bibliographystyle{splncs}

### 2020-04-25 2020-tsd-brno-restructuring/splits.tex
add cicling2019 republication

--- 2020-04-25 2020-tsd-brno-restructuring/splits-anonymous.tex
@@
 %
-\author{Anonymous Author\inst{1}}
+\author{Jan Oldřich Krůza\inst{1}\orcidID{0000-0002-7510-6470}}
 %
+\authorrunning{Jan Krůza}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Anonymous Institute,\\
-Anonymous University,\\
-Adress Undisclosed\\
-\email{anonymous@example.org}}
+\institute{Institute of Formal and Applied Linguistics,\\
+Faculty of Mathematics and Physics,\\
+Charles University\\
+\email{kruza@ufal.mff.cuni.cz}}
 %
@@
 
+\section*{Acknowledgments}
+
+This work has been using language resources developed, stored and distributed by
+the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
+Czech Republic (project LM2018101).
+
 \bibliographystyle{splncs}

### 2020-04-25 2020-tsd-brno-svolocz/svolocz-anonymous.tex
svolocz: use samplepaper as template; add anonymized

--- 2020-04-24 2020-tsd-brno-svolocz/svolocz.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
 %
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+\documentclass[runningheads]{llncs}
 %
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
 %
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[runningheads,a4paper]{llncs}
-
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{rotating}
-
-\usepackage{url}
-\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-
-\mainmatter  % start of an individual contribution
-
-% first the title is needed
+%
 \title{Czech parliament meeting recordings as ASR training data}
-
-% a short form should be given in case it is too long for the running head
-\titlerunning{Czech parliament meeting recordings as ASR training data}
-
-% the name(s) of the author(s) follow(s) next
 %
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
 %
-\author{Jan Oldřich Krůza}
+\author{Anonymous Clerk\inst{1}\orcidID{0000-0000-0000-0000}}
 %
-\authorrunning{Jan Oldřich Krůza}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
-\institute{Institute of Formal and Applied Linguistics,\\
-Faculty of Mathematics and Physics,\\
-Charles University
-\mailsa\\
-\url{ufal.mff.cuni.cz}}
-
+\authorrunning{A. Clerk}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
 %
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
+\institute{The Ministry of Silly Walks,\\
+Monty Python,\\
+Hall of Fame\\
+\email{monty-python@example.org}}
 %
-
-\toctitle{Czech parliament meeting recordings as ASR training data}
-\tocauthor{Jan Oldřich Krůza}
-\maketitle
-
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}
@@
 
-\section*{Acknowledgments}
-
-This work has been using language resources developed, stored and distributed by
-the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
-Czech Republic (project LM2018101).
-
 \bibliographystyle{splncs}

### 2020-04-25 2020-tsd-brno-denoise/denoise-anonymous.tex
2020 denoise: use correct template and add anonymized version

--- 2020-04-15 2020-tsd-brno-denoise/denoise.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
 %
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+\documentclass[runningheads]{llncs}
 %
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
 %
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[runningheads,a4paper]{llncs}
-
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{rotating}
-
-\usepackage{url}
-\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-
-\mainmatter  % start of an individual contribution
-
-% first the title is needed
-\title{Mitigating Low Audio Quality by Domain Transfer for Better ASR}
-
-% a short form should be given in case it is too long for the running head
-\titlerunning{Audio Domain Transfer for Better ASR}
-
-% the name(s) of the author(s) follow(s) next
 %
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
+\title{Speech Enhancement for Low Quality Audio}
 %
-\author{Jan Oldřich Krůza}
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
 %
-\authorrunning{Jan Oldřich Krůza}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
-\institute{Institute of Formal and Applied Linguistics,\\
-Faculty of Mathematics and Physics,\\
-Charles University
-\mailsa\\
-\url{ufal.mff.cuni.cz}}
-
+\author{Rincewind\inst{1}\orcidID{0000-0000-0000-0000}
+\and Luggage\inst{1}\orcidID{0000-0000-0000-0001}}
 %
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
+\authorrunning{Rincewind \and Luggage}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
 %
-
-\toctitle{Mitigating Low Audio Quality by Domain Transfer for Better ASR}
-\tocauthor{Jan Oldřich Krůza}
-\maketitle
-
+\institute{Unseen University,\\
+Ankh-Morpork,\\
+The Discworld\\
+\email{\{rincewind,luggage\}@unseen-university.ankh-morpork.discworld}}
+%
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}
 The paper presents an attempt to solve the problem of bad audio quality
-affecting the accuracy of speech-to-text systems. We
-describe an experiment with Gycle-GAN domain transfer from the defect to the
+affecting the listening experience and accuracy of speech-to-text systems. We
+describe two experiments: one with automated spectral subtraction and one
+with Gycle-GAN domain transfer from the defect to the
 pure as a way to reduce the impact of noise, overdrive, reverberation and other
@@
 training on augmented data, and 2) to adapt the input data. This article
-presents an experiment in the latter approach.
+presents one experiment in the former and one in the latter approach.
 
@@
 performance and the listening experience. Whereas the word error rate on the
-overall test set is 21\%, it is 45\% on a sample of recordings suffering from
+overall test set is 19\%, it is 45\% on a sample of recordings suffering from
 overdrive, and 68\% on recordings taken on a magnetophone tape with a slow
@@
 in several stages. For instance, the echo is essentially convolution with a certain vector of
-amplitude values, known as the impulse response. It even might not be
+amplitude values, known as the impulse response. It might even not be
 constant if there are moving objects in the room or if the speaker or
@@
 response of recorder's amplifiers. Finally it is non-linearly distorted in
-the process of making a magnetic recording on the tape. Another batch of distortions
-comes during the playback. Clearly, modeling, let alone reverting these
+the process of making a magnetic recording on the tape. Another batch of
+distortions comes during the playback. On top of that, in every step a certain
+amount of additive noise is introduced. Clearly, modeling, let alone reverting these
 interferences is a difficult task.
+
+\section{Denoising}
+
+We conducted a baseline experiment employing the standard denoising based on
+spectral subtraction. We assumed that each recording in the
+corpus suffers from a consistent kind of noise. This does not hold for all the
+recordings but we dismissed this fact for the sake of simplicity. We used sox
+fox the denoising routine. The process for each recording is as follows:
+\begin{enumerate}
+\item{Identify and isolate a sample of pure stationary noise,}
+\item{extract the noise profile,}
+\item{apply noise reduction based on the noise profile.}
+\end{enumerate}
+
+Items 2 and 3 are taken care of by sox. As for extracting fitting silences, we
+used the following method:
+\begin{enumerate}
+\item{
+    Identify and extract all precicted silences using an
+    existing aligned automatic transcript.
+}
+\item{
+    Select 100 silences around the 25th percentile ordered by length.
+    This secures silences that are neither too short nor too long. Long silences
+    tend to contain non-stationary noises, that's why we avoid using them.
+}
+\item{
+    Create a distance matrix on the silences using
+    musly\cite{schnitzer2011using}.
+}
+\item{
+    Select 10 silences with the lowest median on the distance matrix. These
+    silences are most like other silences, thus being least likely to contain
+    non-speech events.
+}
+\item{
+    Finally, catenate and output the selected silences.
+}
+\end{enumerate}
+
+Subjective evaluation confirms the expectable outcome that relatively
+high-quality recordings that only suffer from some additive noise get
+easier to listen to. Recordings suffering from other defects and overall lower
+audio quality sometimes get rather harder to understand.
+
+We have also trained a speech recognition system on the denoised data. The
+performance has sunken to word error rate of 0.94, which basically means the
+transcription doesn't work at all. Probably with some parameter tweaking this
+could be remedied but it doesn't seem to be a promising path. Curiously enough, the
+original model trained on non-denoised data performs much better on the denoised
+test data with word error rate of 0.70. Table~\ref{tab:results-denoise} sums
+these up.
+
+\begin{table}[htpb]
+\begin{center}
+\begin{tabular}{|l||r|r|}
+\hline
+WER    & original model & denoised model \\
+\hline
+original testing data & 0.189 & 0.940 \\
+denoised testing data & 0.698 & 0.941 \\
+\hline
+\end{tabular}
+\caption{Word error rate for the original and denoised test data by a model trained
+on the original and denoised training data.}\label{tab:results-denoise}
+\end{center}
+\end{table}
 
@@
 We have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
-is based on mel-frequency cepstra. A distance matrix has been created using the
-tool musly\cite{schnitzer2011using} and the clusters on top of that using
+is based on mel-frequency cepstra. A distance matrix has been created using
+musly and the clusters on top of that using
 hierarchical clustering\cite{johnson1967hierarchical}.
@@
 
-The small but apparent reduction in word error rate in case of the overdriven
-cluster is the positive, promising result. The improvement is rather small, and
-as such it does not seem worthwhile applying it on the entire corpus. It
-proves, however, that there is some potential. An important immediate benefit is
+The word error rate decrease in the overdriven cluster is unfortunately statistically
+insignificant with standard deviations 0.054 and 0.079 of the transferred and
+original versions respectively. An important benefit is
 that the transferred overdriven (or should we say de-overdriven) sound files are
@@
 sample. It seems that when the signal is too hard to separate from the noise,
-the covertor prefers to generate silence.
+the convertor prefers to generate silence.
 
@@
 Neural-network domain transfer offers a universally applicable method to
-mitigate any kind of and combination of acoustic defects without the necessity
-of modeling the specific effects.
-The method was demonstrated to help to a small extent with speech recognition
-results. However, where the damage is too severe, it discards the only
+mitigate any kind and combination of acoustic defects without the necessity
+of specifically modeling them. However,
+we have not recorded any significant increase in speech recognition performance.
+Also, where the damage is too severe, the processing discards the only
 information at hand instead of improving its quality.
@@
 
+The conducted experiments on mitigating defects in acoustic quality of recorded speech
+show that 1) it is not easy to reach an increase in speech recognition
+performance, 2) it is easier to reach an improvement for the subjective
+human listening experience and 3) heavily impaired recordings are particularly difficult to
+make any better.
+
 \section*{Acknowledgments}
 
-This work has been using language resources developed, stored and distributed by
-the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
-Czech Republic (project LM2018101).
+This work has been supported by undisclosed institutions and entities.
 

### 2020-05-04 2020-tsd-brno-svolocz/svolocz-anonymous.tex
svolocz: add cross-corpora ASR

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 hours with training :
-development : test ratio of 48 : 1 : 1; batch size 50; learning rate 0.0001; dropout
+development : test ratio of 18 : 1 : 1; batch size 50; learning rate 0.0001; dropout
 rate 0.2. The training took 12 epochs to reach optimal dev fit and the final
@@
 
+I have also tried training a speech recognition system with other datasets and
+the combination of them all. Of the datasets listed in
+section~\ref{sec:intro}, only Vystadial, Otázky Václava Moravce (ovm) and the
+corpus of Karel Makoň (makon) proved useful without much effort.
+
+Apart from them, I used the publicly not available corpora of Charles University
+Corpus of Financial News (CUCFN, 65 hours)\cite{byrne1999large}, the Balanced
+corpus of informal spoken Czech (Oral2013, 293 hours)\cite{oral2013} and the
+spoken Bible (100 hours) available with no license terms from {\em
+poslouchamebibli.cz}. Table~\ref{tab:csasr:results} shows the speech recognition
+results for each corpus on test data from itself and on a common test set from
+all the corpora.
+
+\begin{table}[htpb]
+\begin{center}
+\begin{tabular}{|l||r|r|}
+\hline
+source    & WER on self & WER on all \\
+\hline
+bible     & 9.20\%  & 94.7\% \\
+cucfgn    & 31.6\%  & 72.8\% \\
+makon     & 30.4\%  & 77.3\% \\
+oral2013  & 78.4\%  & 60.7\% \\
+ovm       & 21.6\%  & 72.9\% \\
+\textbf{parliament}
+          & 8.74\%  & \textbf{39.7\%} \\
+vystadial & 51.0\%  & 74.0\% \\
+\hline
+all       & 28.4\%  & 28.4\% \\
+\hline
+\end{tabular}
+\caption{Word error rate of speech recognition on the individual corpora and on
+their concatenation.}\label{tab:csasr:results}
+\end{center}
+\end{table}
+
+All speech recognition systems were trained with the same hyperparameters as
+described above. For a fairer comparison, a common, general language model was
+used for all the tests. Hence the higher WER for the Parliament corpus tested on
+itself than the reported 8.40\%.
+
 \section{Conclusion}
@@
 I have presented a new corpus of spoken Czech suitable for training speech
-recognition systems based on data in the public domain. The corpus size exceed
+recognition systems based on data in the public domain. The corpus size exceeds
 by an order the size of other freely available such corpora. A speech
@@
 
+Among the compared corpora, the Czech parliament corpus performs by far best
+even in speech recognition outside its domain.
+
 Source code for scraping and building the corpus is in the public domain and
 available on \texttt{GitHub.com/Sixtease/cz-parliament-speech-corpus}.
+
+\section*{Acknowledgments}
+
+This work has been using language resources developed, stored and distributed by
+the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
+Czech Republic (project LM2018101).
 

### 2020-05-04 2020-tsd-brno-svolocz/svolocz.tex
svolocz: add cross-corpora ASR

@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
 %
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+\documentclass[runningheads]{llncs}
 %
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
 %
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[runningheads,a4paper]{llncs}
-
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{rotating}
-
-\usepackage{url}
-\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-
-\mainmatter  % start of an individual contribution
-
-% first the title is needed
+%
 \title{Czech parliament meeting recordings as ASR training data}
-
-% a short form should be given in case it is too long for the running head
-\titlerunning{Czech parliament meeting recordings as ASR training data}
-
-% the name(s) of the author(s) follow(s) next
 %
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
 %
-\author{Jan Oldřich Krůza}
+\author{Jan Oldřich Krůza\inst{1}\orcidID{0000-0002-7510-6470}}
 %
-\authorrunning{Jan Oldřich Krůza}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
+\authorrunning{Jan Krůza}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
 \institute{Institute of Formal and Applied Linguistics,\\
 Faculty of Mathematics and Physics,\\
-Charles University
-\mailsa\\
-\url{ufal.mff.cuni.cz}}
-
+Charles University\\
+\email{kruza@ufal.mff.cuni.cz}}
 %
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
+\maketitle              % typeset the header of the contribution
 %
-
-\toctitle{Czech parliament meeting recordings as ASR training data}
-\tocauthor{Jan Oldřich Krůza}
-\maketitle
-
 \begin{abstract}
@@
 of the dataset to the purpose.
+
+Among the compared corpora, the Czech parliament corpus performs by far best
+even in speech recognition outside its domain.
 

### 2020-06-24 2020-tsd-brno-denoise/denoise.tex
add DKl as author
DKl: additive noise note
things and stuff

@@
-%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.20 of 2017/10/04
 %
-% This is the LaTeX source for the instructions to authors using
-% the LaTeX document class 'llncs.cls' for contributions to
-% the Lecture Notes in Computer Sciences series.
-% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+\documentclass[runningheads]{llncs}
 %
-% It may be used as a template for your own input - copy it
-% to a new file with a new name and use it as the basis
-% for your article.
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
 %
-% NB: the document class 'llncs' has its own and detailed documentation, see
-% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[runningheads,a4paper]{llncs}
-
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{rotating}
-
-\usepackage{url}
-\urldef{\mailsa}\path|kruza@ufal.mff.cuni.cz|    
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+% If you use the hyperref package, please uncomment the following line
+% to display URLs in blue roman font according to Springer's eBook style:
+% \renewcommand\UrlFont{\color{blue}\rmfamily}
 
 \begin{document}
-
-\mainmatter  % start of an individual contribution
-
-% first the title is needed
-\title{Mitigating Low Audio Quality by Domain Transfer for Better ASR}
-
-% a short form should be given in case it is too long for the running head
-\titlerunning{Audio Domain Transfer for Better ASR}
-
-% the name(s) of the author(s) follow(s) next
 %
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
+\title{Speech Enhancement for Low Quality Audio}
 %
-\author{Jan Oldřich Krůza}
+%\titlerunning{Abbreviated paper title}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
 %
-\authorrunning{Jan Oldřich Krůza}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
+\author{Jan Oldřich Krůza\inst{1}\orcidID{0000-0002-7510-6470}\\
+\and David Klusáček\inst{1}\orcidID{0000-0002-0144-2092}}
+%
+\authorrunning{Krůza \and Klusáček}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
 \institute{Institute of Formal and Applied Linguistics,\\
 Faculty of Mathematics and Physics,\\
-Charles University
-\mailsa\\
-\url{ufal.mff.cuni.cz}}
-
+Charles University\\
+\email{\{kruza,klusacek\}@ufal.mff.cuni.cz}}
 %
-% NB: a more complex sample for affiliations and the mapping to the
-% corresponding authors can be found in the file "llncs.dem"
-% (search for the string "\mainmatter" where a contribution starts).
-% "llncs.dem" accompanies the document class "llncs.cls".
+\maketitle              % typeset the header of the contribution
 %
-
-\toctitle{Mitigating Low Audio Quality by Domain Transfer for Better ASR}
-\tocauthor{Jan Oldřich Krůza}
-\maketitle
-
 \begin{abstract}
 The paper presents an attempt to solve the problem of bad audio quality
-affecting the accuracy of speech-to-text systems. We
-describe an experiment with Gycle-GAN domain transfer from the defect to the
+affecting the listening experience and accuracy of speech-to-text systems. We
+describe two experiments: one with automated spectral subtraction and one
+with Gycle-GAN domain transfer from the defect to the
 pure as a way to reduce the impact of noise, overdrive, reverberation and other
@@
 evaluation would be needed to assess it properly.
+
+The conducted experiments on mitigating defects in acoustic quality of recorded speech
+show that 1) it is not easy to reach an increase in speech recognition
+performance, 2) it is easier to reach an improvement for the subjective
+human listening experience and 3) heavily impaired recordings are particularly difficult to
+make any better.
 
@@
 
+This research was supported by SVV project number 260 575.
+
 \bibliographystyle{splncs}

### 2020-06-24 2020-fedcsis-sofia-svolocz/svolocz.tex
add fedcsis svolocz paper

--- 2020-05-04 2020-tsd-brno-svolocz/svolocz.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
+% This file demonstrates how to use the IEEEConf LaTeX2e macro package,
+% to prepare a manuscript for proceedings on CD of the conference
+% FedCSIS
 %
-\documentclass[runningheads]{llncs}
+\documentclass[conference]{IEEEtran}
+%\documentclass[a4paper]{IEEEconf}
+
+% This package serves to balance the column lengths on the last page of the document.
+% please, insert \balance command in the left column of the last page
+\usepackage{balance}
+
+%% to enable \thank command
+\IEEEoverridecommandlockouts 
+%% The usage of the following packages is recommended
+%% to insert graphics
+\usepackage[dvips]{graphicx}
+% to typeset algorithms
+% to typeset code fragments
+\usepackage{listings}
+% provides various features to facilitate writing math formulas and to improve the typographical quality of their output.
+\usepackage[cmex10]{amsmath}
+\interdisplaylinepenalty=2500
+% por urls typesetting and breaking
+\usepackage{url}
+% for vertical merging table cells
+\usepackage{multirow}
+
+% define environments for remarks and examples
+\newtheorem{remark}{Remark}[section]
+\newtheorem{example}[remark]{Example}
+
 %
-\usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\begin{document}
 %
@@
 %
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
 %
-\author{Jan Oldřich Krůza\inst{1}\orcidID{0000-0002-7510-6470}}
-%
-\authorrunning{Jan Krůza}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Institute of Formal and Applied Linguistics,\\
+\author{
+\IEEEauthorblockN{Jan Oldřich Krůza}
+\IEEEauthorblockA{
+Institute of Formal and Applied Linguistics,\\
 Faculty of Mathematics and Physics,\\
 Charles University\\
-\email{kruza@ufal.mff.cuni.cz}}
+Email: kruza@ufal.mff.cuni.cz}
+}
+
+% conference papers do not typically use \thanks and this command
+% is locked out in conference mode. If really needed, such as for
+% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
+% after \documentclass
+
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
 % 
-\maketitle              % typeset the header of the contribution
-%
+%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
+%Homer Simpson\IEEEauthorrefmark{2},
+%James Kirk\IEEEauthorrefmark{3}, 
+%Montgomery Scott\IEEEauthorrefmark{3} and
+%Eldon Tyrell\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
+%Georgia Institute of Technology,
+%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212, Fax: (888) 555--1212}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
+
+\begin{document}
+\maketitle              % typeset the title of the contribution
+
 \begin{abstract}
@@
 reliable parts of the imprecise transcript. Finally, I present an ASR system
-trained on these data.
-\keywords{speech corpus, speech recognition, czech}
+trained on these and other data.
 \end{abstract}
@@
 
-\begin{figure}[htpb]
-\includegraphics[scale=0.25]{rc/overlap.eps}
+\begin{figure*}[tpb]
+\centering
+\includegraphics[width=0.9\hsize]{rc/overlap.eps}
 \caption{Alignment and overlap of audio files and transcript. The examples are
@@
 \label{fig:overlap}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[htpb]
-\includegraphics[scale=0.4]{rc/align.eps}
+\begin{figure*}[htpb]
+\includegraphics[width=0.9\hsize]{rc/align.eps}
 \caption{Schema of aligning the audio to the stenographic transcript on word
@@
 \label{fig:align}
-\end{figure}
+\end{figure*}
 
@@
 
+\section{Numerals and Abbreviations}
+
 \section{ASR Based on the Dataset}
@@
 corpus of informal spoken Czech (Oral2013, 293 hours)\cite{oral2013} and the
-spoken Bible (100 hours) available with no license terms from {\em
-poslouchamebibli.cz}. Table~\ref{tab:csasr:results} shows the speech recognition
+spoken Bible (100 hours) available with no license terms from
+\url{poslouchamebibli.cz}. Table~\ref{tab:csasr:results} shows the speech recognition
 results for each corpus on test data from itself and on a common test set from
@@
 \begin{table}[htpb]
-\begin{center}
+\caption{Word error rate of speech recognition on the individual corpora and on
+their concatenation.}
+\centering
 \begin{tabular}{|l||r|r|}
@@
 bible     & 9.20\%  & 94.7\% \\
-cucfgn    & 31.6\%  & 72.8\% \\
+cucfn     & 31.6\%  & 72.8\% \\
 makon     & 30.4\%  & 77.3\% \\
@@
 \end{tabular}
-\caption{Word error rate of speech recognition on the individual corpora and on
-their concatenation.}\label{tab:csasr:results}
-\end{center}
+\label{tab:csasr:results}
 \end{table}
@@
 Source code for scraping and building the corpus is in the public domain and
-available on \texttt{GitHub.com/Sixtease/cz-parliament-speech-corpus}.
+available on 
+\url{GitHub.com/Sixtease/cz-parliament-speech-corpus}.
 
@@
 
-\bibliographystyle{splncs}
+This research was supported by SVV project number 260 575.
 
+\bibliographystyle{myIEEEtran}
 \bibliography{citace}

### 2020-07-01 2020-fedcsis-sofia-svolocz/svolocz.tex
fedcsis svolocz: numerals

@@
 
+There are many numeral expressions in the transcripts. They amount to 489,880
+out of 25,010,269 tokens in the complete stenographic transcript, which is
+almost two percent. In the training dataset, 24\% of the samples contain one or
+more numerals.
+
+Originally, I have included the digits into the alphabet for speech recognition,
+thus attempting to train the system to transcribe numeral expressions directly
+into digits. The speech recognition system described in the following section
+would however transcribe numeral expressions as empty strings.
+
+There are four ways to deal with the problem:
+\begin{enumerate}
+\item{ignore it,}
+\item{remove digits from the training data,}
+\item{manually expand digits to words,}
+\item{automatically expand digits to words.}
+\end{enumerate}
+
+The first option needs no elaboration. The second one, removing samples with
+digits, is an easy and viable option but it is a waste of a quarter of the
+dataset and of the vast majority of samples with numerals in them. Manual
+expansion would surely be ideal but very costly. It remains to attempt the
+fourth variant of automated expansion.
+
+For automated expansion of digits into words, we can use the available initial
+transcript and the algorithm for alignment with the stenographic transcript.
+
+The expansion is done in two steps:
+\begin{enumerate}
+\item{generation of verbal variants,}
+\item{selection of the most likely variant.}
+\end{enumerate}
+
+I have used the Perl module \texttt{Lingua::CS::Num2Word} as a base for the
+expansion. I modified the module in the following way: 1) I added support
+the order of billions, which is very common in the corpus. 2) A number is no
+longer expanded into a single phrase but instead into all possible phrases
+expressing the given number. 3) I added support for genitive and accusative
+cases, decimal numerals, ordinals, dates and times.
+
+All tokens in the stenographs that include digits are expanded into their
+verbalization variants before further processing. Upon alignment, the variant
+with least edit distance from the initial transcript is selected.
+
+Common abbreviations and symbols are expanded together with the digits. For
+example, the very common character \textit{``§'' (paragraph)} is expanded into
+the forms \textit{paragraf, paragrafu, paragrafů, paragrafem,
+paragrafech} that represent common inflections of the word. Some common
+abbreviations that undergo inflection include \textit{``čl.'' (article)},
+\textit{``odst.''}, which also means ``paragraph'' and \textit{``tzv.''
+(co-called)}.
+
+With the expansion in the process, the similarity of the stenographic transcript
+and the initial one raised, which raised the number of accepted segments from
+26\% to 35\%. The amount of training data grew by 86 hours to 1144.
+
 \section{ASR Based on the Dataset}

### 2020-07-02 2020-fedcsis-sofia-svolocz/svolocz.tex
match thesis chapter

@@
 10:20\footnote{https://www.psp.cz/eknih/2017ps/stenprot/040schuz/s040372.htm}.
+Both can be corrected by devising a smarter scraper and neither has any
+significance for speech recognition: speaker annotation fundamentally and
+neglecting the links for their infrequency.
 
@@
 \end{figure*}
+
+Nota bene, a GMM-based system was chosen for the initial transcript instead of
+a DNN-based for three reasons: 1) Foremost, it is straightforward to obtain
+precise alignment from a GMM-based system. 2) The training doesn't require so
+much computational resources and data. 3) It isn't crucial to have maximum possible
+accuracy in this stage.
 
@@
 I have used the Perl module \texttt{Lingua::CS::Num2Word} as a base for the
-expansion. I modified the module in the following way: 1) I added support
+expansion. I modified the module in the following way: 1) I added support for
 the order of billions, which is very common in the corpus. 2) A number is no
@@
 abbreviations that undergo inflection include \textit{``čl.'' (article)},
-\textit{``odst.''}, which also means ``paragraph'' and \textit{``tzv.''
-(co-called)}.
+\textit{``odst.'' (also paragraph)} and \textit{``tzv.'' (co-called)}.
 
-With the expansion in the process, the similarity of the stenographic transcript
-and the initial one raised, which raised the number of accepted segments from
-26\% to 35\%. The amount of training data grew by 86 hours to 1144.
+After incorporating the expansion into the pipeline, the similarity of the
+stenographic transcript and the initial one raised, which also raised the number
+of accepted segments from 26\% to 35\%. The amount of training data grew by 86
+hours to 1144.
 
@@
 rate 0.2. The training took 12 epochs to reach optimal dev fit and the final
-word error rate on testing data from the corpus itself is 8.40\%.
+word error rate on testing data from the corpus itself is 8.40\% before digit
+expansion and 7.89\% afterwards.
 
@@
 ovm       & 21.6\%  & 72.9\% \\
-\textbf{parliament}
-          & 8.74\%  & \textbf{39.7\%} \\
+parliament w/digits
+          & 8.74\%  & 39.7\% \\
+\textbf{parliament expanded}
+          & 7.89\%  & \textbf{36.0\%} \\
 vystadial & 51.0\%  & 74.0\% \\
@@
 All speech recognition systems were trained with the same hyperparameters as
-described above. For a fairer comparison, a common, general language model was
-used for all the tests. Hence the higher WER for the Parliament corpus tested on
-itself than the reported 8.40\%.
+described above.% For a fairer comparison, a common, general language model was
+%used for all the tests. Hence the higher WER for the Parliament corpus tested on
+%itself than the reported 8.40\%.
 

### 2020-07-03 2020-fedcsis-sofia-denoise/denoise.tex
add fedcsis denoise and splits papers

--- 2020-06-24 2020-tsd-brno-denoise/denoise.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
+% This file demonstrates how to use the IEEEConf LaTeX2e macro package,
+% to prepare a manuscript for proceedings on CD of the conference
+% FedCSIS
 %
-\documentclass[runningheads]{llncs}
+\documentclass[conference]{IEEEtran}
+%\documentclass[a4paper]{IEEEconf}
+
+% This package serves to balance the column lengths on the last page of the document.
+% please, insert \balance command in the left column of the last page
+\usepackage{balance}
+
+%% to enable \thank command
+\IEEEoverridecommandlockouts 
+%% The usage of the following packages is recommended
+%% to insert graphics
+\usepackage[dvips]{graphicx}
+% to typeset algorithms
+% to typeset code fragments
+\usepackage{listings}
+% provides various features to facilitate writing math formulas and to improve the typographical quality of their output.
+\usepackage[cmex10]{amsmath}
+\interdisplaylinepenalty=2500
+% por urls typesetting and breaking
+\usepackage{url}
+% for vertical merging table cells
+\usepackage{multirow}
+\usepackage{flushend}
+
+% define environments for remarks and examples
+\newtheorem{remark}{Remark}[section]
+\newtheorem{example}[remark]{Example}
+
 %
-\usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-% \renewcommand\UrlFont{\color{blue}\rmfamily}
-
-\begin{document}
 %
@@
 %
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
 %
-\author{Jan Oldřich Krůza\inst{1}\orcidID{0000-0002-7510-6470}\\
-\and David Klusáček\inst{1}\orcidID{0000-0002-0144-2092}}
-%
-\authorrunning{Krůza \and Klusáček}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Institute of Formal and Applied Linguistics,\\
+\author{
+\IEEEauthorblockN{Jan Oldřich Krůza, David Klusáček}
+\IEEEauthorblockA{
+Institute of Formal and Applied Linguistics,\\
 Faculty of Mathematics and Physics,\\
 Charles University\\
-\email{\{kruza,klusacek\}@ufal.mff.cuni.cz}}
+Email: \{kruza,klusacek\}@ufal.mff.cuni.cz}
+}
+
+% conference papers do not typically use \thanks and this command
+% is locked out in conference mode. If really needed, such as for
+% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
+% after \documentclass
+
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
 % 
-\maketitle              % typeset the header of the contribution
-%
+%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
+%Homer Simpson\IEEEauthorrefmark{2},
+%James Kirk\IEEEauthorrefmark{3}, 
+%Montgomery Scott\IEEEauthorrefmark{3} and
+%Eldon Tyrell\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
+%Georgia Institute of Technology,
+%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212, Fax: (888) 555--1212}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
+
+\begin{document}
+\maketitle              % typeset the title of the contribution
+
 \begin{abstract}
@@
 defects.
-\keywords{speech recognition, GycleGAN, domain transfer, signal processing}
 \end{abstract}
@@
 \begin{table}[htpb]
-\begin{center}
+\caption{Word error rate for the original and denoised test data by a model trained
+on the original and denoised training data.}\label{tab:results-denoise}
+\centering
 \begin{tabular}{|l||r|r|}
@@
 \end{tabular}
-\caption{Word error rate for the original and denoised test data by a model trained
-on the original and denoised training data.}\label{tab:results-denoise}
-\end{center}
 \end{table}
@@
 \begin{table}[htpb]
-\begin{center}
+\caption{Word error rate for the two damaged clusters before and after CycleGAN
+transfer.}\label{tab:results}
+\centering
 \begin{tabular}{|l||r|r|}
@@
 \end{tabular}
-\caption{Word error rate for the two damaged clusters before and after CycleGAN
-transfer.}\label{tab:results}
-\end{center}
 \end{table}
@@
 
-\begin{figure}[htpb]
-\includegraphics[scale=0.4]{rc/plzen.eps}
+\begin{figure*}[tpb]
+\centering
+\includegraphics[width=0.9\hsize]{rc/plzen.eps}
 \caption{Wave form (above) and spectrogram (below) of a recording taken in low
@@
 \label{fig:plzen}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[htpb]
-\includegraphics[scale=0.4]{rc/overdrive.eps}
+\begin{figure*}[tpb]
+\centering
+\includegraphics[width=0.9\hsize]{rc/overdrive.eps}
 \caption{Wave form (above) and spectrogram (below) of an overdriven recording
@@
 \label{fig:overdrive}
-\end{figure}
+\end{figure*}
 
@@
 
-\bibliographystyle{splncs}
-
+\bibliographystyle{myIEEEtran}
 \bibliography{citace}

### 2020-07-03 2020-fedcsis-sofia-restructuring/splits.tex
add fedcsis denoise and splits papers

--- 2020-04-25 2020-tsd-brno-restructuring/splits.tex
@@
 %
-%\titlerunning{Abbreviated paper title}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
 %
-\author{Jan Oldřich Krůza\inst{1}\orcidID{0000-0002-7510-6470}}
-%
-\authorrunning{Jan Krůza}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Institute of Formal and Applied Linguistics,\\
+\author{
+\IEEEauthorblockN{Jan Oldřich Krůza}
+\IEEEauthorblockA{
+Institute of Formal and Applied Linguistics,\\
 Faculty of Mathematics and Physics,\\
 Charles University\\
-\email{kruza@ufal.mff.cuni.cz}}
+Email: kruza@ufal.mff.cuni.cz}
+}
+
+% conference papers do not typically use \thanks and this command
+% is locked out in conference mode. If really needed, such as for
+% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
+% after \documentclass
+
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
 % 
-\maketitle              % typeset the header of the contribution
-%
+%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
+%Homer Simpson\IEEEauthorrefmark{2},
+%James Kirk\IEEEauthorrefmark{3}, 
+%Montgomery Scott\IEEEauthorrefmark{3} and
+%Eldon Tyrell\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
+%Georgia Institute of Technology,
+%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212, Fax: (888) 555--1212}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
+
+\begin{document}
+\maketitle              % typeset the title of the contribution
 
@@
 the talks, Mr. Karel Makoň\cite{hajek2007cesky}, a Czech mystic, started giving talks in private
-groups after his deportation into a concentration camp in 1939, and stopped in
-1991, two years before his death. His regular listeners were recording his
+groups after his deportation into a concentration camp in 1939, and ceased in
+1992, one year before his death. His regular listeners were recording his
 talks. I don't know exactly when the recording started but the earliest denoted
@@
 There is also a three-hour video on youtube: \\
-\texttt{https://www.youtube.com/watch?v=UaNm9jnnJiA}
+\url{https://www.youtube.com/watch?v=UaNm9jnnJiA}
 
@@
 
-Option 3 - storing the corpus in smaller chunks has the downside of serving unnecessary context when requesting a specific range.
-Also, if we want to retain full download capability, we must either stitch the recording together or host both split and integral versions of all recordings.
+Option 3 - storing the corpus in smaller chunks has the downside of serving
+unnecessary context when requesting a specific range. Also, if we want to retain
+full download capability, we must either stitch the recording together or host
+both split and integral versions of all recordings.
 
-Option 4 - waiting for streaming support in Web Audio API might seem silly but the issue has been discussed by the Web Audio API developer team, see https://github.com/WebAudio/web-audio-api/issues/1305 .
+Option 4 - waiting for streaming support in Web Audio API might seem silly but
+the issue has been discussed by the Web Audio API developer team, see
+\url{https://github.com/WebAudio/web-audio-api/issues/1305}.
 
-I opted for storing the corpus in smaller chunks.
-The downside of unnecessary context is a minor one, especially if we choose a fitting length of the chunks.
-Having to host two versions of the corpus raises the costs in my case by less than one US Dollar per month.
+I opted for storing the corpus in smaller chunks. The downside of unnecessary
+context is a minor one, especially if we choose a fitting length of the chunks.
+Having to host two versions of the corpus raises the costs in my case by less
+than one US Dollar per month.
 
@@
 The corpus can be accessed through its web interface at
-\texttt{http://radio.makon.cz/}. The playback and other operations like playing
+\url{http://radio.makon.cz/}. The playback and other operations like playing
 back or downloading an arbitrary span (which can span a segment boundary) can be
@@
 GitHub: \\
-\texttt{https://github.com/Sixtease/CorpusMakoni/tree/master/scripts}.
+\url{https://github.com/Sixtease/CorpusMakoni/tree/master/scripts}.
 Similarly, the source code for the JavaScript part that works with the segments
 can be found at its GitHub repository:
-\texttt{https://github.com/Sixtease/MakonReact}, specifically in the files
-\texttt{src/store/audio.js} and \texttt{src/store/AudioChunks.js}.
+\url{https://github.com/Sixtease/MakonReact}, specifically in the files
+\url{src/store/audio.js} and \url{src/store/AudioChunks.js}.
 
@@
 
-The new segmented structuring is used in the web application to emulate streaming capability while exploiting the advantages offered by Web Audio API.
-Other uses of the corpus, like direct downloads, archiving and on-demand forced alignment still rely on the original structuring where one recording of length in order of tens of minutes corresponds to a single file.
+The new segmented structuring is used in the web application to emulate
+streaming capability while exploiting the advantages offered by Web Audio API.
+Other uses of the corpus, like direct downloads, archiving and on-demand forced
+alignment still rely on the original structuring where one recording of length
+in order of tens of minutes corresponds to a single file.
 
-The redundant storage needs impose a negligible financial overhead but the work necessary for creation and maintenance of these extra files bring about an even stronger wish for the Web Audio API to support streaming.
-However, storing the corpus in files of a smaller size and devising the mechanism to seamlessly glue the chunks together during playback and transcription was a worthy lesson with potential use in other settings.
+The redundant storage needs impose a negligible financial overhead but the work
+necessary for creation and maintenance of these extra files bring about an even
+stronger wish for the Web Audio API to support streaming. However, storing the
+corpus in files of a smaller size and devising the mechanism to seamlessly glue
+the chunks together during playback and transcription was a worthy lesson with
+potential use in other settings.
 
